---
title: "Build priority land-use rasters for Denmark and Silkeborg"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  basemap_file: "Datasets/lu_01_2018.tif"
  basemap_vat_file: "Datasets/lu_01_2018.tif.vat.dbf"
  basemap_reclass_file: "Datasets/Basemap_all_cat_reclass_SN_05_2023.csv"
  subclasses_file: "Datasets/Subclasses.tif"
  markblokkort_file: "Datasets/Rast_markblokkort_Croped.tif"
  markblokkort_field: "MB_TYPE"
  silkeborg_file: "Datasets/Silkeborg_kommune.shp"
  output_dir: "Predictors/silkeborg_landuse"
  forest_nature_rule: "subclasses_skovnatur"
  overwrite: false
---

<!-- Edit this Rmd, then knit it. The code is also safe to run chunk by chunk. -->

# Purpose

This workflow combines four sources into one categorical land-use system:

1. the eight `SN_ModelClass` nature types derived from Basemap03;
2. agriculture from Markblokkort;
3. urban and infrastructure classes from Basemap03; and
4. production forest from the **national Basemap forest footprint**.

The last point is deliberate. `Subclasses == "Drevet Skov"` only identifies
managed forest inside the protection-scheme footprint. It is useful as an
audit, but it cannot be the source for production forest across the whole
country. Here, all Basemap forest (`311000` and `312000`) is treated as
production forest unless it is explicitly retained as forest nature.

It writes a national current raster for model fitting and threshold calibration,
then crops that exact raster to Silkeborg for current and scenario projections.
The two rasters therefore use the same codes and factor levels.

The priority order is:

1. nature type;
2. agriculture;
3. urban/infrastructure;
4. production forest; and
5. other land.

Thus, a natural habitat is not overwritten merely because it intersects a
Markblokkort polygon. Agriculture still takes priority over production forest,
and urban land is never included in a restoration scenario.

# Input folder

Place the source files under `Datasets/` using the parameter names above. The
Basemap value-attribute table supplied with `lu_01_2018.tif` is essential: its
`C_19` field provides the stable aggregated classes used for urban and forest.

`markblokkort_file` may be either:

- a categorical raster whose cells are `INT_AGG`, `PGR`, or the original
  Markblokkort classes; or
- a vector file with the class field named by `markblokkort_field`.

The categorical baseline is built nationally because model fitting should use
the Danish occurrence records and a national predictor domain. Only the final
projection copy is cropped to Silkeborg. CHELSA is **not** upscaled nationally;
the second document aligns climate at 10 m only for the small municipality.

# Data provenance and redistribution

Basemap03 is an Aarhus University/DCE analytical product and is not legally
binding. Cite the Basemap03 technical documentation (Levin 2019) in analyses
that use this derivative. The derived rasters, Markblokkort data and project
`Subclasses` layer must retain their respective source terms and internal data
agreements. In particular, do not add large or restricted source rasters to a
public Git repository merely because this script can reproduce the derivative;
track paths/manifests and documentation in Git, and store the spatial inputs in
the approved data location.

```{r setup}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  message = TRUE,
  warning = TRUE
)

required_packages <- c(
  "terra", "foreign", "dplyr", "readr", "stringr", "tibble", "knitr"
)

missing_packages <- setdiff(
  required_packages,
  rownames(utils::installed.packages())
)

if (length(missing_packages) > 0L) {
  stop(
    "Install the missing packages first: ",
    paste(missing_packages, collapse = ", "),
    call. = FALSE
  )
}

library(terra)
library(dplyr)

terra::terraOptions(progress = 1L, threads = 1L)
```

# Fixed class dictionary

The IDs are fixed rather than generated from first-seen order. This guarantees
that current and scenario rasters have exactly the same interpretation.

```{r class-dictionary}
landuse_legend <- tibble::tribble(
  ~id, ~class,                  ~role,
    1, "ForestDryPoor",        "nature",
    2, "ForestWetPoor",        "nature",
    3, "ForestDryRich",        "nature",
    4, "ForestWetRich",        "nature",
    5, "OpenDryPoor",          "nature",
    6, "OpenWetPoor",          "nature",
    7, "OpenDryRich",          "nature",
    8, "OpenWetRich",          "nature",
    9, "Agriculture",          "convertible",
   10, "Urban",                "fixed",
   11, "ProductionForest",     "convertible",
   12, "Other",                "fixed"
)

nature_codes <- stats::setNames(
  landuse_legend$id[landuse_legend$role == "nature"],
  landuse_legend$class[landuse_legend$role == "nature"]
)

code_for <- function(class_name) {
  out <- landuse_legend$id[match(class_name, landuse_legend$class)]
  if (length(out) != 1L || is.na(out)) {
    stop("Unknown land-use class: ", class_name, call. = FALSE)
  }
  out
}

knitr::kable(landuse_legend)
```

# Helpers

```{r helpers}
assert_files_exist <- function(paths) {
  missing <- paths[!file.exists(paths)]
  if (length(missing) > 0L) {
    stop(
      "Missing input files:\n- ",
      paste(missing, collapse = "\n- "),
      call. = FALSE
    )
  }
  invisible(paths)
}

align_categorical <- function(x, template) {
  x <- terra::crop(x, terra::ext(template), snap = "out")

  if (!terra::compareGeom(x, template, stopOnError = FALSE)) {
    x <- terra::project(x, template, method = "near")
  }

  if (!terra::compareGeom(x, template, stopOnError = FALSE)) {
    stop("A categorical input could not be aligned to Basemap.", call. = FALSE)
  }

  terra::mask(x, template)
}

count_true <- function(x) {
  as.numeric(
    terra::global(
      terra::ifel(!is.na(x) & x, 1, 0),
      fun = "sum",
      na.rm = TRUE
    )[1, 1]
  )
}

write_landuse_raster <- function(x, filename, overwrite = FALSE) {
  dir.create(dirname(filename), recursive = TRUE, showWarnings = FALSE)

  terra::writeRaster(
    x,
    filename = filename,
    overwrite = overwrite,
    datatype = "INT1U",
    NAflag = 255,
    gdal = c(
      "COMPRESS=DEFLATE",
      "PREDICTOR=2",
      "TILED=YES",
      "BIGTIFF=IF_SAFER"
    )
  )

  filename
}

read_agriculture_mask <- function(
    path,
    template,
    field = "MB_TYPE",
    accepted = c("OMD", "PGR", "ING", "PAF", "MIX", "VKS", "INT_AGG")) {
  raster_extensions <- c(".tif", ".tiff", ".vrt", ".img")
  extension <- tolower(tools::file_ext(path))

  if (paste0(".", extension) %in% raster_extensions) {
    agriculture <- align_categorical(terra::rast(path), template)
    category_tables <- terra::levels(agriculture)
    category_table <- if (length(category_tables) > 0L) {
      category_tables[[1]]
    } else {
      NULL
    }

    if (!is.null(category_table) && ncol(category_table) >= 2L) {
      return(agriculture %in% accepted)
    }

    # Numeric Markblokkort rasters are commonly 1/NA or 1/0. Zero is treated
    # as background; any other finite value is agricultural land.
    return(!is.na(agriculture) & agriculture != 0)
  }

  agriculture <- terra::vect(path)
  if (!field %in% names(agriculture)) {
    stop(
      "The Markblokkort vector must contain `", field, "`.",
      call. = FALSE
    )
  }

  agriculture <- agriculture[agriculture[[field]] %in% accepted, ]
  if (nrow(agriculture) == 0L) {
    stop("No accepted Markblokkort polygons were found.", call. = FALSE)
  }

  agriculture <- terra::project(agriculture, terra::crs(template))
  terra::rasterize(agriculture, template, field = 1, background = 0) == 1
}
```

# Validate and read inputs

```{r read-inputs}
assert_files_exist(c(
  params$basemap_file,
  params$basemap_vat_file,
  params$basemap_reclass_file,
  params$subclasses_file,
  params$markblokkort_file,
  params$silkeborg_file
))

basemap <- terra::rast(params$basemap_file)
if (terra::nlyr(basemap) != 1L) {
  stop("The Basemap input must contain exactly one layer.", call. = FALSE)
}

silkeborg <- terra::vect(params$silkeborg_file)
silkeborg <- terra::makeValid(silkeborg)
silkeborg <- terra::aggregate(silkeborg)
silkeborg <- terra::project(silkeborg, terra::crs(basemap))

basemap_template <- basemap

names(basemap_template) <- "Basemap_value"

vat <- foreign::read.dbf(params$basemap_vat_file, as.is = TRUE)
reclass <- utils::read.csv(
  params$basemap_reclass_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

required_vat_columns <- c("Value", "C_19")
required_reclass_columns <- c("Value1", "SN_ModelClass")

if (!all(required_vat_columns %in% names(vat))) {
  stop(
    "The Basemap VAT must contain: ",
    paste(required_vat_columns, collapse = ", "),
    call. = FALSE
  )
}

if (!all(required_reclass_columns %in% names(reclass))) {
  stop(
    "The reclassification CSV must contain: ",
    paste(required_reclass_columns, collapse = ", "),
    call. = FALSE
  )
}
```

# Join the Basemap tables

The Basemap03 VAT supplied for this project uses the following aggregate codes:

- `110000`--`160000`: urban, roads, infrastructure, recreation and extraction;
- `311000` and `312000`: forest and wet forest.

Agriculture is taken from Markblokkort, not inferred from these codes.

```{r join-basemap-tables}
reclass_clean <- reclass |>
  dplyr::transmute(
    Value1 = as.numeric(.data$Value1),
    SN_ModelClass = stringr::str_trim(
      sub("Both$", "Rich", .data$SN_ModelClass)
    )
  )

rat <- vat |>
  dplyr::mutate(
    Value = as.numeric(.data$Value),
    C_19 = as.integer(.data$C_19)
  ) |>
  dplyr::left_join(reclass_clean, by = c("Value" = "Value1"))

duplicate_labels <- rat |>
  dplyr::filter(!is.na(.data$SN_ModelClass)) |>
  dplyr::distinct(.data$Value, .data$SN_ModelClass) |>
  dplyr::count(.data$Value) |>
  dplyr::filter(.data$n > 1L)

if (nrow(duplicate_labels) > 0L) {
  stop("Some Basemap values have more than one SN model class.", call. = FALSE)
}

unknown_nature_labels <- setdiff(
  unique(stats::na.omit(rat$SN_ModelClass)),
  c(names(nature_codes), "Exclude")
)

if (length(unknown_nature_labels) > 0L) {
  stop(
    "Unrecognised SN_ModelClass labels: ",
    paste(unknown_nature_labels, collapse = ", "),
    call. = FALSE
  )
}
```

# Build source masks

Open nature is accepted wherever the SN reclassification identifies an open
nature type. Forest nature is more conservative: by default it requires
`Subclasses == "Skovnatur"`. Every remaining Basemap forest cell is available
to the production-forest class. Set `forest_nature_rule` to `all_sn_forest` only
if the SN forest classes have independently been verified to exclude managed
forest.

```{r source-masks}
nature_lookup <- rat |>
  dplyr::transmute(
    old = .data$Value,
    new = unname(nature_codes[.data$SN_ModelClass])
  ) |>
  dplyr::filter(!is.na(.data$old), !is.na(.data$new)) |>
  dplyr::distinct()

if (nrow(nature_lookup) == 0L) {
  stop("No Basemap values mapped to the eight nature types.", call. = FALSE)
}

nature_candidate <- terra::classify(
  basemap_template,
  rcl = as.matrix(nature_lookup),
  others = NA
)

subclasses <- params$subclasses_file |>
  terra::rast() |>
  align_categorical(basemap_template)

agriculture_mask <- read_agriculture_mask(
  path = params$markblokkort_file,
  template = basemap_template,
  field = params$markblokkort_field
)

forest_basemap_values <- rat |>
  dplyr::filter(.data$C_19 %in% c(311000L, 312000L)) |>
  dplyr::pull(.data$Value) |>
  unique()

urban_basemap_values <- rat |>
  dplyr::filter(dplyr::between(.data$C_19, 110000L, 160000L)) |>
  dplyr::pull(.data$Value) |>
  unique()

forest_mask <- basemap_template %in% forest_basemap_values
urban_mask <- basemap_template %in% urban_basemap_values

open_codes <- landuse_legend$id[
  startsWith(landuse_legend$class, "Open")
]
forest_codes <- landuse_legend$id[
  startsWith(landuse_legend$class, "Forest")
]

open_nature_mask <- nature_candidate %in% open_codes

if (identical(params$forest_nature_rule, "subclasses_skovnatur")) {
  forest_nature_mask <-
    nature_candidate %in% forest_codes &
    subclasses %in% "Skovnatur"
} else if (identical(params$forest_nature_rule, "all_sn_forest")) {
  forest_nature_mask <- nature_candidate %in% forest_codes
} else {
  stop(
    "`forest_nature_rule` must be `subclasses_skovnatur` or ",
    "`all_sn_forest`.",
    call. = FALSE
  )
}

nature_mask <- open_nature_mask | forest_nature_mask
nature_layer <- terra::ifel(nature_mask, nature_candidate, NA)
```

# Apply the priority hierarchy

Assignments are made from lowest to highest priority. The final `cover()` puts
the eight nature classes on top.

```{r priority-landuse}
denmark_domain_values <- rat |>
  dplyr::filter(
    !.data$C_19 %in% c(
      420000L, # sea
      800000L, # unmapped
      999999L  # Germany
    )
  ) |>
  dplyr::pull(.data$Value) |>
  unique()

domain <- basemap_template %in% denmark_domain_values

landuse <- terra::ifel(domain, code_for("Other"), NA)
landuse <- terra::ifel(
  forest_mask,
  code_for("ProductionForest"),
  landuse
)
landuse <- terra::ifel(urban_mask, code_for("Urban"), landuse)
landuse <- terra::ifel(
  agriculture_mask,
  code_for("Agriculture"),
  landuse
)
landuse <- terra::cover(nature_layer, landuse)
names(landuse) <- "Landuse"

class_table <- as.data.frame(landuse_legend[c("id", "class")])
terra::levels(landuse) <- class_table

landuse_silkeborg <- landuse |>
  terra::crop(silkeborg, snap = "out") |>
  terra::mask(silkeborg)
names(landuse_silkeborg) <- "Landuse"
terra::levels(landuse_silkeborg) <- class_table
```

# Validation and the production-forest check

The audit explicitly reports how much production forest would have been found
from `Subclasses == "Drevet Skov"` and how much is found from Basemap. A large
difference is expected and is the reason for the revised method.

```{r validate}
if (!terra::compareGeom(landuse, basemap_template, stopOnError = FALSE)) {
  stop("The final land-use raster lost the Basemap geometry.", call. = FALSE)
}

observed_ids <- terra::freq(as.numeric(landuse), value = TRUE)
if (!all(observed_ids$value %in% landuse_legend$id)) {
  stop("The final raster contains an unknown class ID.", call. = FALSE)
}

subclasses_managed_forest <- subclasses %in% "Drevet Skov"
forest_outside_subclasses <- forest_mask & is.na(subclasses)

source_audit <- tibble::tibble(
  measure = c(
    "all_basemap_forest_cells",
    "subclasses_drevet_skov_cells",
    "basemap_forest_outside_subclasses_cells",
    "final_production_forest_cells",
    "final_agriculture_cells",
    "final_urban_cells",
    "final_nature_cells"
  ),
  cells = c(
    count_true(forest_mask),
    count_true(subclasses_managed_forest),
    count_true(forest_outside_subclasses),
    count_true(landuse == code_for("ProductionForest")),
    count_true(landuse == code_for("Agriculture")),
    count_true(landuse == code_for("Urban")),
    count_true(landuse %in% landuse_legend$id[landuse_legend$role == "nature"])
  )
)

if (source_audit$cells[source_audit$measure ==
    "basemap_forest_outside_subclasses_cells"] == 0) {
  warning(
    "No Basemap forest was found outside Subclasses. Confirm that ",
    "Subclasses is really restricted to protection schemes.",
    call. = FALSE
  )
}

cell_area_km2 <- terra::cellSize(landuse, unit = "km")
area_by_class <- terra::zonal(
  cell_area_km2,
  as.numeric(landuse),
  fun = "sum",
  na.rm = TRUE
)
names(area_by_class) <- c("id", "area_km2")

class_audit <- landuse_legend |>
  dplyr::left_join(area_by_class, by = "id") |>
  dplyr::mutate(area_km2 = dplyr::coalesce(.data$area_km2, 0))

silkeborg_cell_area_km2 <- terra::cellSize(
  landuse_silkeborg,
  unit = "km"
)
silkeborg_area_by_class <- terra::zonal(
  silkeborg_cell_area_km2,
  as.numeric(landuse_silkeborg),
  fun = "sum",
  na.rm = TRUE
)
names(silkeborg_area_by_class) <- c("id", "area_km2")

silkeborg_class_audit <- landuse_legend |>
  dplyr::left_join(silkeborg_area_by_class, by = "id") |>
  dplyr::mutate(area_km2 = dplyr::coalesce(.data$area_km2, 0))

knitr::kable(source_audit)
knitr::kable(class_audit, digits = 3)
knitr::kable(silkeborg_class_audit, digits = 3)
```

# Write outputs

```{r write-outputs}
dir.create(params$output_dir, recursive = TRUE, showWarnings = FALSE)

national_landuse_path <- file.path(
  params$output_dir,
  "Landuse_Denmark_current.tif"
)
silkeborg_landuse_path <- file.path(
  params$output_dir,
  "Landuse_Silkeborg_current.tif"
)
legend_path <- file.path(params$output_dir, "Landuse_class_legend.csv")
source_audit_path <- file.path(
  params$output_dir,
  "Landuse_source_audit.csv"
)
class_audit_path <- file.path(
  params$output_dir,
  "Landuse_Denmark_class_area_audit.csv"
)
silkeborg_class_audit_path <- file.path(
  params$output_dir,
  "Landuse_Silkeborg_class_area_audit.csv"
)
national_path_manifest <- file.path(
  params$output_dir,
  "landuse_Denmark_current_path.rds"
)
silkeborg_path_manifest <- file.path(
  params$output_dir,
  "landuse_Silkeborg_current_path.rds"
)

write_landuse_raster(
  landuse,
  filename = national_landuse_path,
  overwrite = params$overwrite
)
write_landuse_raster(
  landuse_silkeborg,
  filename = silkeborg_landuse_path,
  overwrite = params$overwrite
)
readr::write_csv(landuse_legend, legend_path)
readr::write_csv(source_audit, source_audit_path)
readr::write_csv(class_audit, class_audit_path)
readr::write_csv(silkeborg_class_audit, silkeborg_class_audit_path)
saveRDS(national_landuse_path, national_path_manifest)
saveRDS(silkeborg_landuse_path, silkeborg_path_manifest)

c(
  national_landuse = national_landuse_path,
  silkeborg_landuse = silkeborg_landuse_path,
  legend = legend_path,
  source_audit = source_audit_path,
  national_class_audit = class_audit_path,
  silkeborg_class_audit = silkeborg_class_audit_path,
  national_manifest = national_path_manifest,
  silkeborg_manifest = silkeborg_path_manifest
)
```

# Quick map

```{r plot, fig.width=10, fig.height=7}
plot(
  landuse_silkeborg,
  type = "classes",
  main = "Priority land use — Silkeborg"
)
```

# Interpretation

This output is a **current land-use predictor**, not a protection-status map.
`Subclasses` contributes the distinction between forest nature and managed
forest, but Basemap supplies the complete forest footprint. Before modelling,
inspect the audit and map, especially boundaries between `Forest*` and
`ProductionForest`. If a separate authoritative unmanaged/natural-forest layer
becomes available, merge it into `forest_nature_mask`; do not revert to using
`Drevet Skov` alone for production forest.
