---
title: "Build eight Silkeborg nature-restoration scenarios"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  national_current_landuse_file: "Predictors/silkeborg_landuse/Landuse_Denmark_current.tif"
  current_landuse_file: "Predictors/silkeborg_landuse/Landuse_Silkeborg_current.tif"
  landuse_legend_file: "Predictors/silkeborg_landuse/Landuse_class_legend.csv"
  silkeborg_file: "Datasets/Silkeborg_kommune.shp"
  current_climate_file: ""
  current_predictor_manifest_file: "Predictors/current_predictor_paths.rds"
  output_dir: "Predictors/silkeborg_scenarios"
  overwrite: false
---

# Purpose

This workflow starts from the priority land-use raster created by
`01_build_silkeborg_priority_landuse.Rmd`. It creates eight counterfactual
rasters. In each one, **every Agriculture and ProductionForest cell** is changed
to one of the eight nature types.

Existing nature, urban and other cells remain unchanged. The workflow also:

- aligns the existing current CHELSA stack to the Silkeborg 10 m template;
- writes separate national-training and local-projection manifests;
- writes a scenario manifest in the list-column format used by
  `SpeciesPoolR::load_one_scenario()`; and
- audits geometry and the exact set of changed cells.

These are deliberately extreme, single-endpoint scenarios. They answer “what
if every convertible production cell became this nature type?” They do not yet
encode feasibility, restoration time, dispersal, succession or partial uptake.

The national/local split is essential. `FitSpeciesModels()` should use national
land use and national CHELSA so Danish occurrence records remain inside the
predictor domain. The national 10 m land-use raster is reduced to the existing
CHELSA grid with nearest-neighbour resampling; CHELSA is not needlessly
upsampled across all Denmark. Only current/scenario projections are restricted
to Silkeborg. Do not replace the training predictors with the local manifest
unless the scientific intention is to fit from Silkeborg records alone.

```{r setup}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  message = TRUE,
  warning = TRUE
)

required_packages <- c("terra", "dplyr", "readr", "tibble", "knitr")
missing_packages <- setdiff(
  required_packages,
  rownames(utils::installed.packages())
)

if (length(missing_packages) > 0L) {
  stop(
    "Install the missing packages first: ",
    paste(missing_packages, collapse = ", "),
    call. = FALSE
  )
}

library(terra)
library(dplyr)

# Parallelism belongs at the targets/crew level. Each raster worker should use
# one GDAL/OpenMP thread to avoid nested oversubscription on the server.
Sys.setenv(
  OMP_NUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1"
)
terra::terraOptions(progress = 1L, threads = 1L)
```

# Helpers

```{r helpers}
assert_files_exist <- function(paths) {
  paths <- unique(as.character(paths))
  paths <- paths[!is.na(paths) & nzchar(paths)]
  missing <- paths[!file.exists(paths)]

  if (length(missing) > 0L) {
    stop(
      "Missing input files:\n- ",
      paste(missing, collapse = "\n- "),
      call. = FALSE
    )
  }

  invisible(paths)
}

snake_case <- function(x) {
  x <- gsub("([a-z0-9])([A-Z])", "\\1_\\2", x)
  x <- gsub("[^A-Za-z0-9]+", "_", x)
  tolower(gsub("^_+|_+$", "", x))
}

count_true <- function(x) {
  as.numeric(
    terra::global(
      terra::ifel(!is.na(x) & x, 1, 0),
      fun = "sum",
      na.rm = TRUE
    )[1, 1]
  )
}

sum_where <- function(values, condition) {
  as.numeric(
    terra::global(
      terra::ifel(condition, values, NA),
      fun = "sum",
      na.rm = TRUE
    )[1, 1]
  )
}

write_landuse_raster <- function(x, filename, overwrite = FALSE) {
  dir.create(dirname(filename), recursive = TRUE, showWarnings = FALSE)

  terra::writeRaster(
    x,
    filename = filename,
    overwrite = overwrite,
    datatype = "INT1U",
    NAflag = 255,
    gdal = c(
      "COMPRESS=DEFLATE",
      "PREDICTOR=2",
      "TILED=YES",
      "BIGTIFF=IF_SAFER"
    )
  )

  filename
}

is_landuse_raster <- function(path) {
  x <- terra::rast(path)
  category_tables <- terra::levels(x)
  has_categories <- length(category_tables) > 0L &&
    !is.null(category_tables[[1]]) &&
    nrow(category_tables[[1]]) > 0L

  any(tolower(names(x)) == "landuse") ||
    (terra::nlyr(x) == 1L && has_categories)
}

resolve_current_climate <- function(
    explicit_path,
    predictor_manifest,
    current_landuse_path) {
  if (!is.null(explicit_path) && nzchar(explicit_path)) {
    assert_files_exist(explicit_path)
    return(explicit_path)
  }

  assert_files_exist(predictor_manifest)
  manifest_paths <- unique(unlist(readRDS(predictor_manifest), use.names = FALSE))
  manifest_paths <- manifest_paths[
    !is.na(manifest_paths) & nzchar(manifest_paths)
  ]
  assert_files_exist(manifest_paths)

  normal_current_landuse <- normalizePath(
    current_landuse_path,
    winslash = "/",
    mustWork = TRUE
  )
  normal_manifest <- normalizePath(
    manifest_paths,
    winslash = "/",
    mustWork = TRUE
  )

  candidates <- manifest_paths[normal_manifest != normal_current_landuse]
  candidates <- candidates[!vapply(candidates, is_landuse_raster, logical(1))]

  if (length(candidates) != 1L) {
    stop(
      "Could not identify exactly one current climate stack in `", 
      predictor_manifest, "`. Set `current_climate_file` explicitly. ",
      "Candidates found: ", length(candidates),
      call. = FALSE
    )
  }

  candidates[[1]]
}
```

# Read current land use and dictionary

```{r read-landuse}
assert_files_exist(c(
  params$national_current_landuse_file,
  params$current_landuse_file,
  params$landuse_legend_file,
  params$silkeborg_file
))

landuse_legend <- readr::read_csv(
  params$landuse_legend_file,
  show_col_types = FALSE
)

required_legend_columns <- c("id", "class", "role")
if (!all(required_legend_columns %in% names(landuse_legend))) {
  stop(
    "The land-use legend must contain: ",
    paste(required_legend_columns, collapse = ", "),
    call. = FALSE
  )
}

expected_nature_classes <- c(
  "ForestDryPoor", "ForestWetPoor", "ForestDryRich", "ForestWetRich",
  "OpenDryPoor", "OpenWetPoor", "OpenDryRich", "OpenWetRich"
)

if (!setequal(
  landuse_legend$class[landuse_legend$role == "nature"],
  expected_nature_classes
)) {
  stop("The legend does not contain the expected eight nature classes.", call. = FALSE)
}

code_for <- function(class_name) {
  out <- landuse_legend$id[match(class_name, landuse_legend$class)]
  if (length(out) != 1L || is.na(out)) {
    stop("Unknown land-use class: ", class_name, call. = FALSE)
  }
  as.integer(out)
}

current_landuse <- terra::rast(params$current_landuse_file)
if (terra::nlyr(current_landuse) != 1L) {
  stop("Current land use must have exactly one layer.", call. = FALSE)
}

names(current_landuse) <- "Landuse"
current_landuse_numeric <- as.numeric(current_landuse)

national_current_landuse <- terra::rast(
  params$national_current_landuse_file
)
if (terra::nlyr(national_current_landuse) != 1L) {
  stop("National current land use must have exactly one layer.", call. = FALSE)
}
names(national_current_landuse) <- "Landuse"

silkeborg <- params$silkeborg_file |>
  terra::vect() |>
  terra::makeValid() |>
  terra::aggregate() |>
  terra::project(terra::crs(current_landuse))

current_landuse_numeric <- terra::mask(
  current_landuse_numeric,
  silkeborg
)
```

# Align current CHELSA to the small template

Only the Silkeborg extent is resampled to 10 m. Climate is continuous, so
bilinear interpolation is used; land use is never bilinearly interpolated.

```{r align-climate}
current_climate_source <- resolve_current_climate(
  explicit_path = params$current_climate_file,
  predictor_manifest = params$current_predictor_manifest_file,
  current_landuse_path = params$current_landuse_file
)

assert_files_exist(current_climate_source)
dir.create(params$output_dir, recursive = TRUE, showWarnings = FALSE)

climate_output_path <- file.path(
  params$output_dir,
  "CHELSA_current_Silkeborg_10m.tif"
)

current_climate_source_raster <- terra::rast(current_climate_source)

# Categorical training land use is placed on the existing national CHELSA
# grid. This is far smaller and more honest than manufacturing national 10 m
# climate values. Nearest neighbour is mandatory for the class IDs.
training_landuse_output_path <- file.path(
  params$output_dir,
  "Landuse_Denmark_training_on_CHELSA_grid.tif"
)

training_landuse <- terra::project(
  national_current_landuse,
  current_climate_source_raster[[1]],
  method = "near"
)
training_landuse <- terra::mask(
  training_landuse,
  current_climate_source_raster[[1]]
)
names(training_landuse) <- "Landuse"
terra::levels(training_landuse) <- as.data.frame(
  landuse_legend[c("id", "class")]
)

write_landuse_raster(
  training_landuse,
  filename = training_landuse_output_path,
  overwrite = params$overwrite
)

if (!terra::compareGeom(
  training_landuse,
  current_climate_source_raster,
  stopOnError = FALSE
)) {
  stop("National training land use does not match CHELSA.", call. = FALSE)
}

# Use a separate temporary file so GDAL never needs to overwrite a raster that
# is still open. This matters especially on Windows servers.
climate_projected_path <- tempfile(
  pattern = "CHELSA_current_Silkeborg_projected_",
  fileext = ".tif"
)

current_climate_projected <- terra::project(
  current_climate_source_raster,
  current_landuse_numeric,
  method = "bilinear",
  filename = climate_projected_path,
  overwrite = TRUE,
  wopt = list(
    datatype = "FLT4S",
    gdal = c(
      "COMPRESS=DEFLATE",
      "PREDICTOR=3",
      "TILED=YES",
      "BIGTIFF=IF_SAFER"
    )
  )
)

current_climate <- terra::mask(
  current_climate_projected,
  current_landuse_numeric,
  filename = climate_output_path,
  overwrite = params$overwrite,
  wopt = list(
    datatype = "FLT4S",
    gdal = c("COMPRESS=DEFLATE", "PREDICTOR=3", "TILED=YES")
  )
)

rm(current_climate_projected)
invisible(gc())
unlink(climate_projected_path)

if (!terra::compareGeom(
  current_climate,
  current_landuse_numeric,
  stopOnError = FALSE
)) {
  stop("The aligned CHELSA stack does not match land use.", call. = FALSE)
}

finite_climate_cells <- terra::global(
  terra::app(!is.na(current_climate), sum) == terra::nlyr(current_climate),
  fun = "sum",
  na.rm = TRUE
)[1, 1]

if (!is.finite(finite_climate_cells) || finite_climate_cells == 0) {
  stop("Aligned CHELSA contains no cells complete across all bands.", call. = FALSE)
}
```

# Scenario specification

The conversion mask contains exactly two current classes. Urban and Other are
fixed, and existing nature cells keep their current class.

```{r scenario-specification}
convertible_ids <- c(
  code_for("Agriculture"),
  code_for("ProductionForest")
)

scenario_table <- landuse_legend |>
  dplyr::filter(.data$role == "nature") |>
  dplyr::transmute(
    scenario = paste0("all_to_", vapply(.data$class, snake_case, character(1))),
    target_id = as.integer(.data$id),
    target_class = .data$class
  ) |>
  dplyr::arrange(.data$target_id)

if (nrow(scenario_table) != 8L) {
  stop("Exactly eight nature scenarios are required.", call. = FALSE)
}

convertible_mask <- current_landuse_numeric %in% convertible_ids
convertible_cells <- count_true(convertible_mask)

if (convertible_cells == 0L) {
  stop(
    "No Agriculture or ProductionForest cells occur in the current map.",
    call. = FALSE
  )
}

cell_area_km2 <- terra::cellSize(current_landuse_numeric, unit = "km")
convertible_area_km2 <- sum_where(cell_area_km2, convertible_mask)

knitr::kable(scenario_table)
```

# Write the eight land-use rasters

```{r build-scenarios}
scenario_directory <- file.path(params$output_dir, "landuse")
dir.create(scenario_directory, recursive = TRUE, showWarnings = FALSE)

scenario_results <- vector("list", nrow(scenario_table))

for (i in seq_len(nrow(scenario_table))) {
  scenario_name <- scenario_table$scenario[[i]]
  target_id <- scenario_table$target_id[[i]]

  scenario_numeric <- terra::ifel(
    convertible_mask,
    target_id,
    current_landuse_numeric
  )
  scenario_raster <- scenario_numeric
  names(scenario_raster) <- "Landuse"
  terra::levels(scenario_raster) <- as.data.frame(
    landuse_legend[c("id", "class")]
  )

  scenario_path <- file.path(
    scenario_directory,
    paste0("Landuse_Silkeborg_", scenario_name, ".tif")
  )

  write_landuse_raster(
    scenario_raster,
    filename = scenario_path,
    overwrite = params$overwrite
  )

  changed <- scenario_numeric != current_landuse_numeric
  changed_outside_convertible <- changed & !convertible_mask
  remaining_convertible <- scenario_numeric %in% convertible_ids

  scenario_results[[i]] <- tibble::tibble(
    scenario = scenario_name,
    target_id = target_id,
    target_class = scenario_table$target_class[[i]],
    changed_cells = count_true(changed),
    expected_changed_cells = convertible_cells,
    changed_area_km2 = sum_where(cell_area_km2, changed),
    expected_changed_area_km2 = convertible_area_km2,
    changed_outside_convertible_cells = count_true(
      changed_outside_convertible
    ),
    remaining_convertible_cells = count_true(remaining_convertible),
    geometry_matches_current = terra::compareGeom(
      scenario_raster,
      current_landuse_numeric,
      stopOnError = FALSE
    ),
    landuse_path = scenario_path
  )
}

scenario_audit <- dplyr::bind_rows(scenario_results)

if (any(
  scenario_audit$changed_cells != scenario_audit$expected_changed_cells |
    scenario_audit$changed_outside_convertible_cells != 0L |
    scenario_audit$remaining_convertible_cells != 0L |
    !scenario_audit$geometry_matches_current
)) {
  stop("At least one land-use scenario failed its audit.", call. = FALSE)
}

area_tolerance <- max(1e-8, convertible_area_km2 * 1e-10)
if (any(abs(
  scenario_audit$changed_area_km2 -
    scenario_audit$expected_changed_area_km2
) > area_tolerance)) {
  stop("At least one scenario changed the wrong area.", call. = FALSE)
}

knitr::kable(
  scenario_audit |>
    dplyr::select(
      .data$scenario,
      .data$target_class,
      .data$changed_cells,
      .data$changed_area_km2,
      .data$changed_outside_convertible_cells,
      .data$remaining_convertible_cells
    ),
  digits = 3
)
```

# Write SpeciesPoolR manifests

The training manifest combines the new national land-use layer with the
original national CHELSA stack. The projection manifest combines local current
land use with local CHELSA. Every restoration scenario uses local current
CHELSA. Later, a future-climate path can replace the second element of any
scenario's `path` vector without changing the land-use files.

```{r write-manifests}
training_current_predictor_paths <- c(
  training_landuse_output_path,
  current_climate_source
)

projection_current_predictor_paths <- c(
  params$current_landuse_file,
  climate_output_path
)

scenario_specs <- data.frame(
  scenario = scenario_audit$scenario,
  stringsAsFactors = FALSE
)
scenario_specs$path <- unname(lapply(
  scenario_audit$landuse_path,
  function(landuse_path) c(landuse_path, climate_output_path)
))

training_manifest_path <- file.path(
  params$output_dir,
  "training_current_predictor_paths_Denmark.rds"
)
projection_current_manifest_path <- file.path(
  params$output_dir,
  "projection_current_predictor_paths_silkeborg.rds"
)
scenario_manifest_path <- file.path(
  params$output_dir,
  "mixed_scenario_specs_silkeborg.rds"
)
scenario_audit_path <- file.path(
  params$output_dir,
  "landuse_scenario_audit.csv"
)
scenario_table_path <- file.path(
  params$output_dir,
  "landuse_scenario_table.csv"
)

saveRDS(training_current_predictor_paths, training_manifest_path)
saveRDS(
  projection_current_predictor_paths,
  projection_current_manifest_path
)
saveRDS(scenario_specs, scenario_manifest_path)
readr::write_csv(scenario_audit, scenario_audit_path)
readr::write_csv(scenario_table, scenario_table_path)

c(
  training_manifest = training_manifest_path,
  projection_current_manifest = projection_current_manifest_path,
  scenario_manifest = scenario_manifest_path,
  scenario_audit = scenario_audit_path,
  scenario_table = scenario_table_path,
  climate = climate_output_path
)
```

# Final checks

```{r final-checks}
stopifnot(
  length(training_current_predictor_paths) == 2L,
  all(file.exists(training_current_predictor_paths)),
  length(projection_current_predictor_paths) == 2L,
  all(file.exists(projection_current_predictor_paths)),
  nrow(scenario_specs) == 8L,
  all(vapply(scenario_specs$path, length, integer(1)) == 2L),
  all(file.exists(unlist(scenario_specs$path, use.names = FALSE)))
)

manifest_audit <- tibble::tibble(
  training_predictor_files = length(training_current_predictor_paths),
  projection_current_predictor_files = length(
    projection_current_predictor_paths
  ),
  scenarios = nrow(scenario_specs),
  convertible_cells = convertible_cells,
  convertible_area_km2 = convertible_area_km2,
  climate_layers = terra::nlyr(current_climate),
  complete_climate_cells = as.numeric(finite_climate_cells)
)

knitr::kable(manifest_audit, digits = 3)
```

# Use in the targets pipeline

The existing climate pipeline uses one `Current_predictors` object for both
model fitting and current projection. For this experiment it should be split:

- `Training_current_predictors`: national land use + national current CHELSA;
- `Projection_current_predictors`: Silkeborg land use + Silkeborg current
  CHELSA; and
- `Scenario_predictors`: the eight local scenario pairs.

Point new file targets to the manifests:

```r
targets::tar_target(
  Study_area_file,
  "Datasets/Silkeborg_kommune.shp",
  format = "file"
)

targets::tar_target(
  Training_predictor_manifest_file,
  "Predictors/silkeborg_scenarios/training_current_predictor_paths_Denmark.rds",
  format = "file"
)

targets::tar_target(
  Projection_current_predictor_manifest_file,
  "Predictors/silkeborg_scenarios/projection_current_predictor_paths_silkeborg.rds",
  format = "file"
)

targets::tar_target(
  Scenario_manifest_file,
  "Predictors/silkeborg_scenarios/mixed_scenario_specs_silkeborg.rds",
  format = "file"
)
```

Use `Training_current_predictors` in `FitSpeciesModels()`. Thresholds must also
be calibrated from national current predictions (or from an equivalent,
validated occurrence-level calculation) before those thresholds are applied to
the local current and scenario projections. This requires a small pipeline
refactor; simply pointing the old `Current_predictor_manifest_file` at the local
manifest would silently discard most occurrence records.

Keep `categorical = "Landuse"`. Start the server run with 20--30 crew workers,
one `terra`/BLAS thread per worker, and increase only after observing RAM, disk
throughput and worker recycling. With the domain restricted to Silkeborg, disk
I/O is likely to become limiting before the available terabyte of RAM.

# Plot current and one scenario

```{r plot, fig.width=11, fig.height=6}
example_scenario <- terra::rast(scenario_audit$landuse_path[[1]])

plot(
  c(current_landuse, example_scenario),
  type = "classes",
  main = c("Current", scenario_audit$scenario[[1]])
)
```
