---
title: "Prepare area for optimization"
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```

## Objective 

The objective of this document is to prepare the dataset for the optimization, the first thing that we will do is to load the needed libraries

```{r loadpackages}
library(terra)
library(SpeciesPoolR)
library(purrr)
library(tidyterra)
```


### Deliminting the space

First we will take the Polygon of the Gudenaa plus the polygons made by the workshop of potential areas drawn by experts to be the potential areas to be restored and added to current natural areas, and transform them into one big polygon

```{r generate-polygon, eval =FALSE}
Gudenaa <- terra::vect("Datasets/Gudenåen_hovedløb/HovedlobMarts24.shp") |>  terra::makeValid()

SelectedAreas <- list.files(
  path = "Datasets/Selected_Areas/",
  full.names = TRUE
) |>
  purrr::map(terra::vect) |>
  purrr::map(terra::makeValid) |> 
  terra::vect() |>
  terra::aggregate()

GudenaaDomain <- terra::vect(
  list(Gudenaa, SelectedAreas)
) |>
  terra::aggregate()

writeVector(GudenaaDomain, "Datasets/GudenaaDomain.shp")
```


This area is a total of `r prettyNum(expanse(GudenaaDomain, unit = "ha"), big.mark = ",")`, We are aiming a at selecting for natural areas a total of 25,000 hectares which would be around `r round(25000/expanse(GudenaaDomain, unit = "ha")*100, 2)` percent of the total area. However of course not all cells are elegible, so first we will determine the total area of current nature, and the total area of selectable cells, in order to do that, we will crop and mask the greater landuse of Gudenna to the domain, and calculate the number of cells in those categoires

```{r}
GudenaaDomain <- terra::vect("Datasets/GudenaaDomain.shp")
GudenaaLanduse <- terra::rast("Predictors/gudenaa_landuse/Landuse_Denmark_current.tif") |> 
  terra::project(terra::crs(GudenaaDomain), method = "near") |> 
  terra::crop(GudenaaDomain) |> 
  terra::mask(GudenaaDomain)
```


