---
title: "New Gudenå area: anchored 200 m hard-network refinement"
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Purpose and exact interpretation

This is the second-stage 200 m optimization. It does not replace or overwrite
the successful existing-adjacency warm-up. Instead, it reads that solution,
turns it into a complete feasible MIP start, and adds the requested hard rules:

1. current nature plus restoration is approximately 30% in **each** of the
   four subareas (and therefore 30% in the complete polygon);
2. at least one connected final-nature area contains 5,000 ha;
3. each subarea has at least 1,000 ha of one connected final-nature area inside
   that subarea;
4. every certified area contains new restoration and has exactly one root in
   an existing-nature patch;
5. connecting several distinct existing-nature patches in the certified
   5,000 ha network receives an additional normalized objective bonus; and
6. a production-forest-dominated 200 m unit can receive only a forest-nature
   destination.

Network area means **current plus newly restored nature**. Habitat types may
connect to each other; connectivity is not restricted to one habitat label.
The five certificates may overlap, so one physical cross-boundary component
may satisfy the global and several subarea requirements. This implements
"at least one" without silently requiring five disjoint reserves.

The hard optimization uses the established 200 m rook graph plus exact 10 m
contacts to current-nature patch supernodes. After the solve, this notebook
independently reconstructs final nature at 10 m and audits the same 5,000/1,000
ha rules. A failed 10 m audit means the coarse solution needs a targeted 10 m
refinement; it must not be reported as an exact fine-resolution network.

# Packages

```{r packages}
library(terra)
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(digest)
library(igraph)
library(knitr)
```

# Editable configuration

Only this section should normally need editing.

```{r configuration}
# ---- Spatial inputs ------------------------------------------------------
fine_landuse_path <- "Datasets/FinalLanduse.tif"
coarse_landuse_path <- file.path(
  "Predictors",
  "gudenaa_optimization",
  "Landuse_Gudenaa_200m_hierarchical.tif"
)
subareas_path <- file.path(
  "Datasets",
  "gudenaaplanen_afgreansning_15sep_delområder.shp"
)

subarea_label_field <- NULL
expected_number_of_subareas <- 4L
expected_subarea_areas_ha <- c(
  7991.14551899414,
  8653.25670292969,
  13009.8300014893,
  14526.539438208
)
expected_total_area_ha <- 44180.77
polygon_area_audit_tolerance_ha <- 5

# Needed to recreate exactly the biodiversity-supported eligibility mask used
# by notebook 07.
metrics_dir <- file.path(
  "Results",
  "gudenaa_plant_biodiversity",
  "full",
  "metrics"
)
biodiversity_support_metric <- "Richness"
biodiversity_support_scenario <- "current"

# ---- Upstream successful warm-up ----------------------------------------
warmup_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_30pct_existing_adjacency"
)
warmup_name <- "gudenaa_new_area_200m_30pct_existing_adjacency"

# ---- Policy --------------------------------------------------------------
global_network_minimum_ha <- 5000
subarea_network_minimum_ha <- 1000

# Keep the same area tolerance and spatial weights used by the warm-up.
spatial_contiguity_bonus <- 1
new_to_new_adjacency_weight <- 0.5
existing_to_new_adjacency_weight <- 0.5

# Additional preference for joining several distinct existing-nature patches
# inside the certified global network. This is a normalized soft reward, not a
# hard requirement to connect a particular number of patches.
connected_existing_patch_bonus <- 0.25

# Established coarse rule: when at least half of a cell's eligible area is
# ProductionForest, the entire 200 m decision receives a forest destination.
production_forest_share_threshold <- 0.50

# ---- Outputs and solver --------------------------------------------------
problem_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_30pct_hard_networks"
)
problem_name <- "gudenaa_new_area_200m_30pct_hard_networks"
model_path <- "Gudenaa_30pct_anchored_networks_200m.mod"

solver_name <- "gurobi"
solver_options <- paste(
  "lim:time=21600",
  "mip:gap=0.005",
  "tech:outlev=1",
  paste0(
    "tech:logfile=",
    file.path(problem_dir, paste0(problem_name, "_gurobi.log"))
  ),
  collapse = " "
)

reuse_matching_checkpoints <- TRUE
```

# Names and validation

```{r names-and-validation}
nature_landuses <- c(
  "ForestDryPoor",
  "ForestDryRich",
  "ForestWetPoor",
  "ForestWetRich",
  "OpenDryPoor",
  "OpenDryRich",
  "OpenWetPoor",
  "OpenWetRich"
)
forest_landuses <- c(
  "ForestDryPoor",
  "ForestDryRich",
  "ForestWetPoor",
  "ForestWetRich"
)

required_networks <- c(
  "Global",
  paste0("Subarea", seq_len(expected_number_of_subareas))
)
network_minimums <- c(
  Global = global_network_minimum_ha,
  stats::setNames(
    rep(subarea_network_minimum_ha, expected_number_of_subareas),
    paste0("Subarea", seq_len(expected_number_of_subareas))
  )
)

stopifnot(
  length(nature_landuses) == 8L,
  length(forest_landuses) == 4L,
  all(forest_landuses %in% nature_landuses),
  global_network_minimum_ha > 0,
  subarea_network_minimum_ha > 0,
  spatial_contiguity_bonus >= 0,
  connected_existing_patch_bonus >= 0,
  new_to_new_adjacency_weight >= 0,
  existing_to_new_adjacency_weight >= 0,
  abs(
    new_to_new_adjacency_weight +
      existing_to_new_adjacency_weight - 1
  ) < 1e-10,
  production_forest_share_threshold > 0,
  production_forest_share_threshold <= 1
)
```

# Reusable helpers

```{r helpers}
shapefile_components <- function(path) {
  if (tolower(tools::file_ext(path)) != "shp") return(path)
  stem <- tools::file_path_sans_ext(path)
  paste0(stem, c(".shp", ".shx", ".dbf", ".prj"))
}

stop_for_missing_files <- function(paths, description) {
  paths <- unique(paths)
  missing <- paths[!file.exists(paths)]
  if (length(missing)) {
    stop(
      paste0(
        "Missing ", description, ":\n- ",
        paste(missing, collapse = "\n- ")
      ),
      call. = FALSE
    )
  }
  invisible(paths)
}

same_crs <- function(x, y) isTRUE(terra::same.crs(x, y))

align_to_template <- function(x, template, method) {
  if (!same_crs(x, template)) {
    x <- terra::project(x, template, method = method)
  } else if (!isTRUE(terra::compareGeom(
    x,
    template,
    lyrs = FALSE,
    stopOnError = FALSE
  ))) {
    x <- terra::resample(x, template, method = method)
  }
  terra::mask(x, template)
}

category_codes <- function(x, requested_labels) {
  lev <- terra::levels(x)[[1]]
  if (is.null(lev) || !nrow(lev)) {
    stop("The fine land-use raster has no category table.", call. = FALSE)
  }

  numeric_columns <- names(lev)[vapply(lev, is.numeric, logical(1))]
  label_columns <- names(lev)[vapply(
    lev,
    function(z) is.character(z) || is.factor(z),
    logical(1)
  )]
  if (!length(numeric_columns) || !length(label_columns)) {
    stop("Could not identify the land-use category columns.", call. = FALSE)
  }

  labels <- as.character(lev[[label_columns[[1]]]])
  missing_labels <- setdiff(requested_labels, labels)
  if (length(missing_labels)) {
    stop(
      "Missing land-use classes: ",
      paste(missing_labels, collapse = ", "),
      call. = FALSE
    )
  }

  stats::setNames(
    lev[[numeric_columns[[1]]]][
      match(requested_labels, labels)
    ],
    requested_labels
  )
}

indicator_for_codes <- function(x, codes) {
  terra::ifel(
    is.na(x),
    NA,
    terra::ifel(x %in% codes, 1, 0)
  )
}

zonal_sum_table <- function(x, zones, id_name) {
  out <- terra::zonal(
    x,
    zones,
    fun = "sum",
    na.rm = TRUE,
    wide = TRUE
  ) |>
    as.data.frame()
  names(out)[1] <- id_name
  out[[id_name]] <- as.numeric(out[[id_name]])
  out
}

global_number <- function(x, fun) {
  out <- if (fun %in% c("notNA", "isNA")) {
    terra::global(x, fun)
  } else {
    terra::global(x, fun, na.rm = TRUE)
  }
  as.numeric(out[1, 1])
}

unwrap_checkpoint <- function(path) {
  x <- readRDS(path)
  if (is.list(x) && "value" %in% names(x)) x$value else x
}

file_fingerprint <- function(paths) {
  paths <- normalizePath(paths, winslash = "/", mustWork = FALSE)
  info <- file.info(paths)
  tibble(path = paths, size = info$size, modified = as.numeric(info$mtime))
}

load_or_build_checkpoint <- function(
    path,
    signature,
    builder,
    required_sidecars = character()
) {
  if (
    reuse_matching_checkpoints &&
    file.exists(path) &&
    all(file.exists(required_sidecars))
  ) {
    checkpoint <- readRDS(path)
    if (
      is.list(checkpoint) &&
      identical(checkpoint$signature, signature) &&
      !is.null(checkpoint$value)
    ) {
      return(checkpoint$value)
    }
  }
  value <- builder()
  saveRDS(list(signature = signature, value = value), path)
  value
}

ampl_number <- function(x) {
  format(x, scientific = FALSE, trim = TRUE, digits = 15)
}

append_dat <- function(path, lines) {
  cat(
    paste(lines, collapse = "\n"),
    "\n",
    file = path,
    append = TRUE,
    sep = ""
  )
}

chunk_text <- function(x, chunk_size = 20L) {
  if (!length(x)) return(character())
  groups <- ceiling(seq_along(x) / chunk_size)
  vapply(
    split(x, groups),
    paste,
    collapse = " ",
    FUN.VALUE = character(1)
  )
}

write_ampl_set <- function(path, name, values) {
  append_dat(
    path,
    c(paste("set", name, ":="), chunk_text(values), ";", "")
  )
}

write_ampl_pair_set <- function(path, name, first, second) {
  tuples <- paste0("(", first, ",", second, ")")
  append_dat(
    path,
    c(
      paste("set", name, ":="),
      chunk_text(tuples, chunk_size = 8L),
      ";",
      ""
    )
  )
}

write_ampl_triple_set <- function(path, name, first, second, third) {
  tuples <- paste0("(", first, ",", second, ",", third, ")")
  append_dat(
    path,
    c(
      paste("set", name, ":="),
      chunk_text(tuples, chunk_size = 5L),
      ";",
      ""
    )
  )
}

write_ampl_scalar <- function(path, name, value) {
  append_dat(
    path,
    c(paste0("param ", name, " := ", ampl_number(value), ";"), "")
  )
}

write_ampl_vector <- function(path, name, key, value) {
  rows <- paste(key, ampl_number(value))
  append_dat(path, c(paste("param", name, ":="), rows, ";", ""))
}

write_ampl_matrix <- function(path, name, row_key, column_key, value) {
  rows <- paste(row_key, column_key, ampl_number(value))
  append_dat(path, c(paste("param", name, ":="), rows, ";", ""))
}

commit_generated_file <- function(temporary_path, final_path) {
  unchanged <- file.exists(final_path) &&
    unname(tools::md5sum(temporary_path)) ==
      unname(tools::md5sum(final_path))
  if (unchanged) {
    unlink(temporary_path)
    return(invisible(FALSE))
  }
  copied <- file.copy(temporary_path, final_path, overwrite = TRUE)
  unlink(temporary_path)
  if (!copied) stop("Could not write: ", final_path, call. = FALSE)
  invisible(TRUE)
}

write_lines_if_changed <- function(lines, path) {
  temporary_path <- paste0(path, ".building")
  writeLines(lines, temporary_path)
  commit_generated_file(temporary_path, path)
}

ampl_path <- function(path) chartr(intToUtf8(92L), "/", path)

append_start <- function(path, lines) {
  if (!length(lines)) return(invisible(NULL))
  cat(
    paste(lines, collapse = "\n"),
    "\n",
    file = path,
    append = TRUE,
    sep = ""
  )
}
```

# Read and validate the successful warm-up

```{r read-upstream}
dir.create(problem_dir, recursive = TRUE, showWarnings = FALSE)

warmup_selected_path <- file.path(
  warmup_dir,
  paste0(warmup_name, "_selected_cells.tsv")
)
warmup_summary_path <- file.path(
  warmup_dir,
  paste0(warmup_name, "_solution_summary.tsv")
)
candidate_table_path <- file.path(
  warmup_dir,
  "candidate_cell_dictionary.csv"
)
candidate_subarea_path <- file.path(
  warmup_dir,
  "candidate_area_by_subarea.csv"
)
candidate_edges_path <- file.path(
  warmup_dir,
  "candidate_rook_edges.csv"
)
subarea_budget_path <- file.path(
  warmup_dir,
  "subarea_30pct_budget.csv"
)
existing_contact_path <- file.path(
  warmup_dir,
  "existing_nature_contact_by_action.csv"
)
biodiversity_checkpoint_path <- file.path(
  warmup_dir,
  "coarse_biodiversity_benefits.rds"
)

biodiversity_support_path <- file.path(
  metrics_dir,
  paste0(
    biodiversity_support_metric,
    "_",
    biodiversity_support_scenario,
    ".tif"
  )
)

required_paths <- c(
  fine_landuse_path,
  coarse_landuse_path,
  shapefile_components(subareas_path),
  biodiversity_support_path,
  model_path,
  warmup_selected_path,
  warmup_summary_path,
  candidate_table_path,
  candidate_subarea_path,
  candidate_edges_path,
  subarea_budget_path,
  existing_contact_path,
  biodiversity_checkpoint_path
)
stop_for_missing_files(required_paths, "hard-network input file(s)")

warmup_summary <- readr::read_tsv(
  warmup_summary_path,
  show_col_types = FALSE,
  col_types = cols(.default = col_character())
)
if (!any(str_detect(tolower(warmup_summary$value), "solved|optimal"))) {
  stop(
    "The upstream summary does not report a solved/optimal warm-up.",
    call. = FALSE
  )
}

warmup_selected <- readr::read_tsv(
  warmup_selected_path,
  show_col_types = FALSE
) |>
  transmute(
    coarse_cell_id = as.integer(coarse_cell_id),
    landuse = as.character(landuse)
  )

candidate_table <- readr::read_csv(
  candidate_table_path,
  show_col_types = FALSE
) |>
  mutate(
    coarse_cell_id = as.integer(coarse_cell_id),
    ProductionForestShare = if_else(
      EligibleAreaHa > 0,
      ProductionForestHa / EligibleAreaHa,
      0
    ),
    IsProductionForestCell = as.integer(
      ProductionForestShare >= production_forest_share_threshold
    )
  ) |>
  filter(IsCandidate) |>
  arrange(coarse_cell_id)

candidate_cells <- candidate_table$coarse_cell_id

restorable_area_by_subarea_cell <- readr::read_csv(
  candidate_subarea_path,
  show_col_types = FALSE
) |>
  transmute(
    subarea_id = as.integer(subarea_id),
    coarse_cell_id = as.integer(coarse_cell_id),
    RestorableAreaHaBySubarea = as.numeric(
      RestorableAreaHaBySubarea
    )
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

candidate_edges <- readr::read_csv(
  candidate_edges_path,
  show_col_types = FALSE
) |>
  transmute(i = as.integer(i), j = as.integer(j)) |>
  filter(i %in% candidate_cells, j %in% candidate_cells) |>
  distinct() |>
  arrange(i, j)

subarea_budget <- readr::read_csv(
  subarea_budget_path,
  show_col_types = FALSE
) |>
  mutate(subarea_id = as.integer(subarea_id)) |>
  arrange(subarea_id)

existing_nature_contact <- readr::read_csv(
  existing_contact_path,
  show_col_types = FALSE
) |>
  transmute(
    landuse = as.character(landuse),
    coarse_cell_id = as.integer(coarse_cell_id),
    ExistingNatureContact = as.numeric(ExistingNatureContact)
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

biodiversity_checkpoint <- unwrap_checkpoint(
  biodiversity_checkpoint_path
)
if (
  !is.list(biodiversity_checkpoint) ||
  !"benefits" %in% names(biodiversity_checkpoint)
) {
  stop("Unexpected biodiversity checkpoint structure.", call. = FALSE)
}
biodiversity_table <- biodiversity_checkpoint$benefits |>
  transmute(
    coarse_cell_id = as.integer(coarse_cell_id),
    landuse = as.character(landuse),
    BioDiversity = as.numeric(BioDiversity)
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

stopifnot(
  nrow(subarea_budget) == expected_number_of_subareas,
  nrow(warmup_selected) > 0L,
  !anyDuplicated(warmup_selected$coarse_cell_id),
  all(warmup_selected$coarse_cell_id %in% candidate_cells),
  all(warmup_selected$landuse %in% nature_landuses),
  length(candidate_cells) > 0L,
  nrow(candidate_edges) > 0L,
  nrow(biodiversity_table) ==
    length(candidate_cells) * length(nature_landuses),
  nrow(existing_nature_contact) ==
    length(candidate_cells) * length(nature_landuses),
  !anyNA(biodiversity_table),
  !anyNA(existing_nature_contact)
)

input_audit <- file_fingerprint(required_paths)
readr::write_csv(
  input_audit,
  file.path(problem_dir, "input_file_audit.csv")
)
knitr::kable(input_audit)
```

# Reconstruct the aligned 10 m and 200 m domain

```{r aligned-domain}
fine_landuse_full <- terra::rast(fine_landuse_path)
coarse_landuse_full <- terra::rast(coarse_landuse_path)

stopifnot(
  terra::nlyr(fine_landuse_full) == 1L,
  terra::nlyr(coarse_landuse_full) == 1L,
  all(abs(terra::res(fine_landuse_full) - 10) < 1e-8),
  all(abs(terra::res(coarse_landuse_full) - 200) < 1e-8),
  same_crs(fine_landuse_full, coarse_landuse_full)
)

fine_cell_area_ha <- prod(terra::res(fine_landuse_full)) / 10000

subareas <- terra::vect(subareas_path) |>
  terra::makeValid()
if (nrow(subareas) != expected_number_of_subareas) {
  stop(
    "Expected four subareas; found ", nrow(subareas), ".",
    call. = FALSE
  )
}
if (!same_crs(subareas, fine_landuse_full)) {
  subareas <- terra::project(subareas, terra::crs(fine_landuse_full))
}

if (is.null(subarea_label_field)) {
  subarea_labels <- paste0("subarea_", seq_len(nrow(subareas)))
} else {
  if (!subarea_label_field %in% names(subareas)) {
    stop("Invalid subarea_label_field.", call. = FALSE)
  }
  subarea_labels <- make.unique(as.character(
    as.data.frame(subareas)[[subarea_label_field]]
  ))
}

subareas$subarea_id <- seq_len(nrow(subareas))
subareas$subarea_label <- subarea_labels
polygon_areas_ha <- as.numeric(
  terra::expanse(subareas, unit = "ha", transform = FALSE)
)
planning_domain <- terra::aggregate(subareas)

if (
  any(
    abs(sort(polygon_areas_ha) - sort(expected_subarea_areas_ha)) >
      polygon_area_audit_tolerance_ha
  ) ||
  abs(sum(polygon_areas_ha) - expected_total_area_ha) >
    polygon_area_audit_tolerance_ha
) {
  stop(
    "The polygon areas do not match the intended smaller Gudenå polygon.",
    call. = FALSE
  )
}

fine_landuse <- terra::crop(
  fine_landuse_full,
  planning_domain,
  snap = "out"
) |>
  terra::mask(planning_domain, touches = FALSE)

coarse_landuse <- terra::crop(
  coarse_landuse_full,
  planning_domain,
  snap = "out"
) |>
  terra::mask(planning_domain, touches = TRUE)

coarse_cell_id <- terra::init(coarse_landuse, "cell")
coarse_cell_id <- terra::ifel(
  is.na(coarse_landuse),
  NA,
  coarse_cell_id
)
names(coarse_cell_id) <- "coarse_cell_id"

coarse_id_on_fine <- align_to_template(
  coarse_cell_id,
  fine_landuse,
  method = "near"
)
names(coarse_id_on_fine) <- "coarse_cell_id"

subarea_id_on_fine <- terra::rasterize(
  subareas[, "subarea_id"],
  fine_landuse,
  field = "subarea_id",
  background = NA,
  touches = FALSE
)
subarea_id_on_fine <- terra::mask(
  subarea_id_on_fine,
  fine_landuse
)
names(subarea_id_on_fine) <- "subarea_id"

all_required_classes <- c(
  nature_landuses,
  "Agriculture",
  "ProductionForest",
  "Urban",
  "Other"
)
landuse_codes <- category_codes(fine_landuse, all_required_classes)
nature_codes <- unname(landuse_codes[nature_landuses])
eligible_codes <- unname(
  landuse_codes[c("Agriculture", "ProductionForest")]
)

fine_nature <- indicator_for_codes(fine_landuse, nature_codes)
names(fine_nature) <- "CurrentNatureCells"

biodiversity_support <- terra::rast(biodiversity_support_path)
biodiversity_support <- align_to_template(
  biodiversity_support,
  fine_landuse,
  method = "bilinear"
)
fine_biodiversity_support <- terra::ifel(
  is.na(fine_landuse),
  NA,
  terra::ifel(!is.na(biodiversity_support), 1, 0)
)

fine_eligible <-
  indicator_for_codes(fine_landuse, eligible_codes) *
  fine_biodiversity_support
fine_eligible_mask <- terra::ifel(fine_eligible == 1, 1, NA)
names(fine_eligible_mask) <- "EligibleCells"

candidate_mask_on_fine <- terra::ifel(
  coarse_id_on_fine %in% candidate_cells,
  1,
  NA
)
fine_candidate_eligible_mask <- terra::mask(
  fine_eligible_mask,
  candidate_mask_on_fine
)

if (
  global_number(
    terra::ifel(
      !is.na(fine_landuse) & is.na(coarse_id_on_fine),
      1,
      0
    ),
    "sum"
  ) > 0
) {
  stop("Some valid 10 m cells have no 200 m ID.", call. = FALSE)
}

coarse_ids_reconstructed <- sort(unique(na.omit(as.integer(
  terra::values(coarse_cell_id, mat = FALSE)
))))
if (!all(candidate_cells %in% coarse_ids_reconstructed)) {
  stop(
    "Reconstructed coarse IDs do not match notebook 07 outputs.",
    call. = FALSE
  )
}
```

# Current-nature patches and candidate contacts

Current nature is binarized across all eight nature types and split into exact
10 m rook-connected patches. Each retained patch becomes one graph supernode.

```{r patch-network}
patch_raster_path <- file.path(
  problem_dir,
  "current_nature_patches_10m.tif"
)

patch_signature <- digest::digest(list(
  files = file_fingerprint(c(
    fine_landuse_path,
    coarse_landuse_path,
    shapefile_components(subareas_path),
    biodiversity_support_path,
    candidate_table_path,
    candidate_edges_path
  )),
  candidate_cells = candidate_cells,
  directions = 4L,
  fine_resolution = terra::res(fine_landuse)
))

patch_network <- load_or_build_checkpoint(
  file.path(problem_dir, "existing_patch_network.rds"),
  patch_signature,
  function() {
    current_patch_raster <- terra::patches(
      fine_nature,
      directions = 4,
      zeroAsNA = TRUE
    )
    names(current_patch_raster) <- "patch_id"
    terra::writeRaster(
      current_patch_raster,
      patch_raster_path,
      overwrite = TRUE,
      datatype = "INT4U",
      gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
    )

    patch_table_all <- zonal_sum_table(
      fine_nature,
      current_patch_raster,
      "patch_id"
    ) |>
      transmute(
        patch_id = as.integer(round(patch_id)),
        ExistingPatchAreaHa =
          CurrentNatureCells * fine_cell_area_ha
      )

    eligible_indices <- which(!is.na(terra::values(
      fine_candidate_eligible_mask,
      mat = FALSE
    )))
    if (!length(eligible_indices)) {
      stop("No fine candidate cells are available.", call. = FALSE)
    }

    adjacent_pairs <- terra::adjacent(
      fine_landuse,
      cells = eligible_indices,
      directions = 4,
      pairs = TRUE,
      include = FALSE
    ) |>
      as.matrix()

    patch_values <- terra::values(
      current_patch_raster,
      mat = FALSE
    )
    coarse_values <- terra::values(coarse_id_on_fine, mat = FALSE)

    cell_patch_links <- tibble(
      coarse_cell_id = as.integer(
        round(coarse_values[adjacent_pairs[, 1]])
      ),
      patch_id = as.integer(
        round(patch_values[adjacent_pairs[, 2]])
      )
    ) |>
      filter(
        !is.na(coarse_cell_id),
        !is.na(patch_id),
        coarse_cell_id %in% candidate_cells
      ) |>
      distinct() |>
      arrange(coarse_cell_id, patch_id)

    retained_patch_ids <- sort(unique(cell_patch_links$patch_id))
    patch_table <- patch_table_all |>
      filter(patch_id %in% retained_patch_ids) |>
      arrange(patch_id)

    pair_multiplier <- expected_number_of_subareas + 1L
    patch_subarea_zone <- terra::ifel(
      is.na(current_patch_raster) | is.na(subarea_id_on_fine),
      NA,
      current_patch_raster * pair_multiplier +
        subarea_id_on_fine
    )

    patch_area_by_subarea <- zonal_sum_table(
      fine_nature,
      patch_subarea_zone,
      "pair_id"
    ) |>
      mutate(
        patch_id = as.integer(floor(pair_id / pair_multiplier)),
        subarea_id = as.integer(
          round(pair_id - patch_id * pair_multiplier)
        ),
        ExistingPatchAreaHaBySubarea =
          CurrentNatureCells * fine_cell_area_ha
      ) |>
      filter(
        patch_id %in% retained_patch_ids,
        subarea_id %in% seq_len(expected_number_of_subareas),
        ExistingPatchAreaHaBySubarea > 0
      ) |>
      select(
        patch_id,
        subarea_id,
        ExistingPatchAreaHaBySubarea
      ) |>
      arrange(subarea_id, patch_id)

    list(
      patches = patch_table,
      patch_area_by_subarea = patch_area_by_subarea,
      links = cell_patch_links
    )
  },
  required_sidecars = patch_raster_path
)

current_patch_raster <- terra::rast(patch_raster_path)
patch_table <- patch_network$patches
patch_area_by_subarea <- patch_network$patch_area_by_subarea
cell_patch_links <- patch_network$links
existing_patches <- patch_table$patch_id

stopifnot(
  nrow(patch_table) > 0L,
  nrow(cell_patch_links) > 0L,
  all(cell_patch_links$coarse_cell_id %in% candidate_cells),
  all(cell_patch_links$patch_id %in% existing_patches),
  all(patch_table$ExistingPatchAreaHa > 0),
  !anyDuplicated(patch_table$patch_id),
  !anyDuplicated(cell_patch_links)
)

patch_area_crosscheck <- patch_area_by_subarea |>
  group_by(patch_id) |>
  summarise(
    SplitAreaHa = sum(ExistingPatchAreaHaBySubarea),
    .groups = "drop"
  ) |>
  right_join(patch_table, by = "patch_id")

if (
  anyNA(patch_area_crosscheck$SplitAreaHa) ||
  any(abs(
    patch_area_crosscheck$SplitAreaHa -
      patch_area_crosscheck$ExistingPatchAreaHa
  ) > fine_cell_area_ha + 1e-8)
) {
  stop("Patch areas do not reconcile by subarea.", call. = FALSE)
}

readr::write_csv(
  patch_table,
  file.path(problem_dir, "existing_patch_dictionary.csv")
)
readr::write_csv(
  patch_area_by_subarea,
  file.path(problem_dir, "existing_patch_area_by_subarea.csv")
)
readr::write_csv(
  cell_patch_links,
  file.path(problem_dir, "candidate_to_existing_patch_links.csv")
)

knitr::kable(tibble(
  retained_existing_patches = nrow(patch_table),
  current_nature_ha_in_retained_patches =
    sum(patch_table$ExistingPatchAreaHa),
  candidate_patch_contacts = nrow(cell_patch_links),
  candidate_cells_touching_current_nature =
    n_distinct(cell_patch_links$coarse_cell_id)
), digits = 2)
```

# Potential graph and tight flow bounds

```{r potential-graph}
potential_vertices <- bind_rows(
  candidate_table |>
    transmute(
      vertex = paste0("c_", coarse_cell_id),
      node_type = "cell",
      node_id = coarse_cell_id,
      FlowAreaHa = EligibleAreaHa
    ),
  patch_table |>
    transmute(
      vertex = paste0("p_", patch_id),
      node_type = "patch",
      node_id = patch_id,
      FlowAreaHa = ExistingPatchAreaHa
    )
)

potential_edges <- bind_rows(
  candidate_edges |>
    transmute(from = paste0("c_", i), to = paste0("c_", j)),
  cell_patch_links |>
    transmute(
      from = paste0("c_", coarse_cell_id),
      to = paste0("p_", patch_id)
    )
) |>
  distinct()

potential_graph <- igraph::graph_from_data_frame(
  potential_edges,
  directed = FALSE,
  vertices = potential_vertices
)
potential_membership_raw <- igraph::components(
  potential_graph
)$membership
potential_membership <- tibble(
  vertex = names(potential_membership_raw),
  potential_component = as.integer(potential_membership_raw)
)

potential_vertices <- potential_vertices |>
  left_join(potential_membership, by = "vertex")

potential_component_summary <- potential_vertices |>
  group_by(potential_component) |>
  summarise(
    ComponentMaxAreaHa = sum(FlowAreaHa),
    candidate_cells = sum(node_type == "cell"),
    existing_patches = sum(node_type == "patch"),
    .groups = "drop"
  )

cell_component_bounds <- potential_vertices |>
  filter(node_type == "cell") |>
  select(coarse_cell_id = node_id, potential_component) |>
  left_join(potential_component_summary, by = "potential_component") |>
  select(coarse_cell_id, ComponentMaxAreaHa)

patch_component_bounds <- potential_vertices |>
  filter(node_type == "patch") |>
  select(patch_id = node_id, potential_component) |>
  left_join(potential_component_summary, by = "potential_component") |>
  select(patch_id, ComponentMaxAreaHa)

# Area coefficients for the Global and four Subarea certificates.
network_cell_area <- bind_rows(
  candidate_table |>
    transmute(
      network_name = "Global",
      coarse_cell_id,
      NetworkCellAreaHa = EligibleAreaHa
    ),
  restorable_area_by_subarea_cell |>
    transmute(
      network_name = paste0("Subarea", subarea_id),
      coarse_cell_id,
      NetworkCellAreaHa = RestorableAreaHaBySubarea
    )
) |>
  filter(NetworkCellAreaHa > 0)

network_patch_area <- bind_rows(
  patch_table |>
    transmute(
      network_name = "Global",
      patch_id,
      NetworkPatchAreaHa = ExistingPatchAreaHa
    ),
  patch_area_by_subarea |>
    transmute(
      network_name = paste0("Subarea", subarea_id),
      patch_id,
      NetworkPatchAreaHa = ExistingPatchAreaHaBySubarea
    )
) |>
  filter(NetworkPatchAreaHa > 0)

potential_component_network_area <- bind_rows(
  network_cell_area |>
    left_join(
      potential_vertices |>
        filter(node_type == "cell") |>
        transmute(
          coarse_cell_id = node_id,
          potential_component
        ),
      by = "coarse_cell_id"
    ) |>
    transmute(
      network_name,
      potential_component,
      CertificateAreaHa = NetworkCellAreaHa
    ),
  network_patch_area |>
    left_join(
      potential_vertices |>
        filter(node_type == "patch") |>
        transmute(patch_id = node_id, potential_component),
      by = "patch_id"
    ) |>
    transmute(
      network_name,
      potential_component,
      CertificateAreaHa = NetworkPatchAreaHa
    )
) |>
  group_by(network_name, potential_component) |>
  summarise(
    PotentialCertificateAreaHa = sum(CertificateAreaHa),
    .groups = "drop"
  ) |>
  left_join(
    potential_component_summary,
    by = "potential_component"
  ) |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    CanAnchor = candidate_cells > 0 & existing_patches > 0,
    PassesPotentialArea =
      PotentialCertificateAreaHa >= RequiredAreaHa - 1e-8,
    FeasiblePotentialComponent = CanAnchor & PassesPotentialArea
  )

potential_network_audit <- tibble(
  network_name = required_networks,
  RequiredAreaHa = unname(network_minimums[required_networks])
) |>
  left_join(
    potential_component_network_area |>
      group_by(network_name) |>
      summarise(
        RequiredAreaHa = first(RequiredAreaHa),
        feasible_potential_components =
          sum(FeasiblePotentialComponent),
        largest_potential_certificate_ha = max(
          PotentialCertificateAreaHa[CanAnchor],
          na.rm = TRUE
        ),
        .groups = "drop"
      ),
    by = c("network_name", "RequiredAreaHa")
  )

if (
  anyNA(potential_network_audit$feasible_potential_components) ||
  any(potential_network_audit$feasible_potential_components < 1L)
) {
  stop(
    "At least one requested network is impossible in the potential graph. ",
    "Inspect potential_network_feasibility_audit.csv.",
    call. = FALSE
  )
}

patch_connection_normalizer <- max(
  1,
  max(potential_component_summary$existing_patches) - 1
)

readr::write_csv(
  potential_component_summary,
  file.path(problem_dir, "potential_graph_components.csv")
)
readr::write_csv(
  potential_component_network_area,
  file.path(problem_dir, "potential_component_network_area.csv")
)
readr::write_csv(
  potential_network_audit,
  file.path(problem_dir, "potential_network_feasibility_audit.csv")
)

knitr::kable(potential_network_audit, digits = 2)
```

# Feasible actions and warm-up label repair

The selected **footprint** from notebook 07 is preserved in the MIP start.
Only a destination label is repaired when a production-forest-dominated cell
was assigned an open habitat in the simplified warm-up.

```{r model-actions}
allowed_actions <- tidyr::crossing(
  landuse = nature_landuses,
  candidate_table |>
    select(coarse_cell_id, IsProductionForestCell)
) |>
  filter(
    IsProductionForestCell == 0 |
      landuse %in% forest_landuses
  ) |>
  select(landuse, coarse_cell_id) |>
  arrange(landuse, coarse_cell_id)

objective_table <- allowed_actions |>
  left_join(
    biodiversity_table,
    by = c("landuse", "coarse_cell_id")
  ) |>
  left_join(
    existing_nature_contact,
    by = c("landuse", "coarse_cell_id")
  )

stopifnot(
  n_distinct(allowed_actions$coarse_cell_id) ==
    length(candidate_cells),
  !anyDuplicated(allowed_actions),
  !anyNA(objective_table),
  all(is.finite(objective_table$BioDiversity)),
  all(is.finite(objective_table$ExistingNatureContact))
)

biodiversity_normalizer <- objective_table |>
  group_by(coarse_cell_id) |>
  summarise(best = max(BioDiversity), .groups = "drop") |>
  summarise(total = sum(best)) |>
  pull(total) |>
  max(1)

contiguity_normalizer <- max(1, nrow(candidate_edges))

existing_nature_contact_normalizer <- objective_table |>
  group_by(coarse_cell_id) |>
  summarise(
    best = max(ExistingNatureContact),
    .groups = "drop"
  ) |>
  summarise(total = sum(best)) |>
  pull(total) |>
  max(1)

allowed_i <- allowed_actions |>
  transmute(landuse, i = coarse_cell_id)
allowed_j <- allowed_actions |>
  transmute(landuse, j = coarse_cell_id)

same_habitat_edges <- tidyr::crossing(
  landuse = nature_landuses,
  candidate_edges
) |>
  semi_join(allowed_i, by = c("landuse", "i")) |>
  semi_join(allowed_j, by = c("landuse", "j")) |>
  arrange(landuse, i, j)

best_forest_action <- objective_table |>
  filter(landuse %in% forest_landuses) |>
  mutate(
    local_start_score =
      BioDiversity / biodiversity_normalizer +
      spatial_contiguity_bonus *
        existing_to_new_adjacency_weight *
        ExistingNatureContact /
          existing_nature_contact_normalizer
  ) |>
  group_by(coarse_cell_id) |>
  arrange(
    desc(local_start_score),
    desc(BioDiversity),
    landuse,
    .by_group = TRUE
  ) |>
  slice_head(n = 1L) |>
  ungroup() |>
  select(
    coarse_cell_id,
    repaired_forest_landuse = landuse
  )

warm_start_assignment <- warmup_selected |>
  rename(original_landuse = landuse) |>
  left_join(
    candidate_table |>
      select(coarse_cell_id, IsProductionForestCell),
    by = "coarse_cell_id"
  ) |>
  left_join(best_forest_action, by = "coarse_cell_id") |>
  mutate(
    forest_rule_repair =
      IsProductionForestCell == 1L &
      !original_landuse %in% forest_landuses,
    landuse = if_else(
      forest_rule_repair,
      repaired_forest_landuse,
      original_landuse
    )
  ) |>
  select(
    coarse_cell_id,
    original_landuse,
    landuse,
    IsProductionForestCell,
    forest_rule_repair
  ) |>
  arrange(coarse_cell_id)

invalid_start_actions <- warm_start_assignment |>
  anti_join(
    allowed_actions,
    by = c("landuse", "coarse_cell_id")
  )
if (nrow(invalid_start_actions)) {
  stop("Warm-start action repair failed.", call. = FALSE)
}

warm_start_area_audit <- restorable_area_by_subarea_cell |>
  filter(
    coarse_cell_id %in% warm_start_assignment$coarse_cell_id
  ) |>
  group_by(subarea_id) |>
  summarise(
    WarmStartRestorationHa = sum(RestorableAreaHaBySubarea),
    .groups = "drop"
  ) |>
  right_join(subarea_budget, by = "subarea_id") |>
  mutate(
    WarmStartRestorationHa = coalesce(WarmStartRestorationHa, 0),
    PassesLower =
      WarmStartRestorationHa >= RestorationRequirementHa - 1e-6,
    PassesUpper =
      WarmStartRestorationHa <=
        RestorationRequirementHa + RestorationToleranceHa + 1e-6
  ) |>
  arrange(subarea_id)

if (!all(
  warm_start_area_audit$PassesLower &
    warm_start_area_audit$PassesUpper
)) {
  stop(
    "The upstream footprint does not satisfy the unchanged 30% bounds.",
    call. = FALSE
  )
}

readr::write_csv(
  warm_start_assignment,
  file.path(problem_dir, "warm_start_assignment_and_forest_repairs.csv")
)
readr::write_csv(
  warm_start_area_audit,
  file.path(problem_dir, "warm_start_30pct_audit.csv")
)

knitr::kable(tibble(
  warm_start_selected_cells = nrow(warm_start_assignment),
  production_forest_dominated_selected_cells = sum(
    warm_start_assignment$IsProductionForestCell == 1L
  ),
  destination_labels_repaired = sum(
    warm_start_assignment$forest_rule_repair
  )
))
```

# Find feasible certificate components in the warm-up solution

```{r warm-start-components}
warm_start_ids <- warm_start_assignment$coarse_cell_id
warm_start_patch_ids <- cell_patch_links |>
  filter(coarse_cell_id %in% warm_start_ids) |>
  distinct(patch_id) |>
  pull(patch_id) |>
  sort()

warm_start_vertices <- bind_rows(
  candidate_table |>
    filter(coarse_cell_id %in% warm_start_ids) |>
    transmute(
      vertex = paste0("c_", coarse_cell_id),
      node_type = "cell",
      node_id = coarse_cell_id,
      FlowAreaHa = EligibleAreaHa
    ),
  patch_table |>
    filter(patch_id %in% warm_start_patch_ids) |>
    transmute(
      vertex = paste0("p_", patch_id),
      node_type = "patch",
      node_id = patch_id,
      FlowAreaHa = ExistingPatchAreaHa
    )
)

warm_start_edges <- bind_rows(
  candidate_edges |>
    filter(i %in% warm_start_ids, j %in% warm_start_ids) |>
    transmute(from = paste0("c_", i), to = paste0("c_", j)),
  cell_patch_links |>
    filter(
      coarse_cell_id %in% warm_start_ids,
      patch_id %in% warm_start_patch_ids
    ) |>
    transmute(
      from = paste0("c_", coarse_cell_id),
      to = paste0("p_", patch_id)
    )
) |>
  distinct()

warm_start_graph <- igraph::graph_from_data_frame(
  warm_start_edges,
  directed = FALSE,
  vertices = warm_start_vertices
)
warm_membership_raw <- igraph::components(warm_start_graph)$membership
warm_membership <- tibble(
  vertex = names(warm_membership_raw),
  warm_component = as.integer(warm_membership_raw)
)
warm_start_vertices <- warm_start_vertices |>
  left_join(warm_membership, by = "vertex")

warm_component_summary <- warm_start_vertices |>
  group_by(warm_component) |>
  summarise(
    selected_cells = sum(node_type == "cell"),
    existing_patches = sum(node_type == "patch"),
    FlowAreaHa = sum(FlowAreaHa),
    .groups = "drop"
  )

warm_cell_membership <- warm_start_vertices |>
  filter(node_type == "cell") |>
  transmute(
    coarse_cell_id = node_id,
    warm_component
  )
warm_patch_membership <- warm_start_vertices |>
  filter(node_type == "patch") |>
  transmute(patch_id = node_id, warm_component)

warm_component_network_area <- bind_rows(
  network_cell_area |>
    filter(coarse_cell_id %in% warm_start_ids) |>
    left_join(warm_cell_membership, by = "coarse_cell_id") |>
    transmute(
      network_name,
      warm_component,
      CertificateAreaHa = NetworkCellAreaHa
    ),
  network_patch_area |>
    filter(patch_id %in% warm_start_patch_ids) |>
    left_join(warm_patch_membership, by = "patch_id") |>
    transmute(
      network_name,
      warm_component,
      CertificateAreaHa = NetworkPatchAreaHa
    )
) |>
  group_by(network_name, warm_component) |>
  summarise(
    CertifiedAreaHa = sum(CertificateAreaHa),
    .groups = "drop"
  ) |>
  left_join(warm_component_summary, by = "warm_component") |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    HasExistingRoot = existing_patches > 0,
    HasNewRestoration = selected_cells > 0,
    PassesArea = CertifiedAreaHa >= RequiredAreaHa - 1e-8,
    FeasibleWarmCertificate =
      HasExistingRoot & HasNewRestoration & PassesArea
  )

chosen_certificate_components <- map_dfr(
  required_networks,
  function(k) {
    choices <- warm_component_network_area |>
      filter(
        network_name == k,
        FeasibleWarmCertificate
      ) |>
      arrange(desc(CertifiedAreaHa), warm_component)

    if (!nrow(choices)) {
      stop(
        "The warm-up result cannot seed required certificate ",
        k,
        ". Inspect warm_start_component_network_audit.csv. ",
        "The hard model itself may still be feasible, but this notebook ",
        "deliberately stops rather than claiming a feasible warm start.",
        call. = FALSE
      )
    }

    choices |>
      slice_head(n = 1L)
  }
)

readr::write_csv(
  warm_component_network_area,
  file.path(problem_dir, "warm_start_component_network_audit.csv")
)
readr::write_csv(
  chosen_certificate_components,
  file.path(problem_dir, "warm_start_chosen_certificates.csv")
)

knitr::kable(
  chosen_certificate_components |>
    select(
      network_name,
      warm_component,
      RequiredAreaHa,
      CertifiedAreaHa,
      selected_cells,
      existing_patches
    ),
  digits = 2
)
```

# Construct complete flow starts for all five certificates

Each chosen component is converted to a deterministic spanning tree directed
toward its largest existing-nature patch. Subtree areas give a complete
feasible start for the continuous flow variables as well as all binary
variables.

```{r build-complete-mip-start}
build_certificate_start <- function(network_name, component_id) {
  component_nodes <- warm_start_vertices |>
    filter(warm_component == component_id)

  root_row <- component_nodes |>
    filter(node_type == "patch") |>
    arrange(desc(FlowAreaHa), node_id) |>
    slice_head(n = 1L)
  if (nrow(root_row) != 1L) {
    stop("Certificate has no existing-nature root.", call. = FALSE)
  }

  root_vertex <- root_row$vertex[[1]]
  component_graph <- igraph::induced_subgraph(
    warm_start_graph,
    vids = component_nodes$vertex
  )

  distances_to_root <- as.numeric(igraph::distances(
    component_graph,
    v = root_vertex,
    to = igraph::V(component_graph),
    mode = "all"
  )[1, ])
  names(distances_to_root) <- igraph::V(component_graph)$name
  if (any(!is.finite(distances_to_root))) {
    stop("Internal error: certificate component is disconnected.")
  }

  all_vertices <- component_nodes$vertex
  parent <- stats::setNames(
    rep(NA_character_, length(all_vertices)),
    all_vertices
  )

  for (vertex in setdiff(all_vertices, root_vertex)) {
    neighbors <- igraph::as_ids(igraph::neighbors(
      component_graph,
      vertex,
      mode = "all"
    ))
    possible_parents <- neighbors[
      distances_to_root[neighbors] ==
        distances_to_root[[vertex]] - 1
    ]
    if (!length(possible_parents)) {
      stop("Could not assign a flow-tree parent to ", vertex)
    }
    parent[[vertex]] <- sort(possible_parents)[[1]]
  }

  subtree_area <- stats::setNames(
    component_nodes$FlowAreaHa[match(
      all_vertices,
      component_nodes$vertex
    )],
    all_vertices
  )
  flow_records <- list()
  flow_index <- 0L
  descending_vertices <- all_vertices[
    order(distances_to_root[all_vertices], decreasing = TRUE)
  ]

  for (vertex in descending_vertices) {
    if (vertex == root_vertex) next
    parent_vertex <- parent[[vertex]]
    flow_index <- flow_index + 1L
    flow_records[[flow_index]] <- tibble(
      network_name = network_name,
      child = vertex,
      parent = parent_vertex,
      flow_ha = subtree_area[[vertex]]
    )
    subtree_area[[parent_vertex]] <-
      subtree_area[[parent_vertex]] + subtree_area[[vertex]]
  }

  flow_table <- bind_rows(flow_records) |>
    mutate(
      child_type = str_sub(child, 1, 1),
      parent_type = str_sub(parent, 1, 1),
      child_id = as.integer(str_remove(child, "^[cp]_")),
      parent_id = as.integer(str_remove(parent, "^[cp]_"))
    )

  if (any(
    flow_table$child_type == "p" &
      flow_table$parent_type == "p"
  )) {
    stop("Unexpected patch-to-patch edge in a flow tree.")
  }

  list(
    network_name = network_name,
    component_id = component_id,
    cells = component_nodes |>
      filter(node_type == "cell") |>
      pull(node_id) |>
      as.integer() |>
      sort(),
    patches = component_nodes |>
      filter(node_type == "patch") |>
      pull(node_id) |>
      as.integer() |>
      sort(),
    root_patch = as.integer(root_row$node_id[[1]]),
    collected_area_ha = subtree_area[[root_vertex]],
    flows = flow_table
  )
}

certificate_starts <- map(
  seq_len(nrow(chosen_certificate_components)),
  function(index) {
    build_certificate_start(
      chosen_certificate_components$network_name[[index]],
      chosen_certificate_components$warm_component[[index]]
    )
  }
)
names(certificate_starts) <- map_chr(
  certificate_starts,
  "network_name"
)

start_certificate_cell_members <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    coarse_cell_id = x$cells
  )
)
start_certificate_patch_members <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    patch_id = x$patches
  )
)
start_certificate_roots <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    patch_id = x$root_patch,
    collected_area_ha = x$collected_area_ha
  )
)
start_certificate_flows <- map_dfr(certificate_starts, "flows")

start_flow_structure_audit <- map_dfr(
  certificate_starts,
  function(x) {
    expected_collected <-
      sum(candidate_table$EligibleAreaHa[
        match(x$cells, candidate_table$coarse_cell_id)
      ]) +
      sum(patch_table$ExistingPatchAreaHa[
        match(x$patches, patch_table$patch_id)
      ])

    tibble(
      network_name = x$network_name,
      member_nodes = length(x$cells) + length(x$patches),
      spanning_tree_arcs = nrow(x$flows),
      expected_collected_area_ha = expected_collected,
      start_collected_area_ha = x$collected_area_ha,
      tree_has_n_minus_one_arcs =
        nrow(x$flows) == length(x$cells) + length(x$patches) - 1L,
      collected_area_reconciles =
        abs(x$collected_area_ha - expected_collected) <= 1e-6,
      all_tree_flows_positive = all(x$flows$flow_ha > 0)
    )
  }
)

if (!all(
  start_flow_structure_audit$tree_has_n_minus_one_arcs &
    start_flow_structure_audit$collected_area_reconciles &
    start_flow_structure_audit$all_tree_flows_positive
)) {
  stop("Internal error in certificate flow starts.", call. = FALSE)
}

# Independent membership and area checks before any AMPL files are written.
start_certificate_area_check <- bind_rows(
  start_certificate_cell_members |>
    left_join(
      network_cell_area,
      by = c("network_name", "coarse_cell_id")
    ) |>
    transmute(
      network_name,
      area_ha = coalesce(NetworkCellAreaHa, 0)
    ),
  start_certificate_patch_members |>
    left_join(
      network_patch_area,
      by = c("network_name", "patch_id")
    ) |>
    transmute(
      network_name,
      area_ha = coalesce(NetworkPatchAreaHa, 0)
    )
) |>
  group_by(network_name) |>
  summarise(CertifiedAreaHa = sum(area_ha), .groups = "drop") |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    Passes = CertifiedAreaHa >= RequiredAreaHa - 1e-6
  )

if (
  nrow(start_certificate_area_check) != length(required_networks) ||
  !all(start_certificate_area_check$Passes)
) {
  stop("Internal error in warm-start certificate areas.", call. = FALSE)
}

if (
  nrow(start_certificate_roots) != length(required_networks) ||
  anyDuplicated(start_certificate_roots$network_name)
) {
  stop("Each certificate must have exactly one root.", call. = FALSE)
}

readr::write_csv(
  start_certificate_area_check,
  file.path(problem_dir, "warm_start_certificate_area_check.csv")
)
readr::write_csv(
  start_certificate_roots,
  file.path(problem_dir, "warm_start_certificate_roots.csv")
)
readr::write_csv(
  start_flow_structure_audit,
  file.path(problem_dir, "warm_start_flow_structure_audit.csv")
)

print(knitr::kable(start_certificate_area_check, digits = 2))
knitr::kable(start_flow_structure_audit, digits = 6)
```

# Export AMPL data, complete MIP start, and run file

```{r export-ampl}
dat_path <- file.path(problem_dir, paste0(problem_name, ".dat"))
mip_start_path <- file.path(
  problem_dir,
  paste0(problem_name, "_mip_start.run")
)
run_path <- file.path(problem_dir, paste0(problem_name, ".run"))

solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_selected_cells.tsv")
)
selected_subarea_path <- file.path(
  problem_dir,
  paste0(problem_name, "_selected_cell_subareas.tsv")
)
subarea_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_subarea_summary.tsv")
)
network_cell_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_cells.tsv")
)
network_patch_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_patches.tsv")
)
network_root_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_roots.tsv")
)
network_solution_summary_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_summary.tsv")
)
summary_path <- file.path(
  problem_dir,
  paste0(problem_name, "_solution_summary.tsv")
)

dat_temporary_path <- paste0(dat_path, ".building")
writeLines(
  c(
    "# Generated by 08_gudenaa_new_area_200m_hard_networks.Rmd",
    paste("# Polygon area (ha):", ampl_number(sum(polygon_areas_ha))),
    paste(
      "# Global network minimum (ha):",
      ampl_number(global_network_minimum_ha)
    ),
    paste(
      "# Per-subarea network minimum (ha):",
      ampl_number(subarea_network_minimum_ha)
    ),
    "# Carbon, MustIncludeCells, and forest/wet quotas are absent.",
    ""
  ),
  dat_temporary_path
)

write_ampl_set(dat_temporary_path, "Cells", candidate_cells)
write_ampl_set(dat_temporary_path, "Landuses", nature_landuses)
write_ampl_set(dat_temporary_path, "ForestLanduses", forest_landuses)
write_ampl_set(
  dat_temporary_path,
  "Subareas",
  subarea_budget$subarea_id
)
write_ampl_set(
  dat_temporary_path,
  "RequiredNetworks",
  required_networks
)
write_ampl_set(dat_temporary_path, "GlobalNetworks", "Global")
write_ampl_pair_set(
  dat_temporary_path,
  "AllowedActions",
  allowed_actions$landuse,
  allowed_actions$coarse_cell_id
)
write_ampl_pair_set(
  dat_temporary_path,
  "E",
  candidate_edges$i,
  candidate_edges$j
)
write_ampl_triple_set(
  dat_temporary_path,
  "SameHabitatEdges",
  same_habitat_edges$landuse,
  same_habitat_edges$i,
  same_habitat_edges$j
)
write_ampl_set(
  dat_temporary_path,
  "ExistingPatches",
  existing_patches
)
write_ampl_pair_set(
  dat_temporary_path,
  "CellPatchEdges",
  cell_patch_links$coarse_cell_id,
  cell_patch_links$patch_id
)

write_ampl_matrix(
  dat_temporary_path,
  "BioDiversity",
  objective_table$landuse,
  objective_table$coarse_cell_id,
  objective_table$BioDiversity
)
write_ampl_matrix(
  dat_temporary_path,
  "ExistingNatureContact",
  objective_table$landuse,
  objective_table$coarse_cell_id,
  objective_table$ExistingNatureContact
)
write_ampl_vector(
  dat_temporary_path,
  "RestorableAreaHa",
  candidate_table$coarse_cell_id,
  candidate_table$EligibleAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "RestorableAreaHaBySubarea",
  restorable_area_by_subarea_cell$subarea_id,
  restorable_area_by_subarea_cell$coarse_cell_id,
  restorable_area_by_subarea_cell$RestorableAreaHaBySubarea
)
write_ampl_vector(
  dat_temporary_path,
  "SubareaAreaHa",
  subarea_budget$subarea_id,
  subarea_budget$PolygonAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "CurrentNatureAreaHa",
  subarea_budget$subarea_id,
  subarea_budget$CurrentNatureAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "FinalNatureTargetHa",
  subarea_budget$subarea_id,
  subarea_budget$FinalNatureTargetHa
)
write_ampl_vector(
  dat_temporary_path,
  "RestorationRequirementHa",
  subarea_budget$subarea_id,
  subarea_budget$RestorationRequirementHa
)
write_ampl_vector(
  dat_temporary_path,
  "RestorationToleranceHa",
  subarea_budget$subarea_id,
  subarea_budget$RestorationToleranceHa
)

write_ampl_vector(
  dat_temporary_path,
  "ExistingPatchAreaHa",
  patch_table$patch_id,
  patch_table$ExistingPatchAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "NetworkCellAreaHa",
  network_cell_area$network_name,
  network_cell_area$coarse_cell_id,
  network_cell_area$NetworkCellAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "NetworkPatchAreaHa",
  network_patch_area$network_name,
  network_patch_area$patch_id,
  network_patch_area$NetworkPatchAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "NetworkMinimumHa",
  names(network_minimums),
  unname(network_minimums)
)
write_ampl_vector(
  dat_temporary_path,
  "ComponentMaxAreaCellHa",
  cell_component_bounds$coarse_cell_id,
  cell_component_bounds$ComponentMaxAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "ComponentMaxAreaPatchHa",
  patch_component_bounds$patch_id,
  patch_component_bounds$ComponentMaxAreaHa
)

production_forest_cells <- candidate_table$coarse_cell_id[
  candidate_table$IsProductionForestCell == 1L
]
write_ampl_vector(
  dat_temporary_path,
  "IsProductionForestCell",
  production_forest_cells,
  rep(1, length(production_forest_cells))
)

write_ampl_scalar(
  dat_temporary_path,
  "BiodiversityNormalizer",
  biodiversity_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "ContiguityNormalizer",
  contiguity_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "ExistingNatureContactNormalizer",
  existing_nature_contact_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "PatchConnectionNormalizer",
  patch_connection_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "SpatialContiguityBonus",
  spatial_contiguity_bonus
)
write_ampl_scalar(
  dat_temporary_path,
  "NewToNewAdjacencyWeight",
  new_to_new_adjacency_weight
)
write_ampl_scalar(
  dat_temporary_path,
  "ExistingToNewAdjacencyWeight",
  existing_to_new_adjacency_weight
)
write_ampl_scalar(
  dat_temporary_path,
  "ConnectedExistingPatchBonus",
  connected_existing_patch_bonus
)

dat_text <- readLines(dat_temporary_path, warn = FALSE)
forbidden_tokens <- grepl(
  "(^|[[:space:]])(NA|NaN|Inf|-Inf)([[:space:];]|$)",
  dat_text
)
stopifnot(!any(forbidden_tokens))
commit_generated_file(dat_temporary_path, dat_path)

# ---- Complete MIP start --------------------------------------------------
mip_start_temporary_path <- paste0(mip_start_path, ".building")
writeLines(
  c(
    "# Complete feasible start built from the successful notebook-07 footprint.",
    "let {c in Cells} Selected[c] := 0;",
    "let {(l,c) in AllowedActions} LanduseDecision[l,c] := 0;",
    "let {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j] := 0;",
    "let {k in RequiredNetworks, c in Cells} NetworkCell[k,c] := 0;",
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "NetworkPatch[k,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "NetworkRoot[k,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (i,j) in NetworkArcs} ",
      "AreaFlowHa[k,i,j] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (c,p) in CellPatchEdges} ",
      "CellToPatchFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (c,p) in CellPatchEdges} ",
      "PatchToCellFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "CollectedAreaHa[k,p] := 0;"
    )
  ),
  mip_start_temporary_path
)

append_start(
  mip_start_temporary_path,
  paste0(
    "let Selected[",
    warm_start_assignment$coarse_cell_id,
    "] := 1;"
  )
)
append_start(
  mip_start_temporary_path,
  warm_start_assignment |>
    transmute(line = paste0(
      "let LanduseDecision['",
      landuse,
      "',",
      coarse_cell_id,
      "] := 1;"
    )) |>
    pull(line)
)

start_labels_i <- warm_start_assignment |>
  transmute(landuse, i = coarse_cell_id)
start_labels_j <- warm_start_assignment |>
  transmute(landuse, j = coarse_cell_id)
start_contiguity <- same_habitat_edges |>
  semi_join(start_labels_i, by = c("landuse", "i")) |>
  semi_join(start_labels_j, by = c("landuse", "j"))
append_start(
  mip_start_temporary_path,
  paste0(
    "let Contiguity['",
    start_contiguity$landuse,
    "',",
    start_contiguity$i,
    ",",
    start_contiguity$j,
    "] := 1;"
  )
)

append_start(
  mip_start_temporary_path,
  start_certificate_cell_members |>
    transmute(line = paste0(
      "let NetworkCell['",
      network_name,
      "',",
      coarse_cell_id,
      "] := 1;"
    )) |>
    pull(line)
)
append_start(
  mip_start_temporary_path,
  start_certificate_patch_members |>
    transmute(line = paste0(
      "let NetworkPatch['",
      network_name,
      "',",
      patch_id,
      "] := 1;"
    )) |>
    pull(line)
)
append_start(
  mip_start_temporary_path,
  start_certificate_roots |>
    transmute(line = paste0(
      "let NetworkRoot['",
      network_name,
      "',",
      patch_id,
      "] := 1;"
    )) |>
    pull(line)
)

flow_start_lines <- start_certificate_flows |>
  mutate(line = case_when(
    child_type == "c" & parent_type == "c" ~ paste0(
      "let AreaFlowHa['",
      network_name,
      "',",
      child_id,
      ",",
      parent_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    child_type == "c" & parent_type == "p" ~ paste0(
      "let CellToPatchFlowHa['",
      network_name,
      "',",
      child_id,
      ",",
      parent_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    child_type == "p" & parent_type == "c" ~ paste0(
      "let PatchToCellFlowHa['",
      network_name,
      "',",
      parent_id,
      ",",
      child_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    TRUE ~ NA_character_
  ))
stopifnot(!anyNA(flow_start_lines$line))
append_start(mip_start_temporary_path, flow_start_lines$line)

append_start(
  mip_start_temporary_path,
  start_certificate_roots |>
    transmute(line = paste0(
      "let CollectedAreaHa['",
      network_name,
      "',",
      patch_id,
      "] := ",
      ampl_number(collected_area_ha),
      ";"
    )) |>
    pull(line)
)
commit_generated_file(mip_start_temporary_path, mip_start_path)
```

```{r export-run-and-audits}
run_lines <- c(
  "reset;",
  "",
  paste0('model "', ampl_path(model_path), '";'),
  paste0('data "', ampl_path(dat_path), '";'),
  paste0('include "', ampl_path(mip_start_path), '";'),
  "",
  paste("option solver", solver_name, ";"),
  if (nzchar(solver_options)) {
    paste0(
      "option ", solver_name, "_options '",
      ampl_path(solver_options), "';"
    )
  },
  "solve;",
  "display solve_result_num, solve_result;",
  "",
  paste0(
    'printf "coarse_cell_id\\tlanduse\\trestored_ha',
    '\\tbiodiversity\\texisting_nature_contact_edges\\n" > "',
    ampl_path(solution_path), '";'
  ),
  paste0(
    'printf {(l,c) in AllowedActions: ',
    'LanduseDecision[l,c] > 0.5} ',
    '"%d\\t%s\\t%.8f\\t%.8f\\t%.8f\\n", ',
    'c, l, RestorableAreaHa[c], BioDiversity[l,c], ',
    'ExistingNatureContact[l,c] >> "',
    ampl_path(solution_path), '";'
  ),
  "",
  paste0(
    'printf "subarea_id\\tcoarse_cell_id\\trestored_ha\\n" > "',
    ampl_path(selected_subarea_path), '";'
  ),
  paste0(
    'printf {s in Subareas, c in Cells: Selected[c] > 0.5 ',
    'and RestorableAreaHaBySubarea[s,c] > 0} ',
    '"%d\\t%d\\t%.8f\\n", s, c, ',
    'RestorableAreaHaBySubarea[s,c] >> "',
    ampl_path(selected_subarea_path), '";'
  ),
  "",
  paste0(
    'printf "subarea_id\\tpolygon_ha\\ttarget_ha\\tcurrent_ha',
    '\\trequired_restoration_ha\\trestored_ha\\tfinal_nature_ha',
    '\\tfinal_nature_percent\\n" > "',
    ampl_path(subarea_solution_path), '";'
  ),
  paste0(
    'printf {s in Subareas} ',
    '"%d\\t%.8f\\t%.8f\\t%.8f\\t%.8f\\t%.8f\\t%.8f',
    '\\t%.8f\\n", ',
    's, SubareaAreaHa[s], FinalNatureTargetHa[s], ',
    'CurrentNatureAreaHa[s], RestorationRequirementHa[s], ',
    'sum {c in Cells} RestorableAreaHaBySubarea[s,c] * Selected[c], ',
    'CurrentNatureAreaHa[s] + sum {c in Cells} ',
    'RestorableAreaHaBySubarea[s,c] * Selected[c], ',
    '100 * (CurrentNatureAreaHa[s] + sum {c in Cells} ',
    'RestorableAreaHaBySubarea[s,c] * Selected[c]) / ',
    'SubareaAreaHa[s] >> "', ampl_path(subarea_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tcoarse_cell_id\\tcounted_area_ha',
    '\\ttotal_restorable_ha\\n" > "',
    ampl_path(network_cell_solution_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks, c in Cells: ',
    'NetworkCell[k,c] > 0.5} ',
    '"%s\\t%d\\t%.8f\\t%.8f\\n", ',
    'k, c, NetworkCellAreaHa[k,c], RestorableAreaHa[c] >> "',
    ampl_path(network_cell_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tpatch_id\\tcounted_area_ha',
    '\\ttotal_patch_ha\\n" > "',
    ampl_path(network_patch_solution_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks, p in ExistingPatches: ',
    'NetworkPatch[k,p] > 0.5} ',
    '"%s\\t%d\\t%.8f\\t%.8f\\n", ',
    'k, p, NetworkPatchAreaHa[k,p], ExistingPatchAreaHa[p] >> "',
    ampl_path(network_patch_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tpatch_id\\tcollected_area_ha\\n" > "',
    ampl_path(network_root_solution_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks, p in ExistingPatches: ',
    'NetworkRoot[k,p] > 0.5} ',
    '"%s\\t%d\\t%.8f\\n", k, p, CollectedAreaHa[k,p] >> "',
    ampl_path(network_root_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tminimum_ha\\tcertified_area_ha',
    '\\tmember_cells\\tmember_patches\\troot_count',
    '\\tcollected_total_area_ha\\n" > "',
    ampl_path(network_solution_summary_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks} ',
    '"%s\\t%.8f\\t%.8f\\t%d\\t%d\\t%d\\t%.8f\\n", ',
    'k, NetworkMinimumHa[k], ',
    'sum {c in Cells} NetworkCellAreaHa[k,c] * NetworkCell[k,c] + ',
    'sum {p in ExistingPatches} NetworkPatchAreaHa[k,p] * ',
    'NetworkPatch[k,p], ',
    'sum {c in Cells} NetworkCell[k,c], ',
    'sum {p in ExistingPatches} NetworkPatch[k,p], ',
    'sum {p in ExistingPatches} NetworkRoot[k,p], ',
    'sum {p in ExistingPatches} CollectedAreaHa[k,p] >> "',
    ampl_path(network_solution_summary_path), '";'
  ),
  "",
  paste0(
    'printf "item\\tvalue\\n" > "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "solve_result\\t%s\\n", solve_result >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "solve_result_num\\t%d\\n", solve_result_num >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "objective\\t%.10f\\n", ConservationIndex >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_biodiversity\\t%.10f\\n", ',
    '(sum {(l,c) in AllowedActions} BioDiversity[l,c] * ',
    'LanduseDecision[l,c]) / BiodiversityNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_new_to_new_adjacency\\t%.10f\\n", ',
    '(sum {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j]) / ',
    'ContiguityNormalizer >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_existing_to_new_adjacency\\t%.10f\\n", ',
    '(sum {(l,c) in AllowedActions} ExistingNatureContact[l,c] * ',
    'LanduseDecision[l,c]) / ExistingNatureContactNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_existing_patches_joined_beyond_root\\t%.0f\\n", ',
    'sum {k in GlobalNetworks, p in ExistingPatches} ',
    'NetworkPatch[k,p] - 1 >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_existing_patch_connection\\t%.10f\\n", ',
    '(sum {k in GlobalNetworks, p in ExistingPatches} ',
    'NetworkPatch[k,p] - 1) / PatchConnectionNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "selected_cells\\t%.0f\\n", ',
    'sum {c in Cells} Selected[c] >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "restored_ha\\t%.8f\\n", ',
    'sum {c in Cells} RestorableAreaHa[c] * Selected[c] >> "',
    ampl_path(summary_path), '";'
  )
)
write_lines_if_changed(run_lines, run_path)

problem_audit <- tibble(
  polygon_area_ha = sum(subarea_budget$PolygonAreaHa),
  thirty_percent_target_ha =
    sum(subarea_budget$FinalNatureTargetHa),
  current_nature_ha = sum(subarea_budget$CurrentNatureAreaHa),
  required_new_restoration_ha =
    sum(subarea_budget$RestorationRequirementHa),
  candidate_cells = length(candidate_cells),
  allowed_action_binaries = nrow(allowed_actions),
  same_habitat_binaries = nrow(same_habitat_edges),
  candidate_rook_edges = nrow(candidate_edges),
  retained_existing_patches = length(existing_patches),
  candidate_patch_links = nrow(cell_patch_links),
  required_network_certificates = length(required_networks),
  certificate_membership_binaries =
    length(required_networks) *
      (length(candidate_cells) + 2 * length(existing_patches)),
  certificate_flow_variables =
    length(required_networks) *
      (2 * nrow(candidate_edges) +
         2 * nrow(cell_patch_links) +
         length(existing_patches)),
  warm_start_selected_cells = nrow(warm_start_assignment),
  warm_start_label_repairs =
    sum(warm_start_assignment$forest_rule_repair),
  spatial_contiguity_bonus = spatial_contiguity_bonus,
  connected_existing_patch_bonus = connected_existing_patch_bonus,
  patch_connection_normalizer = patch_connection_normalizer
)

readr::write_csv(
  problem_audit,
  file.path(problem_dir, "problem_size_and_policy_audit.csv")
)

export_audit <- tibble(
  file = c("model", "data", "MIP start", "run"),
  path = c(model_path, dat_path, mip_start_path, run_path),
  exists = file.exists(c(model_path, dat_path, mip_start_path, run_path)),
  size_mib = file.info(
    c(model_path, dat_path, mip_start_path, run_path)
  )$size / 1024^2
)

print(knitr::kable(t(problem_audit), digits = 3))
knitr::kable(export_audit, digits = 3)
```

# Run AMPL

From PowerShell or Command Prompt at the `BiodiversitySignatureProjects`
project root:

```{powershell, eval=FALSE}
ampl Optimization/gudenaa_new_area_200m_30pct_hard_networks/gudenaa_new_area_200m_30pct_hard_networks.run
```

Or from an open AMPL session at the project root:

```{ampl, eval=FALSE}
include "Optimization/gudenaa_new_area_200m_30pct_hard_networks/gudenaa_new_area_200m_30pct_hard_networks.run";
```

In the Gurobi log, look for a line stating that a user MIP start produced or
loaded a solution. A warm start supplies an incumbent; it does not alter the
hard constraints or guarantee a faster proof of optimality.

# Post-solve independent audit and maps

Re-knit this notebook after AMPL finishes. Matching patch checkpoints are
reused. Post-processing runs only when all required solution tables are newer
than the generated model inputs.

```{r post-solve}
required_solution_paths <- c(
  solution_path,
  subarea_solution_path,
  network_cell_solution_path,
  network_patch_solution_path,
  network_root_solution_path,
  network_solution_summary_path,
  summary_path
)

solution_is_current <- all(file.exists(required_solution_paths))
if (solution_is_current) {
  solution_time <- min(file.info(required_solution_paths)$mtime)
  input_time <- max(file.info(c(
    model_path,
    dat_path,
    mip_start_path,
    run_path
  ))$mtime)
  solution_is_current <- isTRUE(solution_time >= input_time)
}

if (!solution_is_current) {
  message(
    "No current hard-network solution found. Run the generated .run file, ",
    "then re-knit this notebook."
  )
} else {
  selected_solution <- readr::read_tsv(
    solution_path,
    show_col_types = FALSE
  ) |>
    mutate(coarse_cell_id = as.integer(coarse_cell_id))
  ampl_subarea_solution <- readr::read_tsv(
    subarea_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(subarea_id = as.integer(subarea_id))
  network_cell_solution <- readr::read_tsv(
    network_cell_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(coarse_cell_id = as.integer(coarse_cell_id))
  network_patch_solution <- readr::read_tsv(
    network_patch_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(patch_id = as.integer(patch_id))
  network_root_solution <- readr::read_tsv(
    network_root_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(patch_id = as.integer(patch_id))
  ampl_network_summary <- readr::read_tsv(
    network_solution_summary_path,
    show_col_types = FALSE
  )

  selected_ids <- selected_solution$coarse_cell_id
  stopifnot(
    nrow(selected_solution) > 0L,
    !anyDuplicated(selected_ids),
    all(selected_ids %in% candidate_cells),
    all(selected_solution$landuse %in% nature_landuses),
    setequal(ampl_network_summary$network_name, required_networks),
    setequal(network_root_solution$network_name, required_networks)
  )

  # ---- Four 30% constraints ---------------------------------------------
  independent_subarea_solution <-
    restorable_area_by_subarea_cell |>
    filter(coarse_cell_id %in% selected_ids) |>
    group_by(subarea_id) |>
    summarise(
      IndependentlyRestoredHa = sum(RestorableAreaHaBySubarea),
      .groups = "drop"
    ) |>
    right_join(subarea_budget, by = "subarea_id") |>
    mutate(
      IndependentlyRestoredHa = coalesce(IndependentlyRestoredHa, 0),
      FinalNatureAreaHa =
        CurrentNatureAreaHa + IndependentlyRestoredHa,
      FinalNaturePercent =
        100 * FinalNatureAreaHa / PolygonAreaHa,
      PassesLower =
        IndependentlyRestoredHa >=
          RestorationRequirementHa - 1e-6,
      PassesUpper =
        IndependentlyRestoredHa <=
          RestorationRequirementHa +
            RestorationToleranceHa + 1e-6
    ) |>
    arrange(subarea_id)

  if (!all(
    independent_subarea_solution$PassesLower &
      independent_subarea_solution$PassesUpper
  )) {
    stop("Independent 30% audit failed.", call. = FALSE)
  }
  if (max(abs(
    ampl_subarea_solution$restored_ha -
      independent_subarea_solution$IndependentlyRestoredHa
  )) > 1e-5) {
    stop("AMPL and R subarea totals disagree.", call. = FALSE)
  }

  # ---- Production-forest destination rule -------------------------------
  forest_rule_audit <- selected_solution |>
    select(coarse_cell_id, landuse) |>
    left_join(
      candidate_table |>
        select(
          coarse_cell_id,
          ProductionForestHa,
          AgricultureHa,
          ProductionForestShare,
          IsProductionForestCell
        ),
      by = "coarse_cell_id"
    ) |>
    mutate(
      PassesForestRule =
        IsProductionForestCell == 0L |
        landuse %in% forest_landuses
    )
  if (!all(forest_rule_audit$PassesForestRule)) {
    stop("Production-forest destination audit failed.", call. = FALSE)
  }

  # ---- Recalculate the five AMPL certificates ---------------------------
  independent_certificate_summary <- tibble(
    network_name = required_networks,
    RequiredAreaHa = unname(network_minimums[required_networks])
  ) |>
    left_join(
      bind_rows(
        network_cell_solution |>
          left_join(
            network_cell_area,
            by = c("network_name", "coarse_cell_id")
          ) |>
          transmute(
            network_name,
            area_ha = coalesce(NetworkCellAreaHa, 0),
            member_cells = 1L,
            member_patches = 0L
          ),
        network_patch_solution |>
          left_join(
            network_patch_area,
            by = c("network_name", "patch_id")
          ) |>
          transmute(
            network_name,
            area_ha = coalesce(NetworkPatchAreaHa, 0),
            member_cells = 0L,
            member_patches = 1L
          )
      ) |>
        group_by(network_name) |>
        summarise(
          IndependentlyCertifiedAreaHa = sum(area_ha),
          member_cells = sum(member_cells),
          member_patches = sum(member_patches),
          .groups = "drop"
        ),
      by = "network_name"
    ) |>
    left_join(
      network_root_solution |>
        count(network_name, name = "root_count"),
      by = "network_name"
    ) |>
    mutate(
      PassesArea =
        IndependentlyCertifiedAreaHa >= RequiredAreaHa - 1e-6,
      PassesOneRoot = root_count == 1L,
      PassesNewRestoration = member_cells >= 1L,
      PassesCoarseCertificate =
        PassesArea & PassesOneRoot & PassesNewRestoration
    )

  if (
    anyNA(independent_certificate_summary) ||
    !all(independent_certificate_summary$PassesCoarseCertificate)
  ) {
    stop("Independent AMPL certificate audit failed.", call. = FALSE)
  }
  if (any(!network_cell_solution$coarse_cell_id %in% selected_ids)) {
    stop("A certificate contains an unselected cell.", call. = FALSE)
  }
  if (max(abs(
    ampl_network_summary$certified_area_ha[
      match(
        independent_certificate_summary$network_name,
        ampl_network_summary$network_name
      )
    ] - independent_certificate_summary$IndependentlyCertifiedAreaHa
  )) > 1e-5) {
    stop("AMPL and R certificate totals disagree.", call. = FALSE)
  }

  readr::write_csv(
    independent_subarea_solution,
    file.path(problem_dir, "independent_subarea_solution_audit.csv")
  )
  readr::write_csv(
    forest_rule_audit,
    file.path(problem_dir, "production_forest_solution_audit.csv")
  )
  readr::write_csv(
    independent_certificate_summary,
    file.path(problem_dir, "independent_coarse_certificate_audit.csv")
  )

  # ---- Exact 10 m final-nature components -------------------------------
  fine_restoration <- terra::ifel(
    fine_eligible == 1 & coarse_id_on_fine %in% selected_ids,
    1,
    0
  )
  fine_restoration <- terra::mask(fine_restoration, fine_landuse)
  names(fine_restoration) <- "RestorationCells"

  final_nature <- terra::ifel(
    fine_nature == 1 | fine_restoration == 1,
    1,
    0
  )
  final_nature <- terra::mask(final_nature, fine_landuse)
  names(final_nature) <- "FinalNatureCells"

  exact_component_raster <- terra::patches(
    final_nature,
    directions = 4,
    zeroAsNA = TRUE
  )
  names(exact_component_raster) <- "component_id"

  exact_component_summary <- zonal_sum_table(
    c(fine_nature, fine_restoration, final_nature),
    exact_component_raster,
    "component_id"
  ) |>
    transmute(
      component_id = as.integer(round(component_id)),
      CurrentNatureAreaHa =
        CurrentNatureCells * fine_cell_area_ha,
      RestorationAreaHa =
        RestorationCells * fine_cell_area_ha,
      FinalNatureAreaHa =
        FinalNatureCells * fine_cell_area_ha,
      HasExistingNature = CurrentNatureAreaHa > 0,
      HasNewRestoration = RestorationAreaHa > 0,
      IsAnchoredRestorationComponent =
        HasExistingNature & HasNewRestoration
    ) |>
    arrange(desc(FinalNatureAreaHa))

  exact_pair_multiplier <- expected_number_of_subareas + 1L
  exact_component_subarea_zone <- terra::ifel(
    is.na(exact_component_raster) | is.na(subarea_id_on_fine),
    NA,
    exact_component_raster * exact_pair_multiplier +
      subarea_id_on_fine
  )

  exact_component_by_subarea <- zonal_sum_table(
    c(fine_nature, fine_restoration, final_nature),
    exact_component_subarea_zone,
    "pair_id"
  ) |>
    mutate(
      component_id = as.integer(floor(
        pair_id / exact_pair_multiplier
      )),
      subarea_id = as.integer(round(
        pair_id - component_id * exact_pair_multiplier
      )),
      CurrentNatureAreaHa =
        CurrentNatureCells * fine_cell_area_ha,
      RestorationAreaHa =
        RestorationCells * fine_cell_area_ha,
      FinalNatureAreaHa =
        FinalNatureCells * fine_cell_area_ha
    ) |>
    filter(
      subarea_id %in% seq_len(expected_number_of_subareas)
    ) |>
    select(
      component_id,
      subarea_id,
      CurrentNatureAreaHa,
      RestorationAreaHa,
      FinalNatureAreaHa
    ) |>
    left_join(
      exact_component_summary |>
        select(
          component_id,
          HasExistingNature,
          HasNewRestoration,
          IsAnchoredRestorationComponent
        ),
      by = "component_id"
    ) |>
    arrange(subarea_id, desc(FinalNatureAreaHa))

  exact_anchored_components <- exact_component_summary |>
    filter(IsAnchoredRestorationComponent)

  if (nrow(exact_anchored_components)) {
    exact_global_row <- exact_anchored_components |>
      slice_max(FinalNatureAreaHa, n = 1, with_ties = FALSE) |>
      transmute(
        scope = "whole_polygon",
        subarea_id = NA_integer_,
        component_id,
        RequiredAreaHa = global_network_minimum_ha,
        ExactFinalNatureAreaHa = FinalNatureAreaHa,
        ExactCurrentNatureAreaHa = CurrentNatureAreaHa,
        ExactRestorationAreaHa = RestorationAreaHa,
        PassesExact10m =
          ExactFinalNatureAreaHa >= RequiredAreaHa - 1e-6
      )
  } else {
    exact_global_row <- tibble(
      scope = "whole_polygon",
      subarea_id = NA_integer_,
      component_id = NA_integer_,
      RequiredAreaHa = global_network_minimum_ha,
      ExactFinalNatureAreaHa = 0,
      ExactCurrentNatureAreaHa = 0,
      ExactRestorationAreaHa = 0,
      PassesExact10m = FALSE
    )
  }

  exact_largest_by_subarea <- exact_component_by_subarea |>
    filter(IsAnchoredRestorationComponent) |>
    group_by(subarea_id) |>
    slice_max(FinalNatureAreaHa, n = 1, with_ties = FALSE) |>
    ungroup() |>
    transmute(
      subarea_id,
      component_id,
      ExactFinalNatureAreaHa = FinalNatureAreaHa,
      ExactCurrentNatureAreaHa = CurrentNatureAreaHa,
      ExactRestorationAreaHa = RestorationAreaHa
    )

  exact_network_goal_audit <- bind_rows(
    exact_global_row,
    tibble(
      subarea_id = seq_len(expected_number_of_subareas),
      scope = subarea_labels,
      RequiredAreaHa = subarea_network_minimum_ha
    ) |>
      left_join(exact_largest_by_subarea, by = "subarea_id") |>
      mutate(
        ExactFinalNatureAreaHa = coalesce(
          ExactFinalNatureAreaHa,
          0
        ),
        ExactCurrentNatureAreaHa = coalesce(
          ExactCurrentNatureAreaHa,
          0
        ),
        ExactRestorationAreaHa = coalesce(
          ExactRestorationAreaHa,
          0
        ),
        PassesExact10m =
          ExactFinalNatureAreaHa >= RequiredAreaHa - 1e-6
      ) |>
      select(
        scope,
        subarea_id,
        component_id,
        RequiredAreaHa,
        ExactFinalNatureAreaHa,
        ExactCurrentNatureAreaHa,
        ExactRestorationAreaHa,
        PassesExact10m
      )
  )

  readr::write_csv(
    exact_component_summary,
    file.path(problem_dir, "exact_10m_final_nature_components.csv")
  )
  readr::write_csv(
    exact_component_by_subarea,
    file.path(problem_dir, "exact_10m_component_by_subarea.csv")
  )
  readr::write_csv(
    exact_network_goal_audit,
    file.path(problem_dir, "exact_10m_network_goal_audit.csv")
  )

  # ---- Rasters for inspection and the next refinement -------------------
  selected_coarse <- coarse_cell_id
  selected_values <- rep(NA_integer_, terra::ncell(selected_coarse))
  selected_values[candidate_cells] <- 0L
  selected_values[selected_ids] <- 1L
  terra::values(selected_coarse) <- selected_values
  names(selected_coarse) <- "selected_200m"

  landuse_lookup <- tibble(
    warm_start_code = seq_along(nature_landuses),
    landuse = nature_landuses
  )
  destination_coarse <- coarse_cell_id
  destination_values <- rep(
    NA_integer_,
    terra::ncell(destination_coarse)
  )
  destination_values[selected_ids] <- landuse_lookup$warm_start_code[
    match(selected_solution$landuse, landuse_lookup$landuse)
  ]
  terra::values(destination_coarse) <- destination_values
  names(destination_coarse) <- "warm_start_landuse"

  destination_fine <- align_to_template(
    destination_coarse,
    fine_landuse,
    method = "near"
  )
  destination_fine <- terra::mask(
    destination_fine,
    fine_eligible_mask
  )

  terra::writeRaster(
    selected_coarse,
    file.path(problem_dir, "selected_coarse_cells_200m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES")
  )
  terra::writeRaster(
    fine_restoration,
    file.path(problem_dir, "hard_solution_restoration_10m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    final_nature,
    file.path(problem_dir, "hard_solution_final_nature_10m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    exact_component_raster,
    file.path(
      problem_dir,
      "hard_solution_final_nature_components_10m.tif"
    ),
    overwrite = TRUE,
    datatype = "INT4U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    destination_fine,
    file.path(
      problem_dir,
      "hard_solution_landuse_warm_start_10m.tif"
    ),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  readr::write_csv(
    landuse_lookup,
    file.path(problem_dir, "warm_start_landuse_lookup.csv")
  )

  print(independent_subarea_solution)
  print(independent_certificate_summary)
  print(exact_network_goal_audit)

  if (!all(exact_network_goal_audit$PassesExact10m)) {
    warning(
      "The AMPL 200 m certificates pass, but one or more exact 10 m ",
      "network goals fail. Use this result as the targeted input to a ",
      "10 m refinement; do not report the failed goal as achieved."
    )
  }
}
```

# Quick plots

```{r quick-plots, eval=FALSE}
result_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_30pct_hard_networks"
)

plot(terra::rast(file.path(
  result_dir,
  "hard_solution_landuse_warm_start_10m.tif"
)))

plot(terra::rast(file.path(
  result_dir,
  "hard_solution_final_nature_components_10m.tif"
)))
```

# Interpretation safeguards

1. The hard model certifies connectivity on the 200 m graph. The independent
   `exact_10m_network_goal_audit.csv` is the authority for fine-resolution
   reporting.
2. Existing and newly restored nature both count toward 5,000/1,000 ha.
   Every certificate contains at least one selected restoration cell and is
   rooted at one existing-nature patch.
3. A component may satisfy several subarea certificates. Its area counted for
   `Subarea3`, for example, is only the part physically inside subarea 3, even
   if the connecting route crosses another subarea.
4. The patch-connection term is a preference. The model must connect at least
   one existing patch because of anchoring, but it is not forced to connect two
   or more if that would make the other objectives worse.
5. The production-forest rule is a 200 m approximation: a unit is classified
   as production forest when at least 50% of its eligible area is that source
   class. A later 10 m model can enforce source-specific transitions exactly.
6. The previous warm-up folder is read-only. This stage writes only to
   `Optimization/gudenaa_new_area_200m_30pct_hard_networks/`; no old `.dat`
   file or output folder should be deleted.
