---
title: "New Gudenå area: 200 m global-30% feasible-network repair (08a-global30)"
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Purpose and exact interpretation

This is an explicit repair stage between the successful notebook-07 warm-up
and the final hard-network refinement. It does not overwrite notebook 07.
Instead, it gives Gurobi an immediately feasible starting point and solves
three objectives in strict order:

1. remove all 5,000/1,000 ha network shortfalls;
2. among feasible networks, minimize the area-weighted symmetric difference
   from the successful notebook-07 footprint; and
3. among equally small repairs, maximize biodiversity and the **same spatial
   contiguity objective used in notebook 07** (bonus 1.2 by default).

The repaired solution obeys these hard rules:

1. current nature plus restoration is **30% across the complete planning
   polygon**, allowing only a 16 ha raster/accounting tolerance above the
   target; there is no separate 30% requirement for any subarea;
2. at least one connected final-nature area contains 5,000 ha;
3. each subarea has at least 1,000 ha of one connected final-nature area inside
   that subarea;
4. every certified area contains new restoration and has exactly one root in
   an existing-nature patch;
5. connecting several distinct existing-nature patches in the certified
   5,000 ha network receives an additional normalized objective bonus; and
6. a production-forest-dominated 200 m unit can receive only a forest-nature
   destination.

Network area means **current plus newly restored nature**. For the Global
5,000 ha certificate, a reached existing-nature patch contributes its complete
national connected area, including the part outside the planning polygon. For
a Subarea certificate, only its hectares inside that delområde count toward
1,000 ha. The whole-polygon 30% calculation counts only land physically
inside the planning polygon. The four 1,000 ha certificates provide the hard
spatial-distribution safeguards that the former subarea-specific 30% bounds
provided.
Habitat types may connect to each other; connectivity is not restricted to one
habitat label.
The five certificates may overlap, so one physical cross-boundary component
may satisfy the global and several subarea requirements. This implements
"at least one" without silently requiring five disjoint reserves.

The repair optimization uses the established 200 m rook graph plus the earlier
model's buffered 200 m contact convention for current-nature patch supernodes.
It restricts each certificate to potential graph components that are large
enough to satisfy that certificate. This is an exact reduction, because one
connected certificate cannot cross between disconnected potential components.

After the solve, this notebook independently reconstructs final nature at 10 m
and audits the same 5,000/1,000 ha rules. A failed 10 m audit means the coarse
solution needs a targeted 10 m refinement; it must not be reported as an exact
fine-resolution network.

# Packages

```{r packages}
library(terra)
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(digest)
library(igraph)
library(knitr)
```

# Editable configuration

Only this section should normally need editing.

```{r configuration}
# ---- Spatial inputs ------------------------------------------------------
fine_landuse_path <- "Datasets/FinalLanduse.tif"
coarse_landuse_path <- file.path(
  "Predictors",
  "gudenaa_optimization",
  "Landuse_Gudenaa_200m_hierarchical.tif"
)
subareas_path <- file.path(
  "Datasets",
  "gudenaaplanen_afgreansning_15sep_delområder.shp"
)

# National current-nature patches preserve connectivity and full patch area
# beyond the new planning boundary. These are the same prepared inputs used by
# the earlier large-polygon anchored model.
national_patch_path <- file.path(
  "Predictors",
  "gudenaa_optimization",
  "Current_nature_patches_Denmark_10m.gpkg"
)
national_patch_layer <- "current_nature_patches"
national_patch_dictionary_path <- file.path(
  "Predictors",
  "gudenaa_optimization",
  "Current_nature_patch_dictionary_Denmark.csv"
)

# Geometry contact settings inherited from the earlier patch-supernode model.
contact_buffer_m <- 0.10
minimum_contact_proxy_m2 <- 0.10

subarea_label_field <- NULL
expected_number_of_subareas <- 4L
expected_subarea_areas_ha <- c(
  7991.14551899414,
  8653.25670292969,
  13009.8300014893,
  14526.539438208
)
expected_total_area_ha <- 44180.77
polygon_area_audit_tolerance_ha <- 5

# Needed to recreate exactly the biodiversity-supported eligibility mask used
# by notebook 07.
metrics_dir <- file.path(
  "Results",
  "gudenaa_plant_biodiversity",
  "full",
  "metrics"
)
biodiversity_support_metric <- "Richness"
biodiversity_support_scenario <- "current"

# ---- Upstream successful warm-up ----------------------------------------
warmup_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_30pct_existing_adjacency_thr1_forest_rule"
)

warmup_name <-
  "gudenaa_new_area_200m_30pct_existing_adjacency_thr1_forest_rule"

# Notebook 07 can legitimately reuse this matching baseline checkpoint
# without copying it into warmup_dir. Notebook 08 therefore resolves the
# biodiversity RDS from the warm-up folder first and this fallback second.
baseline_checkpoint_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_30pct"
)

# ---- Policy --------------------------------------------------------------
global_network_minimum_ha <- 5000
subarea_network_minimum_ha <- 1000
global_nature_target_fraction <- 0.30

# A narrow global upper tolerance prevents an unconstrained overshoot (for
# example, 50% nature) while retaining the successful notebook-07 footprint.
# Sixteen hectares equals the former 4 ha tolerance in each of four subareas.
global_restoration_tolerance_ha <- 16

# Keep the same area tolerance and spatial weights used by the warm-up.
spatial_contiguity_bonus <- 1.2
new_to_new_adjacency_weight <- 0.5
existing_to_new_adjacency_weight <- 0.5

# Strict lexicographic tolerances. Stage 3 may use only this much additional
# footprint-change area above the stage-2 incumbent; 0.01 ha is numerical
# slack, not permission for a materially larger repair.
network_shortfall_tolerance_ha <- 1e-6
repair_change_tolerance_ha <- 0.01

# Additional preference for joining several distinct existing-nature patches
# inside the certified global network. This is a normalized soft reward, not a
# hard requirement to connect a particular number of patches.
connected_existing_patch_bonus <- 0.25

# Established coarse rule: when at least half of a cell's eligible area is
# ProductionForest, the entire 200 m decision receives a forest destination.
production_forest_share_threshold <- 0.50

# ---- Outputs and solver --------------------------------------------------
problem_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_global30_feasible_network_repair"
)

problem_name <-
  "gudenaa_new_area_200m_global30_feasible_network_repair"

model_path <- "Gudenaa_global30_feasible_network_repair_200m.mod"

solver_name <- "gurobi"
repair_feasibility_time_limit_seconds <- 10800
repair_minimum_change_time_limit_seconds <- 7200
repair_quality_time_limit_seconds <- 3600
repair_mip_gap <- 0.005

solver_options_base <- paste(
  "tech:outlev=1",
  collapse = " "
)

reuse_matching_checkpoints <- TRUE
```

# Names and validation

```{r names-and-validation}
nature_landuses <- c(
  "ForestDryPoor",
  "ForestDryRich",
  "ForestWetPoor",
  "ForestWetRich",
  "OpenDryPoor",
  "OpenDryRich",
  "OpenWetPoor",
  "OpenWetRich"
)
forest_landuses <- c(
  "ForestDryPoor",
  "ForestDryRich",
  "ForestWetPoor",
  "ForestWetRich"
)

required_networks <- c(
  "Global",
  paste0("Subarea", seq_len(expected_number_of_subareas))
)
network_minimums <- c(
  Global = global_network_minimum_ha,
  stats::setNames(
    rep(subarea_network_minimum_ha, expected_number_of_subareas),
    paste0("Subarea", seq_len(expected_number_of_subareas))
  )
)

stopifnot(
  length(nature_landuses) == 8L,
  length(forest_landuses) == 4L,
  all(forest_landuses %in% nature_landuses),
  global_network_minimum_ha > 0,
  subarea_network_minimum_ha > 0,
  global_nature_target_fraction > 0,
  global_nature_target_fraction <= 1,
  global_restoration_tolerance_ha >= 0,
  spatial_contiguity_bonus >= 0,
  connected_existing_patch_bonus >= 0,
  new_to_new_adjacency_weight >= 0,
  existing_to_new_adjacency_weight >= 0,
  abs(
    new_to_new_adjacency_weight +
      existing_to_new_adjacency_weight - 1
  ) < 1e-10,
  production_forest_share_threshold > 0,
  production_forest_share_threshold <= 1,
  network_shortfall_tolerance_ha >= 0,
  repair_change_tolerance_ha >= 0,
  repair_feasibility_time_limit_seconds > 0,
  repair_minimum_change_time_limit_seconds > 0,
  repair_quality_time_limit_seconds > 0,
  repair_mip_gap >= 0,
  repair_mip_gap < 1
)
```

# Reusable helpers

```{r helpers}
shapefile_components <- function(path) {
  if (tolower(tools::file_ext(path)) != "shp") return(path)
  stem <- tools::file_path_sans_ext(path)
  paste0(stem, c(".shp", ".shx", ".dbf", ".prj"))
}

stop_for_missing_files <- function(paths, description) {
  paths <- unique(paths)
  missing <- paths[!file.exists(paths)]
  if (length(missing)) {
    stop(
      paste0(
        "Missing ", description, ":\n- ",
        paste(missing, collapse = "\n- ")
      ),
      call. = FALSE
    )
  }
  invisible(paths)
}

same_crs <- function(x, y) isTRUE(terra::same.crs(x, y))

align_to_template <- function(x, template, method) {
  if (!same_crs(x, template)) {
    x <- terra::project(x, template, method = method)
  } else if (!isTRUE(terra::compareGeom(
    x,
    template,
    lyrs = FALSE,
    stopOnError = FALSE
  ))) {
    x <- terra::resample(x, template, method = method)
  }
  terra::mask(x, template)
}

category_codes <- function(x, requested_labels) {
  lev <- terra::levels(x)[[1]]
  if (is.null(lev) || !nrow(lev)) {
    stop("The fine land-use raster has no category table.", call. = FALSE)
  }

  numeric_columns <- names(lev)[vapply(lev, is.numeric, logical(1))]
  label_columns <- names(lev)[vapply(
    lev,
    function(z) is.character(z) || is.factor(z),
    logical(1)
  )]
  if (!length(numeric_columns) || !length(label_columns)) {
    stop("Could not identify the land-use category columns.", call. = FALSE)
  }

  labels <- as.character(lev[[label_columns[[1]]]])
  missing_labels <- setdiff(requested_labels, labels)
  if (length(missing_labels)) {
    stop(
      "Missing land-use classes: ",
      paste(missing_labels, collapse = ", "),
      call. = FALSE
    )
  }

  stats::setNames(
    lev[[numeric_columns[[1]]]][
      match(requested_labels, labels)
    ],
    requested_labels
  )
}

indicator_for_codes <- function(x, codes) {
  terra::ifel(
    is.na(x),
    NA,
    terra::ifel(x %in% codes, 1, 0)
  )
}

zonal_sum_table <- function(x, zones, id_name) {
  out <- terra::zonal(
    x,
    zones,
    fun = "sum",
    na.rm = TRUE,
    wide = TRUE
  ) |>
    as.data.frame()
  names(out)[1] <- id_name
  out[[id_name]] <- as.numeric(out[[id_name]])
  out
}

global_number <- function(x, fun) {
  out <- if (fun %in% c("notNA", "isNA")) {
    terra::global(x, fun)
  } else {
    terra::global(x, fun, na.rm = TRUE)
  }
  as.numeric(out[1, 1])
}

unwrap_checkpoint <- function(path) {
  x <- readRDS(path)
  if (is.list(x) && "value" %in% names(x)) x$value else x
}

file_fingerprint <- function(paths) {
  paths <- normalizePath(paths, winslash = "/", mustWork = FALSE)
  info <- file.info(paths)
  tibble(path = paths, size = info$size, modified = as.numeric(info$mtime))
}

load_or_build_checkpoint <- function(
    path,
    signature,
    builder,
    required_sidecars = character()
) {
  if (
    reuse_matching_checkpoints &&
    file.exists(path) &&
    all(file.exists(required_sidecars))
  ) {
    checkpoint <- readRDS(path)
    if (
      is.list(checkpoint) &&
      identical(checkpoint$signature, signature) &&
      !is.null(checkpoint$value)
    ) {
      return(checkpoint$value)
    }
  }
  value <- builder()
  saveRDS(list(signature = signature, value = value), path)
  value
}

ampl_number <- function(x) {
  format(x, scientific = FALSE, trim = TRUE, digits = 15)
}

append_dat <- function(path, lines) {
  cat(
    paste(lines, collapse = "\n"),
    "\n",
    file = path,
    append = TRUE,
    sep = ""
  )
}

chunk_text <- function(x, chunk_size = 20L) {
  if (!length(x)) return(character())
  groups <- ceiling(seq_along(x) / chunk_size)
  vapply(
    split(x, groups),
    paste,
    collapse = " ",
    FUN.VALUE = character(1)
  )
}

write_ampl_set <- function(path, name, values) {
  append_dat(
    path,
    c(paste("set", name, ":="), chunk_text(values), ";", "")
  )
}

write_ampl_pair_set <- function(path, name, first, second) {
  tuples <- paste0("(", first, ",", second, ")")
  append_dat(
    path,
    c(
      paste("set", name, ":="),
      chunk_text(tuples, chunk_size = 8L),
      ";",
      ""
    )
  )
}

write_ampl_triple_set <- function(path, name, first, second, third) {
  tuples <- paste0("(", first, ",", second, ",", third, ")")
  append_dat(
    path,
    c(
      paste("set", name, ":="),
      chunk_text(tuples, chunk_size = 5L),
      ";",
      ""
    )
  )
}

write_ampl_scalar <- function(path, name, value) {
  append_dat(
    path,
    c(paste0("param ", name, " := ", ampl_number(value), ";"), "")
  )
}

write_ampl_vector <- function(path, name, key, value) {
  rows <- paste(key, ampl_number(value))
  append_dat(path, c(paste("param", name, ":="), rows, ";", ""))
}

write_ampl_matrix <- function(path, name, row_key, column_key, value) {
  rows <- paste(row_key, column_key, ampl_number(value))
  append_dat(path, c(paste("param", name, ":="), rows, ";", ""))
}

commit_generated_file <- function(temporary_path, final_path) {
  unchanged <- file.exists(final_path) &&
    unname(tools::md5sum(temporary_path)) ==
      unname(tools::md5sum(final_path))
  if (unchanged) {
    unlink(temporary_path)
    return(invisible(FALSE))
  }
  copied <- file.copy(temporary_path, final_path, overwrite = TRUE)
  unlink(temporary_path)
  if (!copied) stop("Could not write: ", final_path, call. = FALSE)
  invisible(TRUE)
}

write_lines_if_changed <- function(lines, path) {
  temporary_path <- paste0(path, ".building")
  writeLines(lines, temporary_path)
  commit_generated_file(temporary_path, path)
}

ampl_path <- function(path) chartr(intToUtf8(92L), "/", path)

append_start <- function(path, lines) {
  if (!length(lines)) return(invisible(NULL))
  cat(
    paste(lines, collapse = "\n"),
    "\n",
    file = path,
    append = TRUE,
    sep = ""
  )
}
```

# Read and validate the successful warm-up

```{r read-upstream}
dir.create(problem_dir, recursive = TRUE, showWarnings = FALSE)

warmup_selected_path <- file.path(
  warmup_dir,
  paste0(warmup_name, "_selected_cells.tsv")
)
warmup_summary_path <- file.path(
  warmup_dir,
  paste0(warmup_name, "_solution_summary.tsv")
)
candidate_table_path <- file.path(
  warmup_dir,
  "candidate_cell_dictionary.csv"
)
candidate_subarea_path <- file.path(
  warmup_dir,
  "candidate_area_by_subarea.csv"
)
candidate_edges_path <- file.path(
  warmup_dir,
  "candidate_rook_edges.csv"
)
subarea_budget_path <- file.path(
  warmup_dir,
  "subarea_30pct_budget.csv"
)
existing_contact_path <- file.path(
  warmup_dir,
  "existing_nature_contact_by_action.csv"
)
biodiversity_checkpoint_candidates <- c(
  file.path(warmup_dir, "coarse_biodiversity_benefits.rds"),
  file.path(
    baseline_checkpoint_dir,
    "coarse_biodiversity_benefits.rds"
  )
)
biodiversity_checkpoint_paths_found <-
  biodiversity_checkpoint_candidates[
    file.exists(biodiversity_checkpoint_candidates)
  ]
if (!length(biodiversity_checkpoint_paths_found)) {
  stop(
    paste0(
      "No biodiversity checkpoint was found. Expected either:\n- ",
      paste(biodiversity_checkpoint_candidates, collapse = "\n- ")
    ),
    call. = FALSE
  )
}
biodiversity_checkpoint_path <-
  biodiversity_checkpoint_paths_found[[1]]

biodiversity_support_path <- file.path(
  metrics_dir,
  paste0(
    biodiversity_support_metric,
    "_",
    biodiversity_support_scenario,
    ".tif"
  )
)

required_paths <- c(
  fine_landuse_path,
  coarse_landuse_path,
  shapefile_components(subareas_path),
  national_patch_path,
  national_patch_dictionary_path,
  biodiversity_support_path,
  model_path,
  warmup_selected_path,
  warmup_summary_path,
  candidate_table_path,
  candidate_subarea_path,
  candidate_edges_path,
  subarea_budget_path,
  existing_contact_path,
  biodiversity_checkpoint_path
)
stop_for_missing_files(required_paths, "08a repair input file(s)")

warmup_summary <- readr::read_tsv(
  warmup_summary_path,
  show_col_types = FALSE,
  col_types = cols(.default = col_character())
)
if (!any(str_detect(tolower(warmup_summary$value), "solved|optimal"))) {
  stop(
    "The upstream summary does not report a solved/optimal warm-up.",
    call. = FALSE
  )
}

# Make the repaired output folder a self-contained upstream folder for the
# subsequent 08b refinement. These are immutable support inputs; solution
# tables are written later by AMPL under problem_name.
downstream_support_sources <- c(
  candidate_table_path,
  candidate_subarea_path,
  candidate_edges_path,
  subarea_budget_path,
  existing_contact_path,
  biodiversity_checkpoint_path
)
downstream_support_names <- c(
  "candidate_cell_dictionary.csv",
  "candidate_area_by_subarea.csv",
  "candidate_rook_edges.csv",
  "subarea_30pct_budget.csv",
  "existing_nature_contact_by_action.csv",
  "coarse_biodiversity_benefits.rds"
)
downstream_support_destinations <- file.path(
  problem_dir,
  downstream_support_names
)
support_copy_ok <- map2_lgl(
  downstream_support_sources,
  downstream_support_destinations,
  function(source, destination) {
    if (
      file.exists(destination) &&
        identical(
          unname(tools::md5sum(source)),
          unname(tools::md5sum(destination))
        )
    ) {
      return(TRUE)
    }
    isTRUE(file.copy(source, destination, overwrite = TRUE))
  }
)
if (!all(support_copy_ok)) {
  stop(
    "Could not copy all downstream warm-start support files into problem_dir.",
    call. = FALSE
  )
}

warmup_selected <- readr::read_tsv(
  warmup_selected_path,
  show_col_types = FALSE
) |>
  transmute(
    coarse_cell_id = as.integer(coarse_cell_id),
    landuse = as.character(landuse)
  )

candidate_table <- readr::read_csv(
  candidate_table_path,
  show_col_types = FALSE
) |>
  mutate(
    coarse_cell_id = as.integer(coarse_cell_id),
    ProductionForestShare = if_else(
      EligibleAreaHa > 0,
      ProductionForestHa / EligibleAreaHa,
      0
    ),
    IsProductionForestCell = as.integer(
      ProductionForestShare >= production_forest_share_threshold
    )
  ) |>
  filter(IsCandidate) |>
  arrange(coarse_cell_id)

candidate_cells <- candidate_table$coarse_cell_id

restorable_area_by_subarea_cell <- readr::read_csv(
  candidate_subarea_path,
  show_col_types = FALSE
) |>
  transmute(
    subarea_id = as.integer(subarea_id),
    coarse_cell_id = as.integer(coarse_cell_id),
    RestorableAreaHaBySubarea = as.numeric(
      RestorableAreaHaBySubarea
    )
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

candidate_edges <- readr::read_csv(
  candidate_edges_path,
  show_col_types = FALSE
) |>
  transmute(i = as.integer(i), j = as.integer(j)) |>
  filter(i %in% candidate_cells, j %in% candidate_cells) |>
  distinct() |>
  arrange(i, j)

subarea_budget <- readr::read_csv(
  subarea_budget_path,
  show_col_types = FALSE
) |>
  mutate(subarea_id = as.integer(subarea_id)) |>
  arrange(subarea_id)

global_budget <- subarea_budget |>
  summarise(
    GlobalPolygonAreaHa = sum(PolygonAreaHa),
    GlobalCurrentNatureAreaHa = sum(CurrentNatureAreaHa)
  ) |>
  mutate(
    GlobalNatureTargetFraction = global_nature_target_fraction,
    GlobalFinalNatureTargetHa =
      GlobalNatureTargetFraction * GlobalPolygonAreaHa,
    GlobalRestorationRequirementHa = pmax(
      0,
      GlobalFinalNatureTargetHa - GlobalCurrentNatureAreaHa
    ),
    GlobalRestorationToleranceHa = global_restoration_tolerance_ha,
    GlobalMaximumRestorationHa =
      GlobalRestorationRequirementHa + GlobalRestorationToleranceHa,
    GlobalMaximumFinalNatureHa =
      GlobalFinalNatureTargetHa + GlobalRestorationToleranceHa,
    GlobalMaximumFinalNaturePercent =
      100 * GlobalMaximumFinalNatureHa / GlobalPolygonAreaHa
  )

readr::write_csv(
  global_budget,
  file.path(problem_dir, "whole_polygon_30pct_budget.csv")
)

existing_nature_contact <- readr::read_csv(
  existing_contact_path,
  show_col_types = FALSE
) |>
  transmute(
    landuse = as.character(landuse),
    coarse_cell_id = as.integer(coarse_cell_id),
    ExistingNatureContact = as.numeric(ExistingNatureContact)
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

biodiversity_checkpoint <- unwrap_checkpoint(
  biodiversity_checkpoint_path
)
if (
  !is.list(biodiversity_checkpoint) ||
  !"benefits" %in% names(biodiversity_checkpoint)
) {
  stop("Unexpected biodiversity checkpoint structure.", call. = FALSE)
}
biodiversity_table <- biodiversity_checkpoint$benefits |>
  transmute(
    coarse_cell_id = as.integer(coarse_cell_id),
    landuse = as.character(landuse),
    BioDiversity = as.numeric(BioDiversity)
  ) |>
  filter(coarse_cell_id %in% candidate_cells)

stopifnot(
  nrow(subarea_budget) == expected_number_of_subareas,
  nrow(warmup_selected) > 0L,
  !anyDuplicated(warmup_selected$coarse_cell_id),
  all(warmup_selected$coarse_cell_id %in% candidate_cells),
  all(warmup_selected$landuse %in% nature_landuses),
  length(candidate_cells) > 0L,
  nrow(candidate_edges) > 0L,
  nrow(biodiversity_table) ==
    length(candidate_cells) * length(nature_landuses),
  nrow(existing_nature_contact) ==
    length(candidate_cells) * length(nature_landuses),
  !anyNA(biodiversity_table),
  !anyNA(existing_nature_contact)
)

input_audit <- file_fingerprint(required_paths)
readr::write_csv(
  input_audit,
  file.path(problem_dir, "input_file_audit.csv")
)
readr::write_csv(
  tibble(
    candidate_path = biodiversity_checkpoint_candidates,
    exists = file.exists(biodiversity_checkpoint_candidates),
    selected = biodiversity_checkpoint_candidates ==
      biodiversity_checkpoint_path
  ),
  file.path(problem_dir, "biodiversity_checkpoint_resolution.csv")
)
knitr::kable(input_audit)
```

# Reconstruct the aligned 10 m and 200 m domain

```{r aligned-domain}
fine_landuse_full <- terra::rast(fine_landuse_path)
coarse_landuse_full <- terra::rast(coarse_landuse_path)

stopifnot(
  terra::nlyr(fine_landuse_full) == 1L,
  terra::nlyr(coarse_landuse_full) == 1L,
  all(abs(terra::res(fine_landuse_full) - 10) < 1e-8),
  all(abs(terra::res(coarse_landuse_full) - 200) < 1e-8),
  same_crs(fine_landuse_full, coarse_landuse_full)
)

fine_cell_area_ha <- prod(terra::res(fine_landuse_full)) / 10000

subareas <- terra::vect(subareas_path) |>
  terra::makeValid()
if (nrow(subareas) != expected_number_of_subareas) {
  stop(
    "Expected four subareas; found ", nrow(subareas), ".",
    call. = FALSE
  )
}
if (!same_crs(subareas, fine_landuse_full)) {
  subareas <- terra::project(subareas, terra::crs(fine_landuse_full))
}

if (is.null(subarea_label_field)) {
  subarea_labels <- paste0("subarea_", seq_len(nrow(subareas)))
} else {
  if (!subarea_label_field %in% names(subareas)) {
    stop("Invalid subarea_label_field.", call. = FALSE)
  }
  subarea_labels <- make.unique(as.character(
    as.data.frame(subareas)[[subarea_label_field]]
  ))
}

subareas$subarea_id <- seq_len(nrow(subareas))
subareas$subarea_label <- subarea_labels
polygon_areas_ha <- as.numeric(
  terra::expanse(subareas, unit = "ha", transform = FALSE)
)
planning_domain <- terra::aggregate(subareas)

if (
  any(
    abs(sort(polygon_areas_ha) - sort(expected_subarea_areas_ha)) >
      polygon_area_audit_tolerance_ha
  ) ||
  abs(sum(polygon_areas_ha) - expected_total_area_ha) >
    polygon_area_audit_tolerance_ha
) {
  stop(
    "The polygon areas do not match the intended smaller Gudenå polygon.",
    call. = FALSE
  )
}

fine_landuse <- terra::crop(
  fine_landuse_full,
  planning_domain,
  snap = "out"
) |>
  terra::mask(planning_domain, touches = FALSE)

coarse_landuse <- terra::crop(
  coarse_landuse_full,
  planning_domain,
  snap = "out"
) |>
  terra::mask(planning_domain, touches = TRUE)

coarse_cell_id <- terra::init(coarse_landuse, "cell")
coarse_cell_id <- terra::ifel(
  is.na(coarse_landuse),
  NA,
  coarse_cell_id
)
names(coarse_cell_id) <- "coarse_cell_id"

coarse_id_on_fine <- align_to_template(
  coarse_cell_id,
  fine_landuse,
  method = "near"
)
names(coarse_id_on_fine) <- "coarse_cell_id"

subarea_id_on_fine <- terra::rasterize(
  subareas[, "subarea_id"],
  fine_landuse,
  field = "subarea_id",
  background = NA,
  touches = FALSE
)
subarea_id_on_fine <- terra::mask(
  subarea_id_on_fine,
  fine_landuse
)
names(subarea_id_on_fine) <- "subarea_id"

all_required_classes <- c(
  nature_landuses,
  "Agriculture",
  "ProductionForest",
  "Urban",
  "Other"
)
landuse_codes <- category_codes(fine_landuse, all_required_classes)
nature_codes <- unname(landuse_codes[nature_landuses])
eligible_codes <- unname(
  landuse_codes[c("Agriculture", "ProductionForest")]
)

fine_nature <- indicator_for_codes(fine_landuse, nature_codes)
names(fine_nature) <- "CurrentNatureCells"

biodiversity_support <- terra::rast(biodiversity_support_path)
biodiversity_support <- align_to_template(
  biodiversity_support,
  fine_landuse,
  method = "bilinear"
)
fine_biodiversity_support <- terra::ifel(
  is.na(fine_landuse),
  NA,
  terra::ifel(!is.na(biodiversity_support), 1, 0)
)

fine_eligible <-
  indicator_for_codes(fine_landuse, eligible_codes) *
  fine_biodiversity_support
fine_eligible_mask <- terra::ifel(fine_eligible == 1, 1, NA)
names(fine_eligible_mask) <- "EligibleCells"

candidate_mask_on_fine <- terra::ifel(
  coarse_id_on_fine %in% candidate_cells,
  1,
  NA
)
fine_candidate_eligible_mask <- terra::mask(
  fine_eligible_mask,
  candidate_mask_on_fine
)

if (
  global_number(
    terra::ifel(
      !is.na(fine_landuse) & is.na(coarse_id_on_fine),
      1,
      0
    ),
    "sum"
  ) > 0
) {
  stop("Some valid 10 m cells have no 200 m ID.", call. = FALSE)
}

coarse_ids_reconstructed <- sort(unique(na.omit(as.integer(
  terra::values(coarse_cell_id, mat = FALSE)
))))
if (!all(candidate_cells %in% coarse_ids_reconstructed)) {
  stop(
    "Reconstructed coarse IDs do not match notebook 07 outputs.",
    call. = FALSE
  )
}
```

# National current-nature patches and candidate contacts

The whole-polygon 30% accounting above remains clipped to the union of the
four planning subareas. The subarea percentages are descriptive only. The
network accounting deliberately uses a different quantity: the **complete
national area** of each connected existing-nature patch reached by selected
restoration. Thus a patch crossing the new boundary is not cut at that
boundary for the Global 5,000 ha certificate. Its clipped hectares are also
retained separately for the four Subarea certificates.

```{r patch-network}
national_patches <- terra::vect(
  national_patch_path,
  layer = national_patch_layer
) |>
  terra::makeValid()

if (!same_crs(national_patches, coarse_landuse)) {
  national_patches <- terra::project(
    national_patches,
    terra::crs(coarse_landuse)
  )
}
if (!"patch_id" %in% names(national_patches)) {
  stop("The national patch layer has no patch_id field.", call. = FALSE)
}
national_patches$patch_id <- as.integer(national_patches$patch_id)

patch_dictionary_full <- readr::read_csv(
  national_patch_dictionary_path,
  show_col_types = FALSE
) |>
  transmute(
    patch_id = as.integer(patch_id),
    ExistingPatchAreaHa = as.numeric(area_network_ha)
  )

stopifnot(
  !anyDuplicated(patch_dictionary_full$patch_id),
  all(is.finite(patch_dictionary_full$ExistingPatchAreaHa)),
  all(patch_dictionary_full$ExistingPatchAreaHa > 0)
)

patches_near_domain <- terra::crop(
  national_patches,
  terra::buffer(planning_domain, width = contact_buffer_m)
)
if (!nrow(patches_near_domain)) {
  stop("No national current-nature patches occur near the domain.")
}

# Candidate planning units become individual 200 m polygons. This preserves
# the established coarse contact convention from notebook 04.
candidate_id_raster <- coarse_cell_id
candidate_id_values <- rep(NA_real_, terra::ncell(candidate_id_raster))
candidate_id_values[candidate_cells] <- candidate_cells
terra::values(candidate_id_raster) <- candidate_id_values
names(candidate_id_raster) <- "coarse_cell_id"

candidate_polygons <- terra::as.polygons(
  candidate_id_raster,
  aggregate = FALSE,
  values = TRUE,
  na.rm = TRUE
)
candidate_polygons$coarse_cell_id <- as.integer(
  candidate_polygons$coarse_cell_id
)

# Exact patch area inside each delområde. These coefficients are distinct from
# the full national area used by the Global certificate.
patch_fragments_by_subarea <- terra::intersect(
  patches_near_domain[, "patch_id"],
  subareas[, "subarea_id"]
)
if (!nrow(patch_fragments_by_subarea)) {
  stop("No national patch fragments intersect the four subareas.")
}
patch_fragments_by_subarea$fragment_area_ha <- terra::expanse(
  patch_fragments_by_subarea,
  unit = "ha",
  transform = FALSE
)

patch_area_by_subarea_all <- patch_fragments_by_subarea |>
  as.data.frame() |>
  transmute(
    patch_id = as.integer(patch_id),
    subarea_id = as.integer(subarea_id),
    fragment_area_ha = as.numeric(fragment_area_ha)
  ) |>
  group_by(patch_id, subarea_id) |>
  summarise(
    ExistingPatchAreaHaBySubarea = sum(fragment_area_ha),
    .groups = "drop"
  ) |>
  filter(ExistingPatchAreaHaBySubarea > 0)

national_current_by_subarea <- patch_area_by_subarea_all |>
  group_by(subarea_id) |>
  summarise(
    NationalPatchNatureHa = sum(ExistingPatchAreaHaBySubarea),
    .groups = "drop"
  ) |>
  right_join(
    subarea_budget |>
      select(subarea_id, LanduseRasterNatureHa = CurrentNatureAreaHa),
    by = "subarea_id"
  ) |>
  mutate(
    NationalPatchNatureHa = coalesce(NationalPatchNatureHa, 0),
    DifferenceHa = NationalPatchNatureHa - LanduseRasterNatureHa,
    DifferencePercent =
      100 * DifferenceHa / pmax(LanduseRasterNatureHa, 1e-9)
  ) |>
  arrange(subarea_id)

patch_signature <- digest::digest(list(
  files = file_fingerprint(c(
    coarse_landuse_path,
    shapefile_components(subareas_path),
    national_patch_path,
    national_patch_dictionary_path,
    candidate_table_path,
    candidate_edges_path
  )),
  national_patch_layer = national_patch_layer,
  candidate_cells = candidate_cells,
  contact_buffer_m = contact_buffer_m,
  minimum_contact_proxy_m2 = minimum_contact_proxy_m2
))

patch_network <- load_or_build_checkpoint(
  file.path(problem_dir, "existing_patch_network.rds"),
  patch_signature,
  function() {
    candidate_contact_buffers <- terra::buffer(
      candidate_polygons,
      width = contact_buffer_m
    )
    contacts <- terra::intersect(
      candidate_contact_buffers,
      patches_near_domain[, "patch_id"]
    )
    if (!nrow(contacts)) {
      stop("No candidate-to-current-patch contacts were found.")
    }

    contacts$contact_proxy_m2 <- terra::expanse(
      contacts,
      unit = "m",
      transform = FALSE
    )
    links <- contacts |>
      as.data.frame() |>
      transmute(
        coarse_cell_id = as.integer(coarse_cell_id),
        patch_id = as.integer(patch_id),
        contact_proxy_m2 = as.numeric(contact_proxy_m2)
      ) |>
      group_by(coarse_cell_id, patch_id) |>
      summarise(
        contact_proxy_m2 = sum(contact_proxy_m2),
        .groups = "drop"
      ) |>
      filter(
        contact_proxy_m2 >= minimum_contact_proxy_m2,
        coarse_cell_id %in% candidate_cells
      ) |>
      arrange(coarse_cell_id, patch_id)

    linked_patch_areas <- links |>
      distinct(patch_id) |>
      left_join(patch_dictionary_full, by = "patch_id")
    if (anyNA(linked_patch_areas$ExistingPatchAreaHa)) {
      missing_ids <- linked_patch_areas$patch_id[
        is.na(linked_patch_areas$ExistingPatchAreaHa)
      ]
      stop(
        "Missing full national area for patch IDs: ",
        paste(head(missing_ids, 20), collapse = ", "),
        call. = FALSE
      )
    }

    list(
      patches = linked_patch_areas |> arrange(patch_id),
      links = links
    )
  }
)

patch_table <- patch_network$patches
cell_patch_links <- patch_network$links |>
  select(coarse_cell_id, patch_id) |>
  distinct()
existing_patches <- patch_table$patch_id
patch_area_by_subarea <- patch_area_by_subarea_all |>
  filter(patch_id %in% existing_patches) |>
  arrange(subarea_id, patch_id)

stopifnot(
  nrow(patch_table) > 0L,
  nrow(cell_patch_links) > 0L,
  all(cell_patch_links$coarse_cell_id %in% candidate_cells),
  all(cell_patch_links$patch_id %in% existing_patches),
  all(patch_table$ExistingPatchAreaHa > 0),
  !anyDuplicated(patch_table$patch_id),
  !anyDuplicated(cell_patch_links)
)

patch_area_crosscheck <- patch_area_by_subarea |>
  group_by(patch_id) |>
  summarise(
    ExistingPatchAreaInsidePlanningHa =
      sum(ExistingPatchAreaHaBySubarea),
    .groups = "drop"
  ) |>
  right_join(patch_table, by = "patch_id") |>
  mutate(
    ExistingPatchAreaInsidePlanningHa = coalesce(
      ExistingPatchAreaInsidePlanningHa,
      0
    ),
    ExistingPatchAreaOutsidePlanningHa = pmax(
      0,
      ExistingPatchAreaHa - ExistingPatchAreaInsidePlanningHa
    ),
    OutsideSharePercent =
      100 * ExistingPatchAreaOutsidePlanningHa / ExistingPatchAreaHa
  )

if (any(
  patch_area_crosscheck$ExistingPatchAreaInsidePlanningHa >
    patch_area_crosscheck$ExistingPatchAreaHa + 0.1
)) {
  stop(
    "A clipped patch area exceeds its full national dictionary area.",
    call. = FALSE
  )
}

# Keep the scopes explicit. national_current_by_subarea contains every
# national existing-nature patch fragment inside each delområde, whereas the
# certificate ledger below contains only patches that can actually be reached
# from at least one candidate restoration cell under the contact convention.
# The difference is expected and must never be silently counted as reachable
# certificate area.
linked_patch_area_by_subarea <- patch_area_by_subarea |>
  group_by(subarea_id) |>
  summarise(
    LinkedCertificatePatchNatureHa =
      sum(ExistingPatchAreaHaBySubarea),
    .groups = "drop"
  )

patch_ledger_scope_audit <- national_current_by_subarea |>
  left_join(linked_patch_area_by_subarea, by = "subarea_id") |>
  mutate(
    LinkedCertificatePatchNatureHa = coalesce(
      LinkedCertificatePatchNatureHa,
      0
    ),
    UnlinkedNationalPatchNatureHa =
      NationalPatchNatureHa - LinkedCertificatePatchNatureHa,
    LinkedDoesNotExceedAll =
      LinkedCertificatePatchNatureHa <= NationalPatchNatureHa + 1e-6,
    MaterialPatchVsLanduseDifference =
      abs(DifferencePercent) > 5
  ) |>
  arrange(subarea_id)

if (!all(patch_ledger_scope_audit$LinkedDoesNotExceedAll)) {
  stop(
    "Linked certificate-patch area exceeds the all-patch ledger. ",
    "Delete the 08a output folder and rebuild its checkpoints.",
    call. = FALSE
  )
}

linked_inside_from_fragments <-
  sum(patch_area_by_subarea$ExistingPatchAreaHaBySubarea)
linked_inside_from_patch_audit <-
  sum(patch_area_crosscheck$ExistingPatchAreaInsidePlanningHa)
if (abs(
  linked_inside_from_fragments - linked_inside_from_patch_audit
) > 1e-6) {
  stop(
    "Linked patch fragments do not reconcile to the inside/outside audit.",
    call. = FALSE
  )
}

readr::write_csv(
  patch_table,
  file.path(problem_dir, "existing_patch_dictionary_full_national.csv")
)
readr::write_csv(
  patch_area_by_subarea,
  file.path(problem_dir, "existing_patch_area_by_subarea.csv")
)
readr::write_csv(
  patch_area_crosscheck,
  file.path(problem_dir, "existing_patch_inside_outside_audit.csv")
)
readr::write_csv(
  national_current_by_subarea,
  file.path(problem_dir, "national_patches_vs_landuse_budget_audit.csv")
)
readr::write_csv(
  patch_ledger_scope_audit,
  file.path(problem_dir, "patch_ledger_scope_audit.csv")
)
readr::write_csv(
  cell_patch_links,
  file.path(problem_dir, "candidate_to_existing_patch_links.csv")
)

print(knitr::kable(national_current_by_subarea, digits = 3))
print(knitr::kable(patch_ledger_scope_audit, digits = 3))
knitr::kable(tibble(
  retained_existing_patches = nrow(patch_table),
  full_national_area_of_retained_patches_ha =
    sum(patch_table$ExistingPatchAreaHa),
  area_of_retained_patches_inside_planning_ha =
    sum(patch_area_crosscheck$ExistingPatchAreaInsidePlanningHa),
  area_of_retained_patches_outside_planning_ha =
    sum(patch_area_crosscheck$ExistingPatchAreaOutsidePlanningHa),
  linked_full_patches_already_at_least_5000_ha = sum(
    patch_table$ExistingPatchAreaHa >= global_network_minimum_ha
  ),
  candidate_patch_contacts = nrow(cell_patch_links),
  candidate_cells_touching_current_nature =
    n_distinct(cell_patch_links$coarse_cell_id)
), digits = 2)
```

# Potential graph and tight flow bounds

```{r potential-graph}
potential_vertices <- bind_rows(
  candidate_table |>
    transmute(
      vertex = paste0("c_", coarse_cell_id),
      node_type = "cell",
      node_id = coarse_cell_id,
      FlowAreaHa = EligibleAreaHa
    ),
  patch_table |>
    transmute(
      vertex = paste0("p_", patch_id),
      node_type = "patch",
      node_id = patch_id,
      FlowAreaHa = ExistingPatchAreaHa
    )
)

potential_edges <- bind_rows(
  candidate_edges |>
    transmute(from = paste0("c_", i), to = paste0("c_", j)),
  cell_patch_links |>
    transmute(
      from = paste0("c_", coarse_cell_id),
      to = paste0("p_", patch_id)
    )
) |>
  distinct()

potential_graph <- igraph::graph_from_data_frame(
  potential_edges,
  directed = FALSE,
  vertices = potential_vertices
)
potential_membership_raw <- igraph::components(
  potential_graph
)$membership
potential_membership <- tibble(
  vertex = names(potential_membership_raw),
  potential_component = as.integer(potential_membership_raw)
)

potential_vertices <- potential_vertices |>
  left_join(potential_membership, by = "vertex")

potential_component_summary <- potential_vertices |>
  group_by(potential_component) |>
  summarise(
    ComponentMaxAreaHa = sum(FlowAreaHa),
    candidate_cells = sum(node_type == "cell"),
    existing_patches = sum(node_type == "patch"),
    .groups = "drop"
  )

cell_component_bounds <- potential_vertices |>
  filter(node_type == "cell") |>
  select(coarse_cell_id = node_id, potential_component) |>
  left_join(potential_component_summary, by = "potential_component") |>
  select(coarse_cell_id, ComponentMaxAreaHa)

patch_component_bounds <- potential_vertices |>
  filter(node_type == "patch") |>
  select(patch_id = node_id, potential_component) |>
  left_join(potential_component_summary, by = "potential_component") |>
  select(patch_id, ComponentMaxAreaHa)

# Area coefficients for the Global and four Subarea certificates.
network_cell_area <- bind_rows(
  candidate_table |>
    transmute(
      network_name = "Global",
      coarse_cell_id,
      NetworkCellAreaHa = EligibleAreaHa
    ),
  restorable_area_by_subarea_cell |>
    transmute(
      network_name = paste0("Subarea", subarea_id),
      coarse_cell_id,
      NetworkCellAreaHa = RestorableAreaHaBySubarea
    )
) |>
  filter(NetworkCellAreaHa > 0)

network_patch_area <- bind_rows(
  patch_table |>
    transmute(
      network_name = "Global",
      patch_id,
      NetworkPatchAreaHa = ExistingPatchAreaHa
    ),
  patch_area_by_subarea |>
    transmute(
      network_name = paste0("Subarea", subarea_id),
      patch_id,
      NetworkPatchAreaHa = ExistingPatchAreaHaBySubarea
    )
) |>
  filter(NetworkPatchAreaHa > 0)

potential_component_network_area <- bind_rows(
  network_cell_area |>
    left_join(
      potential_vertices |>
        filter(node_type == "cell") |>
        transmute(
          coarse_cell_id = node_id,
          potential_component
        ),
      by = "coarse_cell_id"
    ) |>
    transmute(
      network_name,
      potential_component,
      CertificateAreaHa = NetworkCellAreaHa
    ),
  network_patch_area |>
    left_join(
      potential_vertices |>
        filter(node_type == "patch") |>
        transmute(patch_id = node_id, potential_component),
      by = "patch_id"
    ) |>
    transmute(
      network_name,
      potential_component,
      CertificateAreaHa = NetworkPatchAreaHa
    )
) |>
  group_by(network_name, potential_component) |>
  summarise(
    PotentialCertificateAreaHa = sum(CertificateAreaHa),
    .groups = "drop"
  ) |>
  left_join(
    potential_component_summary,
    by = "potential_component"
  ) |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    CanAnchor = candidate_cells > 0 & existing_patches > 0,
    PassesPotentialArea =
      PotentialCertificateAreaHa >= RequiredAreaHa - 1e-8,
    FeasiblePotentialComponent = CanAnchor & PassesPotentialArea
  )

potential_network_audit <- tibble(
  network_name = required_networks,
  RequiredAreaHa = unname(network_minimums[required_networks])
) |>
  left_join(
    potential_component_network_area |>
      group_by(network_name) |>
      summarise(
        RequiredAreaHa = first(RequiredAreaHa),
        feasible_potential_components =
          sum(FeasiblePotentialComponent),
        largest_potential_certificate_ha = max(
          PotentialCertificateAreaHa[CanAnchor],
          na.rm = TRUE
        ),
        .groups = "drop"
      ),
    by = c("network_name", "RequiredAreaHa")
  )

if (
  anyNA(potential_network_audit$feasible_potential_components) ||
  any(potential_network_audit$feasible_potential_components < 1L)
) {
  stop(
    "At least one requested network is impossible in the potential graph. ",
    "Inspect potential_network_feasibility_audit.csv.",
    call. = FALSE
  )
}

# Exact network-specific reduction: retain only potential components whose
# counted area can satisfy the corresponding certificate and which contain at
# least one candidate cell and one existing patch.
feasible_potential_components <- potential_component_network_area |>
  filter(FeasiblePotentialComponent) |>
  distinct(network_name, potential_component)

cell_potential_membership <- potential_vertices |>
  filter(node_type == "cell") |>
  transmute(
    coarse_cell_id = as.integer(node_id),
    potential_component
  )
patch_potential_membership <- potential_vertices |>
  filter(node_type == "patch") |>
  transmute(
    patch_id = as.integer(node_id),
    potential_component
  )

eligible_network_cells <- feasible_potential_components |>
  inner_join(cell_potential_membership, by = "potential_component") |>
  select(network_name, coarse_cell_id, potential_component) |>
  distinct() |>
  arrange(network_name, coarse_cell_id)

eligible_network_patches <- feasible_potential_components |>
  inner_join(patch_potential_membership, by = "potential_component") |>
  select(network_name, patch_id, potential_component) |>
  distinct() |>
  arrange(network_name, patch_id)

eligible_i <- eligible_network_cells |>
  transmute(network_name, i = coarse_cell_id)
eligible_j <- eligible_network_cells |>
  transmute(network_name, j = coarse_cell_id)

eligible_network_edges <- candidate_edges |>
  inner_join(eligible_i, by = "i") |>
  inner_join(eligible_j, by = c("network_name", "j")) |>
  select(network_name, i, j) |>
  distinct()

eligible_network_arcs <- eligible_network_edges |>
  bind_rows(
    eligible_network_edges |>
      transmute(network_name, i = j, j = i)
  ) |>
  distinct() |>
  arrange(network_name, i, j)

eligible_network_cell_patch_edges <- cell_patch_links |>
  inner_join(
    eligible_network_cells |>
      select(network_name, coarse_cell_id),
    by = "coarse_cell_id"
  ) |>
  inner_join(
    eligible_network_patches |>
      select(network_name, patch_id),
    by = c("network_name", "patch_id")
  ) |>
  select(network_name, coarse_cell_id, patch_id) |>
  distinct() |>
  arrange(network_name, coarse_cell_id, patch_id)

network_reduction_audit <- tibble(
  network_name = required_networks
) |>
  left_join(
    eligible_network_cells |>
      count(network_name, name = "eligible_cells"),
    by = "network_name"
  ) |>
  left_join(
    eligible_network_patches |>
      count(network_name, name = "eligible_patches"),
    by = "network_name"
  ) |>
  left_join(
    eligible_network_arcs |>
      count(network_name, name = "directed_cell_arcs"),
    by = "network_name"
  ) |>
  left_join(
    eligible_network_cell_patch_edges |>
      count(network_name, name = "cell_patch_edges"),
    by = "network_name"
  ) |>
  mutate(
    across(
      c(
        eligible_cells,
        eligible_patches,
        directed_cell_arcs,
        cell_patch_edges
      ),
      ~ coalesce(.x, 0L)
    )
  )

if (any(
  network_reduction_audit$eligible_cells < 1L |
    network_reduction_audit$eligible_patches < 1L |
    network_reduction_audit$cell_patch_edges < 1L
)) {
  stop(
    "Internal error while reducing network variables to feasible components.",
    call. = FALSE
  )
}

patch_connection_normalizer <- max(
  1,
  max(potential_component_summary$existing_patches) - 1
)

readr::write_csv(
  potential_component_summary,
  file.path(problem_dir, "potential_graph_components.csv")
)
readr::write_csv(
  potential_component_network_area,
  file.path(problem_dir, "potential_component_network_area.csv")
)
readr::write_csv(
  potential_network_audit,
  file.path(problem_dir, "potential_network_feasibility_audit.csv")
)
readr::write_csv(
  feasible_potential_components,
  file.path(problem_dir, "feasible_potential_components.csv")
)
readr::write_csv(
  network_reduction_audit,
  file.path(problem_dir, "network_variable_reduction_audit.csv")
)

print(knitr::kable(potential_network_audit, digits = 2))
knitr::kable(network_reduction_audit, digits = 0)
```

# Feasible actions and warm-up label repair

The selected **footprint** from notebook 07 is preserved in the MIP start.
Only a destination label is repaired when a production-forest-dominated cell
was assigned an open habitat in the simplified warm-up.

```{r model-actions}
allowed_actions <- tidyr::crossing(
  landuse = nature_landuses,
  candidate_table |>
    select(coarse_cell_id, IsProductionForestCell)
) |>
  filter(
    IsProductionForestCell == 0 |
      landuse %in% forest_landuses
  ) |>
  select(landuse, coarse_cell_id) |>
  arrange(landuse, coarse_cell_id)

objective_table <- allowed_actions |>
  left_join(
    biodiversity_table,
    by = c("landuse", "coarse_cell_id")
  ) |>
  left_join(
    existing_nature_contact,
    by = c("landuse", "coarse_cell_id")
  )

stopifnot(
  n_distinct(allowed_actions$coarse_cell_id) ==
    length(candidate_cells),
  !anyDuplicated(allowed_actions),
  !anyNA(objective_table),
  all(is.finite(objective_table$BioDiversity)),
  all(is.finite(objective_table$ExistingNatureContact))
)

biodiversity_normalizer <- objective_table |>
  group_by(coarse_cell_id) |>
  summarise(best = max(BioDiversity), .groups = "drop") |>
  summarise(total = sum(best)) |>
  pull(total) |>
  max(1)

contiguity_normalizer <- max(1, nrow(candidate_edges))

existing_nature_contact_normalizer <- objective_table |>
  group_by(coarse_cell_id) |>
  summarise(
    best = max(ExistingNatureContact),
    .groups = "drop"
  ) |>
  summarise(total = sum(best)) |>
  pull(total) |>
  max(1)

allowed_i <- allowed_actions |>
  transmute(landuse, i = coarse_cell_id)
allowed_j <- allowed_actions |>
  transmute(landuse, j = coarse_cell_id)

same_habitat_edges <- tidyr::crossing(
  landuse = nature_landuses,
  candidate_edges
) |>
  semi_join(allowed_i, by = c("landuse", "i")) |>
  semi_join(allowed_j, by = c("landuse", "j")) |>
  arrange(landuse, i, j)

best_forest_action <- objective_table |>
  filter(landuse %in% forest_landuses) |>
  mutate(
    local_start_score =
      BioDiversity / biodiversity_normalizer +
      spatial_contiguity_bonus *
        existing_to_new_adjacency_weight *
        ExistingNatureContact /
          existing_nature_contact_normalizer
  ) |>
  group_by(coarse_cell_id) |>
  arrange(
    desc(local_start_score),
    desc(BioDiversity),
    landuse,
    .by_group = TRUE
  ) |>
  slice_head(n = 1L) |>
  ungroup() |>
  select(
    coarse_cell_id,
    repaired_forest_landuse = landuse
  )

warm_start_assignment <- warmup_selected |>
  rename(original_landuse = landuse) |>
  left_join(
    candidate_table |>
      select(coarse_cell_id, IsProductionForestCell),
    by = "coarse_cell_id"
  ) |>
  left_join(best_forest_action, by = "coarse_cell_id") |>
  mutate(
    forest_rule_repair =
      IsProductionForestCell == 1L &
      !original_landuse %in% forest_landuses,
    landuse = if_else(
      forest_rule_repair,
      repaired_forest_landuse,
      original_landuse
    )
  ) |>
  select(
    coarse_cell_id,
    original_landuse,
    landuse,
    IsProductionForestCell,
    forest_rule_repair
  ) |>
  arrange(coarse_cell_id)

invalid_start_actions <- warm_start_assignment |>
  anti_join(
    allowed_actions,
    by = c("landuse", "coarse_cell_id")
  )
if (nrow(invalid_start_actions)) {
  stop("Warm-start action repair failed.", call. = FALSE)
}

warm_start_subarea_reference_audit <- restorable_area_by_subarea_cell |>
  filter(
    coarse_cell_id %in% warm_start_assignment$coarse_cell_id
  ) |>
  group_by(subarea_id) |>
  summarise(
    WarmStartRestorationHa = sum(RestorableAreaHaBySubarea),
    .groups = "drop"
  ) |>
  right_join(subarea_budget, by = "subarea_id") |>
  mutate(
    WarmStartRestorationHa = coalesce(WarmStartRestorationHa, 0),
    WarmStartFinalNatureHa =
      CurrentNatureAreaHa + WarmStartRestorationHa,
    WarmStartFinalNaturePercent =
      100 * WarmStartFinalNatureHa / PolygonAreaHa,
    ReferenceThirtyPercentDifferenceHa =
      WarmStartFinalNatureHa - FinalNatureTargetHa
  ) |>
  arrange(subarea_id)

warm_start_global_area_audit <-
  warm_start_subarea_reference_audit |>
  summarise(
    WarmStartRestorationHa = sum(WarmStartRestorationHa)
  ) |>
  bind_cols(global_budget) |>
  mutate(
    WarmStartFinalNatureHa =
      GlobalCurrentNatureAreaHa + WarmStartRestorationHa,
    WarmStartFinalNaturePercent =
      100 * WarmStartFinalNatureHa / GlobalPolygonAreaHa,
    ExcessAboveThirtyPercentHa =
      WarmStartFinalNatureHa - GlobalFinalNatureTargetHa,
    PassesGlobalThirtyPercentMinimum =
      WarmStartFinalNatureHa >= GlobalFinalNatureTargetHa - 1e-6,
    PassesGlobalThirtyPercentMaximum =
      WarmStartFinalNatureHa <= GlobalMaximumFinalNatureHa + 1e-6,
    PassesGlobalThirtyPercentBudget =
      PassesGlobalThirtyPercentMinimum &
        PassesGlobalThirtyPercentMaximum
  )

if (!warm_start_global_area_audit$PassesGlobalThirtyPercentBudget[[1]]) {
  stop(
    paste0(
      "The upstream footprint does not satisfy the whole-polygon 30% ",
      "budget and its configured upper tolerance."
    ),
    call. = FALSE
  )
}

readr::write_csv(
  warm_start_assignment,
  file.path(problem_dir, "warm_start_assignment_and_forest_repairs.csv")
)
readr::write_csv(
  warm_start_global_area_audit,
  file.path(problem_dir, "warm_start_30pct_audit.csv")
)
readr::write_csv(
  warm_start_global_area_audit,
  file.path(problem_dir, "warm_start_global_30pct_audit.csv")
)
readr::write_csv(
  warm_start_subarea_reference_audit,
  file.path(problem_dir, "warm_start_subarea_reference_audit.csv")
)

knitr::kable(tibble(
  warm_start_selected_cells = nrow(warm_start_assignment),
  production_forest_dominated_selected_cells = sum(
    warm_start_assignment$IsProductionForestCell == 1L
  ),
  destination_labels_repaired = sum(
    warm_start_assignment$forest_rule_repair
  ),
  warm_start_global_nature_percent =
    warm_start_global_area_audit$WarmStartFinalNaturePercent,
  passes_global_30pct_budget =
    warm_start_global_area_audit$PassesGlobalThirtyPercentBudget
))
```

# Find feasible certificate components in the warm-up solution

```{r warm-start-components}
warm_start_ids <- warm_start_assignment$coarse_cell_id
warm_start_patch_ids <- cell_patch_links |>
  filter(coarse_cell_id %in% warm_start_ids) |>
  distinct(patch_id) |>
  pull(patch_id) |>
  sort()

warm_start_vertices <- bind_rows(
  candidate_table |>
    filter(coarse_cell_id %in% warm_start_ids) |>
    transmute(
      vertex = paste0("c_", coarse_cell_id),
      node_type = "cell",
      node_id = coarse_cell_id,
      FlowAreaHa = EligibleAreaHa
    ),
  patch_table |>
    filter(patch_id %in% warm_start_patch_ids) |>
    transmute(
      vertex = paste0("p_", patch_id),
      node_type = "patch",
      node_id = patch_id,
      FlowAreaHa = ExistingPatchAreaHa
    )
)

warm_start_edges <- bind_rows(
  candidate_edges |>
    filter(i %in% warm_start_ids, j %in% warm_start_ids) |>
    transmute(from = paste0("c_", i), to = paste0("c_", j)),
  cell_patch_links |>
    filter(
      coarse_cell_id %in% warm_start_ids,
      patch_id %in% warm_start_patch_ids
    ) |>
    transmute(
      from = paste0("c_", coarse_cell_id),
      to = paste0("p_", patch_id)
    )
) |>
  distinct()

warm_start_graph <- igraph::graph_from_data_frame(
  warm_start_edges,
  directed = FALSE,
  vertices = warm_start_vertices
)
warm_membership_raw <- igraph::components(warm_start_graph)$membership
warm_membership <- tibble(
  vertex = names(warm_membership_raw),
  warm_component = as.integer(warm_membership_raw)
)
warm_start_vertices <- warm_start_vertices |>
  left_join(warm_membership, by = "vertex")

warm_component_summary <- warm_start_vertices |>
  group_by(warm_component) |>
  summarise(
    selected_cells = sum(node_type == "cell"),
    existing_patches = sum(node_type == "patch"),
    FlowAreaHa = sum(FlowAreaHa),
    .groups = "drop"
  )

warm_cell_membership <- warm_start_vertices |>
  filter(node_type == "cell") |>
  transmute(
    coarse_cell_id = node_id,
    warm_component
  )
warm_patch_membership <- warm_start_vertices |>
  filter(node_type == "patch") |>
  transmute(patch_id = node_id, warm_component)

warm_component_network_area <- bind_rows(
  network_cell_area |>
    filter(coarse_cell_id %in% warm_start_ids) |>
    left_join(warm_cell_membership, by = "coarse_cell_id") |>
    transmute(
      network_name,
      warm_component,
      CertificateAreaHa = NetworkCellAreaHa
    ),
  network_patch_area |>
    filter(patch_id %in% warm_start_patch_ids) |>
    left_join(warm_patch_membership, by = "patch_id") |>
    transmute(
      network_name,
      warm_component,
      CertificateAreaHa = NetworkPatchAreaHa
    )
) |>
  group_by(network_name, warm_component) |>
  summarise(
    CertifiedAreaHa = sum(CertificateAreaHa),
    .groups = "drop"
  ) |>
  left_join(warm_component_summary, by = "warm_component") |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    HasExistingRoot = existing_patches > 0,
    HasNewRestoration = selected_cells > 0,
    PassesArea = CertifiedAreaHa >= RequiredAreaHa - 1e-8,
    FeasibleWarmCertificate =
      HasExistingRoot & HasNewRestoration & PassesArea
  )

# Preserve the diagnostic, but do not stop when the warm footprint is not yet
# feasible. That is precisely the condition notebook 08a is designed to repair.
readr::write_csv(
  warm_component_network_area,
  file.path(problem_dir, "warm_start_component_network_audit.csv")
)

warm_start_certificate_availability <- tibble(
  network_name = required_networks
) |>
  left_join(
    warm_component_network_area |>
      group_by(network_name) |>
      summarise(
        feasible_warm_components = sum(FeasibleWarmCertificate),
        largest_anchored_warm_certificate_ha = max(c(
          0,
          CertifiedAreaHa[HasExistingRoot & HasNewRestoration]
        )),
        can_seed_complete_start = any(FeasibleWarmCertificate),
        .groups = "drop"
      ),
    by = "network_name"
  ) |>
  mutate(
    feasible_warm_components = coalesce(
      feasible_warm_components,
      0L
    ),
    largest_anchored_warm_certificate_ha = coalesce(
      largest_anchored_warm_certificate_ha,
      0
    ),
    can_seed_complete_start = coalesce(
      can_seed_complete_start,
      FALSE
    ),
    RequiredAreaHa = unname(network_minimums[network_name]),
    shortfall_ha = pmax(
      0,
      RequiredAreaHa - largest_anchored_warm_certificate_ha
    )
  )

complete_feasible_mip_start_available <- all(
  warm_start_certificate_availability$can_seed_complete_start
)

effective_warm_start_strategy <-
  "complete_slack_start_then_lexicographic_repair"

warm_start_strategy_audit <- warm_start_certificate_availability |>
  mutate(
    configured_mode = "08a_feasible_network_repair",
    effective_strategy = effective_warm_start_strategy,
    potential_graph_passed = TRUE
  )

readr::write_csv(
  warm_start_strategy_audit,
  file.path(problem_dir, "warm_start_strategy_audit.csv")
)

chosen_certificate_components <- if (
  complete_feasible_mip_start_available
) {
  map_dfr(
    required_networks,
    function(k) {
      warm_component_network_area |>
        filter(
          network_name == k,
          FeasibleWarmCertificate
        ) |>
        arrange(desc(CertifiedAreaHa), warm_component) |>
        slice_head(n = 1L)
    }
  )
} else {
  warm_component_network_area |>
    slice(0)
}

readr::write_csv(
  chosen_certificate_components,
  file.path(problem_dir, "warm_start_chosen_certificates.csv")
)

print(knitr::kable(warm_start_strategy_audit, digits = 2))

if (complete_feasible_mip_start_available) {
  knitr::kable(
    chosen_certificate_components |>
      select(
        network_name,
        warm_component,
        RequiredAreaHa,
        CertifiedAreaHa,
        selected_cells,
        existing_patches
      ),
    digits = 2
  )
} else {
  cat(
    "Notebook 07 is not yet a complete feasible network start. ",
    "Notebook 08a will start from its valid global-30% footprint with explicit ",
    "certificate shortfalls, repair all five networks, and only then ",
    "minimize footprint changes and optimize ecological quality.\n"
  )
}
```

# Construct complete flow starts for all five certificates

When all five certificates are already present, each chosen component is
converted to a deterministic spanning tree directed toward its largest
existing-nature patch. Subtree areas then give a complete feasible start for
the continuous flow variables as well as all binary variables. In automatic
hint mode this chunk is skipped because Gurobi constructs the network and flow
variables while repairing the warm-up footprint.

```{r build-complete-mip-start, eval=FALSE}
build_certificate_start <- function(network_name, component_id) {
  component_nodes <- warm_start_vertices |>
    filter(warm_component == component_id)

  root_row <- component_nodes |>
    filter(node_type == "patch") |>
    arrange(desc(FlowAreaHa), node_id) |>
    slice_head(n = 1L)
  if (nrow(root_row) != 1L) {
    stop("Certificate has no existing-nature root.", call. = FALSE)
  }

  root_vertex <- root_row$vertex[[1]]
  component_graph <- igraph::induced_subgraph(
    warm_start_graph,
    vids = component_nodes$vertex
  )

  distances_to_root <- as.numeric(igraph::distances(
    component_graph,
    v = root_vertex,
    to = igraph::V(component_graph),
    mode = "all"
  )[1, ])
  names(distances_to_root) <- igraph::V(component_graph)$name
  if (any(!is.finite(distances_to_root))) {
    stop("Internal error: certificate component is disconnected.")
  }

  all_vertices <- component_nodes$vertex
  parent <- stats::setNames(
    rep(NA_character_, length(all_vertices)),
    all_vertices
  )

  for (vertex in setdiff(all_vertices, root_vertex)) {
    neighbors <- igraph::as_ids(igraph::neighbors(
      component_graph,
      vertex,
      mode = "all"
    ))
    possible_parents <- neighbors[
      distances_to_root[neighbors] ==
        distances_to_root[[vertex]] - 1
    ]
    if (!length(possible_parents)) {
      stop("Could not assign a flow-tree parent to ", vertex)
    }
    parent[[vertex]] <- sort(possible_parents)[[1]]
  }

  subtree_area <- stats::setNames(
    component_nodes$FlowAreaHa[match(
      all_vertices,
      component_nodes$vertex
    )],
    all_vertices
  )
  flow_records <- list()
  flow_index <- 0L
  descending_vertices <- all_vertices[
    order(distances_to_root[all_vertices], decreasing = TRUE)
  ]

  for (vertex in descending_vertices) {
    if (vertex == root_vertex) next
    parent_vertex <- parent[[vertex]]
    flow_index <- flow_index + 1L
    flow_records[[flow_index]] <- tibble(
      network_name = network_name,
      child = vertex,
      parent = parent_vertex,
      flow_ha = subtree_area[[vertex]]
    )
    subtree_area[[parent_vertex]] <-
      subtree_area[[parent_vertex]] + subtree_area[[vertex]]
  }

  flow_table <- bind_rows(flow_records) |>
    mutate(
      child_type = str_sub(child, 1, 1),
      parent_type = str_sub(parent, 1, 1),
      child_id = as.integer(str_remove(child, "^[cp]_")),
      parent_id = as.integer(str_remove(parent, "^[cp]_"))
    )

  if (any(
    flow_table$child_type == "p" &
      flow_table$parent_type == "p"
  )) {
    stop("Unexpected patch-to-patch edge in a flow tree.")
  }

  list(
    network_name = network_name,
    component_id = component_id,
    cells = component_nodes |>
      filter(node_type == "cell") |>
      pull(node_id) |>
      as.integer() |>
      sort(),
    patches = component_nodes |>
      filter(node_type == "patch") |>
      pull(node_id) |>
      as.integer() |>
      sort(),
    root_patch = as.integer(root_row$node_id[[1]]),
    collected_area_ha = subtree_area[[root_vertex]],
    flows = flow_table
  )
}

certificate_starts <- map(
  seq_len(nrow(chosen_certificate_components)),
  function(index) {
    build_certificate_start(
      chosen_certificate_components$network_name[[index]],
      chosen_certificate_components$warm_component[[index]]
    )
  }
)
names(certificate_starts) <- map_chr(
  certificate_starts,
  "network_name"
)

start_certificate_cell_members <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    coarse_cell_id = x$cells
  )
)
start_certificate_patch_members <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    patch_id = x$patches
  )
)
start_certificate_roots <- map_dfr(
  certificate_starts,
  function(x) tibble(
    network_name = x$network_name,
    patch_id = x$root_patch,
    collected_area_ha = x$collected_area_ha
  )
)
start_certificate_flows <- map_dfr(certificate_starts, "flows")

start_flow_structure_audit <- map_dfr(
  certificate_starts,
  function(x) {
    expected_collected <-
      sum(candidate_table$EligibleAreaHa[
        match(x$cells, candidate_table$coarse_cell_id)
      ]) +
      sum(patch_table$ExistingPatchAreaHa[
        match(x$patches, patch_table$patch_id)
      ])

    tibble(
      network_name = x$network_name,
      member_nodes = length(x$cells) + length(x$patches),
      spanning_tree_arcs = nrow(x$flows),
      expected_collected_area_ha = expected_collected,
      start_collected_area_ha = x$collected_area_ha,
      tree_has_n_minus_one_arcs =
        nrow(x$flows) == length(x$cells) + length(x$patches) - 1L,
      collected_area_reconciles =
        abs(x$collected_area_ha - expected_collected) <= 1e-6,
      all_tree_flows_positive = all(x$flows$flow_ha > 0)
    )
  }
)

if (!all(
  start_flow_structure_audit$tree_has_n_minus_one_arcs &
    start_flow_structure_audit$collected_area_reconciles &
    start_flow_structure_audit$all_tree_flows_positive
)) {
  stop("Internal error in certificate flow starts.", call. = FALSE)
}

# Independent membership and area checks before any AMPL files are written.
start_certificate_area_check <- bind_rows(
  start_certificate_cell_members |>
    left_join(
      network_cell_area,
      by = c("network_name", "coarse_cell_id")
    ) |>
    transmute(
      network_name,
      area_ha = coalesce(NetworkCellAreaHa, 0)
    ),
  start_certificate_patch_members |>
    left_join(
      network_patch_area,
      by = c("network_name", "patch_id")
    ) |>
    transmute(
      network_name,
      area_ha = coalesce(NetworkPatchAreaHa, 0)
    )
) |>
  group_by(network_name) |>
  summarise(CertifiedAreaHa = sum(area_ha), .groups = "drop") |>
  mutate(
    RequiredAreaHa = unname(network_minimums[network_name]),
    Passes = CertifiedAreaHa >= RequiredAreaHa - 1e-6
  )

if (
  nrow(start_certificate_area_check) != length(required_networks) ||
  !all(start_certificate_area_check$Passes)
) {
  stop("Internal error in warm-start certificate areas.", call. = FALSE)
}

if (
  nrow(start_certificate_roots) != length(required_networks) ||
  anyDuplicated(start_certificate_roots$network_name)
) {
  stop("Each certificate must have exactly one root.", call. = FALSE)
}

readr::write_csv(
  start_certificate_area_check,
  file.path(problem_dir, "warm_start_certificate_area_check.csv")
)
readr::write_csv(
  start_certificate_roots,
  file.path(problem_dir, "warm_start_certificate_roots.csv")
)
readr::write_csv(
  start_flow_structure_audit,
  file.path(problem_dir, "warm_start_flow_structure_audit.csv")
)

print(knitr::kable(start_certificate_area_check, digits = 2))
knitr::kable(start_flow_structure_audit, digits = 6)
```

# Export AMPL data, complete MIP start, and run file

```{r export-ampl}
dat_path <- file.path(problem_dir, paste0(problem_name, ".dat"))
mip_start_path <- file.path(
  problem_dir,
  paste0(problem_name, "_mip_start.run")
)
run_path <- file.path(problem_dir, paste0(problem_name, ".run"))

solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_selected_cells.tsv")
)
selected_subarea_path <- file.path(
  problem_dir,
  paste0(problem_name, "_selected_cell_subareas.tsv")
)
subarea_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_subarea_summary.tsv")
)
network_cell_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_cells.tsv")
)
network_patch_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_patches.tsv")
)
network_root_solution_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_roots.tsv")
)
network_solution_summary_path <- file.path(
  problem_dir,
  paste0(problem_name, "_network_summary.tsv")
)
summary_path <- file.path(
  problem_dir,
  paste0(problem_name, "_solution_summary.tsv")
)
repair_stage_summary_path <- file.path(
  problem_dir,
  paste0(problem_name, "_repair_stage_summary.tsv")
)
repair_shortfall_path <- file.path(
  problem_dir,
  paste0(problem_name, "_stage1_network_shortfall.tsv")
)
repair_completion_path <- file.path(
  problem_dir,
  paste0(problem_name, "_repair_completion_status.txt")
)

dat_temporary_path <- paste0(dat_path, ".building")
writeLines(
  c(
    "# Generated by 08a_gudenaa_new_area_200m_feasible_network_repair.Rmd",
    paste("# Polygon area (ha):", ampl_number(sum(polygon_areas_ha))),
    paste(
      "# Global network minimum (ha):",
      ampl_number(global_network_minimum_ha)
    ),
    paste(
      "# Per-subarea network minimum (ha):",
      ampl_number(subarea_network_minimum_ha)
    ),
    paste(
      "# Global existing-patch area:",
      "full national connected patch area (outside included once)"
    ),
    paste(
      "# Subarea existing-patch area:",
      "only geometry physically inside that subarea"
    ),
    "# All restoration coefficients are restricted to the planning polygon.",
    "# Certificate variables are restricted to target-feasible potential components.",
    "# WarmSelected records the successful notebook-07 footprint.",
    "# Carbon, MustIncludeCells, and forest/wet quotas are absent.",
    ""
  ),
  dat_temporary_path
)

write_ampl_set(dat_temporary_path, "Cells", candidate_cells)
write_ampl_set(dat_temporary_path, "Landuses", nature_landuses)
write_ampl_set(dat_temporary_path, "ForestLanduses", forest_landuses)
write_ampl_set(
  dat_temporary_path,
  "Subareas",
  subarea_budget$subarea_id
)
write_ampl_set(
  dat_temporary_path,
  "RequiredNetworks",
  required_networks
)
write_ampl_set(dat_temporary_path, "GlobalNetworks", "Global")
write_ampl_pair_set(
  dat_temporary_path,
  "AllowedActions",
  allowed_actions$landuse,
  allowed_actions$coarse_cell_id
)
write_ampl_pair_set(
  dat_temporary_path,
  "E",
  candidate_edges$i,
  candidate_edges$j
)
write_ampl_triple_set(
  dat_temporary_path,
  "SameHabitatEdges",
  same_habitat_edges$landuse,
  same_habitat_edges$i,
  same_habitat_edges$j
)
write_ampl_set(
  dat_temporary_path,
  "ExistingPatches",
  existing_patches
)
write_ampl_pair_set(
  dat_temporary_path,
  "CellPatchEdges",
  cell_patch_links$coarse_cell_id,
  cell_patch_links$patch_id
)
write_ampl_pair_set(
  dat_temporary_path,
  "EligibleNetworkCells",
  eligible_network_cells$network_name,
  eligible_network_cells$coarse_cell_id
)
write_ampl_pair_set(
  dat_temporary_path,
  "EligibleNetworkPatches",
  eligible_network_patches$network_name,
  eligible_network_patches$patch_id
)
write_ampl_triple_set(
  dat_temporary_path,
  "NetworkCellArcs",
  eligible_network_arcs$network_name,
  eligible_network_arcs$i,
  eligible_network_arcs$j
)
write_ampl_triple_set(
  dat_temporary_path,
  "NetworkCellPatchEdges",
  eligible_network_cell_patch_edges$network_name,
  eligible_network_cell_patch_edges$coarse_cell_id,
  eligible_network_cell_patch_edges$patch_id
)

write_ampl_matrix(
  dat_temporary_path,
  "BioDiversity",
  objective_table$landuse,
  objective_table$coarse_cell_id,
  objective_table$BioDiversity
)
write_ampl_matrix(
  dat_temporary_path,
  "ExistingNatureContact",
  objective_table$landuse,
  objective_table$coarse_cell_id,
  objective_table$ExistingNatureContact
)
write_ampl_vector(
  dat_temporary_path,
  "RestorableAreaHa",
  candidate_table$coarse_cell_id,
  candidate_table$EligibleAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "RestorableAreaHaBySubarea",
  restorable_area_by_subarea_cell$subarea_id,
  restorable_area_by_subarea_cell$coarse_cell_id,
  restorable_area_by_subarea_cell$RestorableAreaHaBySubarea
)
write_ampl_vector(
  dat_temporary_path,
  "SubareaAreaHa",
  subarea_budget$subarea_id,
  subarea_budget$PolygonAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "CurrentNatureAreaHa",
  subarea_budget$subarea_id,
  subarea_budget$CurrentNatureAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "FinalNatureTargetHa",
  subarea_budget$subarea_id,
  subarea_budget$FinalNatureTargetHa
)
write_ampl_vector(
  dat_temporary_path,
  "RestorationRequirementHa",
  subarea_budget$subarea_id,
  subarea_budget$RestorationRequirementHa
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalPolygonAreaHa",
  global_budget$GlobalPolygonAreaHa
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalCurrentNatureAreaHa",
  global_budget$GlobalCurrentNatureAreaHa
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalNatureTargetFraction",
  global_budget$GlobalNatureTargetFraction
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalFinalNatureTargetHa",
  global_budget$GlobalFinalNatureTargetHa
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalRestorationRequirementHa",
  global_budget$GlobalRestorationRequirementHa
)
write_ampl_scalar(
  dat_temporary_path,
  "GlobalRestorationToleranceHa",
  global_budget$GlobalRestorationToleranceHa
)

write_ampl_vector(
  dat_temporary_path,
  "ExistingPatchAreaHa",
  patch_table$patch_id,
  patch_table$ExistingPatchAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "NetworkCellAreaHa",
  network_cell_area$network_name,
  network_cell_area$coarse_cell_id,
  network_cell_area$NetworkCellAreaHa
)
write_ampl_matrix(
  dat_temporary_path,
  "NetworkPatchAreaHa",
  network_patch_area$network_name,
  network_patch_area$patch_id,
  network_patch_area$NetworkPatchAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "NetworkMinimumHa",
  names(network_minimums),
  unname(network_minimums)
)
write_ampl_vector(
  dat_temporary_path,
  "ComponentMaxAreaCellHa",
  cell_component_bounds$coarse_cell_id,
  cell_component_bounds$ComponentMaxAreaHa
)
write_ampl_vector(
  dat_temporary_path,
  "ComponentMaxAreaPatchHa",
  patch_component_bounds$patch_id,
  patch_component_bounds$ComponentMaxAreaHa
)

production_forest_cells <- candidate_table$coarse_cell_id[
  candidate_table$IsProductionForestCell == 1L
]
write_ampl_vector(
  dat_temporary_path,
  "IsProductionForestCell",
  production_forest_cells,
  rep(1, length(production_forest_cells))
)
write_ampl_vector(
  dat_temporary_path,
  "WarmSelected",
  warm_start_assignment$coarse_cell_id,
  rep(1, nrow(warm_start_assignment))
)

write_ampl_scalar(
  dat_temporary_path,
  "BiodiversityNormalizer",
  biodiversity_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "ContiguityNormalizer",
  contiguity_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "ExistingNatureContactNormalizer",
  existing_nature_contact_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "PatchConnectionNormalizer",
  patch_connection_normalizer
)
write_ampl_scalar(
  dat_temporary_path,
  "SpatialContiguityBonus",
  spatial_contiguity_bonus
)
write_ampl_scalar(
  dat_temporary_path,
  "NewToNewAdjacencyWeight",
  new_to_new_adjacency_weight
)
write_ampl_scalar(
  dat_temporary_path,
  "ExistingToNewAdjacencyWeight",
  existing_to_new_adjacency_weight
)
write_ampl_scalar(
  dat_temporary_path,
  "ConnectedExistingPatchBonus",
  connected_existing_patch_bonus
)
write_ampl_scalar(
  dat_temporary_path,
  "FootprintChangeLimitHa",
  sum(candidate_table$EligibleAreaHa)
)
write_ampl_scalar(
  dat_temporary_path,
  "RepairChangeToleranceHa",
  repair_change_tolerance_ha
)

dat_text <- readLines(dat_temporary_path, warn = FALSE)
forbidden_tokens <- grepl(
  "(^|[[:space:]])(NA|NaN|Inf|-Inf)([[:space:];]|$)",
  dat_text
)
stopifnot(!any(forbidden_tokens))
commit_generated_file(dat_temporary_path, dat_path)

# ---- Complete feasible shortfall start ----------------------------------
mip_start_temporary_path <- paste0(mip_start_path, ".building")
# The two legacy branches below are intentionally disabled and retained only
# to document how earlier complete/hint starts were constructed. The active
# 08a shortfall start is written immediately after them.
if (FALSE) {
writeLines(
  c(
    "# Complete feasible start built from the successful notebook-07 footprint.",
    "let {c in Cells} Selected[c] := 0;",
    "let {(l,c) in AllowedActions} LanduseDecision[l,c] := 0;",
    "let {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j] := 0;",
    "let {k in RequiredNetworks, c in Cells} NetworkCell[k,c] := 0;",
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "NetworkPatch[k,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "NetworkRoot[k,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (i,j) in NetworkArcs} ",
      "AreaFlowHa[k,i,j] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (c,p) in CellPatchEdges} ",
      "CellToPatchFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, (c,p) in CellPatchEdges} ",
      "PatchToCellFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {k in RequiredNetworks, p in ExistingPatches} ",
      "CollectedAreaHa[k,p] := 0;"
    )
  ),
  mip_start_temporary_path
)

append_start(
  mip_start_temporary_path,
  paste0(
    "let Selected[",
    warm_start_assignment$coarse_cell_id,
    "] := 1;"
  )
)
append_start(
  mip_start_temporary_path,
  warm_start_assignment |>
    transmute(line = paste0(
      "let LanduseDecision['",
      landuse,
      "',",
      coarse_cell_id,
      "] := 1;"
    )) |>
    pull(line)
)

start_labels_i <- warm_start_assignment |>
  transmute(landuse, i = coarse_cell_id)
start_labels_j <- warm_start_assignment |>
  transmute(landuse, j = coarse_cell_id)
start_contiguity <- same_habitat_edges |>
  semi_join(start_labels_i, by = c("landuse", "i")) |>
  semi_join(start_labels_j, by = c("landuse", "j"))
append_start(
  mip_start_temporary_path,
  paste0(
    "let Contiguity['",
    start_contiguity$landuse,
    "',",
    start_contiguity$i,
    ",",
    start_contiguity$j,
    "] := 1;"
  )
)

append_start(
  mip_start_temporary_path,
  start_certificate_cell_members |>
    transmute(line = paste0(
      "let NetworkCell['",
      network_name,
      "',",
      coarse_cell_id,
      "] := 1;"
    )) |>
    pull(line)
)
append_start(
  mip_start_temporary_path,
  start_certificate_patch_members |>
    transmute(line = paste0(
      "let NetworkPatch['",
      network_name,
      "',",
      patch_id,
      "] := 1;"
    )) |>
    pull(line)
)
append_start(
  mip_start_temporary_path,
  start_certificate_roots |>
    transmute(line = paste0(
      "let NetworkRoot['",
      network_name,
      "',",
      patch_id,
      "] := 1;"
    )) |>
    pull(line)
)

flow_start_lines <- start_certificate_flows |>
  mutate(line = case_when(
    child_type == "c" & parent_type == "c" ~ paste0(
      "let AreaFlowHa['",
      network_name,
      "',",
      child_id,
      ",",
      parent_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    child_type == "c" & parent_type == "p" ~ paste0(
      "let CellToPatchFlowHa['",
      network_name,
      "',",
      child_id,
      ",",
      parent_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    child_type == "p" & parent_type == "c" ~ paste0(
      "let PatchToCellFlowHa['",
      network_name,
      "',",
      parent_id,
      ",",
      child_id,
      "] := ",
      ampl_number(flow_ha),
      ";"
    ),
    TRUE ~ NA_character_
  ))
stopifnot(!anyNA(flow_start_lines$line))
append_start(mip_start_temporary_path, flow_start_lines$line)

append_start(
  mip_start_temporary_path,
  start_certificate_roots |>
    transmute(line = paste0(
      "let CollectedAreaHa['",
      network_name,
      "',",
      patch_id,
      "] := ",
      ampl_number(collected_area_ha),
      ";"
    )) |>
    pull(line)
)
commit_generated_file(mip_start_temporary_path, mip_start_path)
} else if (FALSE) {
  # The notebook-07 footprint satisfies the whole-polygon 30% budget but is
  # not connected enough to satisfy every hard certificate. Supply only its
  # restoration/action pattern. With mip:start=2, Gurobi interprets these
  # incoming binary values as hints rather than as a claimed feasible start.
  writeLines(
    c(
      "# Infeasible notebook-07 footprint supplied as Gurobi variable hints.",
      "# Network and flow variables are deliberately left for Gurobi to build.",
      "let {c in Cells} Selected[c] := 0;",
      "let {(l,c) in AllowedActions} LanduseDecision[l,c] := 0;",
      "let {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j] := 0;"
    ),
    mip_start_temporary_path
  )

  append_start(
    mip_start_temporary_path,
    paste0(
      "let Selected[",
      warm_start_assignment$coarse_cell_id,
      "] := 1;"
    )
  )
  append_start(
    mip_start_temporary_path,
    warm_start_assignment |>
      transmute(line = paste0(
        "let LanduseDecision['",
        landuse,
        "',",
        coarse_cell_id,
        "] := 1;"
      )) |>
      pull(line)
  )

  hint_labels_i <- warm_start_assignment |>
    transmute(landuse, i = coarse_cell_id)
  hint_labels_j <- warm_start_assignment |>
    transmute(landuse, j = coarse_cell_id)
  hint_contiguity <- same_habitat_edges |>
    semi_join(hint_labels_i, by = c("landuse", "i")) |>
    semi_join(hint_labels_j, by = c("landuse", "j"))
  append_start(
    mip_start_temporary_path,
    paste0(
      "let Contiguity['",
      hint_contiguity$landuse,
      "',",
      hint_contiguity$i,
      ",",
      hint_contiguity$j,
      "] := 1;"
    )
  )

  commit_generated_file(mip_start_temporary_path, mip_start_path)
}

# Write a genuinely complete and
# feasible stage-1 start for the repair model. No infeasible network is claimed:
# all certificates are initially inactive and their requirements are carried by
# explicit shortfall variables. The notebook-07 restoration and land-use
# decisions remain exactly represented.
writeLines(
  c(
    "# Complete feasible start for stage 1 of the 08a repair model.",
    "# Warm footprint is exact; certificate shortfalls are explicit.",
    "let {c in Cells} Selected[c] := 0;",
    "let {(l,c) in AllowedActions} LanduseDecision[l,c] := 0;",
    "let {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j] := 0;",
    "let {k in RequiredNetworks} CertificateBuilt[k] := 0;",
    paste0(
      "let {k in RequiredNetworks} NetworkShortfallHa[k] := ",
      "NetworkMinimumHa[k];"
    ),
    paste0(
      "let {(k,c) in EligibleNetworkCells} ",
      "NetworkCell[k,c] := 0;"
    ),
    paste0(
      "let {(k,p) in EligibleNetworkPatches} ",
      "NetworkPatch[k,p] := 0;"
    ),
    paste0(
      "let {(k,p) in EligibleNetworkPatches} ",
      "NetworkRoot[k,p] := 0;"
    ),
    paste0(
      "let {(k,i,j) in NetworkCellArcs} ",
      "AreaFlowHa[k,i,j] := 0;"
    ),
    paste0(
      "let {(k,c,p) in NetworkCellPatchEdges} ",
      "CellToPatchFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {(k,c,p) in NetworkCellPatchEdges} ",
      "PatchToCellFlowHa[k,c,p] := 0;"
    ),
    paste0(
      "let {(k,p) in EligibleNetworkPatches} ",
      "CollectedAreaHa[k,p] := 0;"
    )
  ),
  mip_start_temporary_path
)

append_start(
  mip_start_temporary_path,
  paste0(
    "let Selected[",
    warm_start_assignment$coarse_cell_id,
    "] := 1;"
  )
)
append_start(
  mip_start_temporary_path,
  warm_start_assignment |>
    transmute(line = paste0(
      "let LanduseDecision['",
      landuse,
      "',",
      coarse_cell_id,
      "] := 1;"
    )) |>
    pull(line)
)

repair_start_labels_i <- warm_start_assignment |>
  transmute(landuse, i = coarse_cell_id)
repair_start_labels_j <- warm_start_assignment |>
  transmute(landuse, j = coarse_cell_id)
repair_start_contiguity <- same_habitat_edges |>
  semi_join(repair_start_labels_i, by = c("landuse", "i")) |>
  semi_join(repair_start_labels_j, by = c("landuse", "j"))
append_start(
  mip_start_temporary_path,
  paste0(
    "let Contiguity['",
    repair_start_contiguity$landuse,
    "',",
    repair_start_contiguity$i,
    ",",
    repair_start_contiguity$j,
    "] := 1;"
  )
)

commit_generated_file(mip_start_temporary_path, mip_start_path)
```

```{r export-run-and-audits}
stage_solver_options <- function(
  time_limit_seconds,
  stage_label,
  mip_gap = repair_mip_gap
) {
  paste(
    paste0("lim:time=", as.integer(time_limit_seconds)),
    paste0("mip:gap=", ampl_number(mip_gap)),
    "mip:start=1",
    solver_options_base,
    paste0(
      "tech:logfile=",
      file.path(
        problem_dir,
        paste0(problem_name, "_", stage_label, "_gurobi.log")
      )
    ),
    collapse = " "
  )
}

stage1_solver_options <- stage_solver_options(
  repair_feasibility_time_limit_seconds,
  "stage1_feasibility",
  mip_gap = 0
)
stage2_solver_options <- stage_solver_options(
  repair_minimum_change_time_limit_seconds,
  "stage2_minimum_change"
)
stage3_solver_options <- stage_solver_options(
  repair_quality_time_limit_seconds,
  "stage3_quality"
)

run_lines <- c(
  "reset;",
  "",
  paste0('model "', ampl_path(model_path), '";'),
  paste0('data "', ampl_path(dat_path), '";'),
  paste0('include "', ampl_path(mip_start_path), '";'),
  "",
  paste("option solver", solver_name, ";"),
  paste0(
    'printf "INCOMPLETE\\n" > "',
    ampl_path(repair_completion_path), '";'
  ),
  "drop WarmFootprintChange;",
  "drop RepairQuality;",
  paste0(
    "option ", solver_name, "_options '",
    ampl_path(stage1_solver_options), "';"
  ),
  "solve;",
  "display solve_result_num, solve_result;",
  "",
  paste0(
    'printf "stage\\tsolve_result\\tsolve_result_num',
    '\\tobjective_value\\ttotal_shortfall_ha',
    '\\tfootprint_change_ha\\tfootprint_limit_ha\\n" > "',
    ampl_path(repair_stage_summary_path), '";'
  ),
  paste0(
    'printf "1_network_feasibility\\t%s\\t%d\\t%.10f',
    '\\t%.8f\\t.\\t%.8f\\n", ',
    'solve_result, solve_result_num, RelativeNetworkShortfall, ',
    'sum {k in RequiredNetworks} NetworkShortfallHa[k], ',
    'FootprintChangeLimitHa >> "',
    ampl_path(repair_stage_summary_path), '";'
  ),
  paste0(
    'printf "network_name\\tminimum_ha\\tshortfall_ha',
    '\\trelative_shortfall\\tbuilt\\n" > "',
    ampl_path(repair_shortfall_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks} ',
    '"%s\\t%.8f\\t%.8f\\t%.10f\\t%d\\n", ',
    'k, NetworkMinimumHa[k], NetworkShortfallHa[k], ',
    'NetworkShortfallHa[k] / NetworkMinimumHa[k], ',
    'CertificateBuilt[k] >> "',
    ampl_path(repair_shortfall_path), '";'
  ),
  "",
  paste0(
    "if sum {k in RequiredNetworks} NetworkShortfallHa[k] <= ",
    ampl_number(network_shortfall_tolerance_ha),
    " then {"
  ),
  "fix {k in RequiredNetworks} NetworkShortfallHa[k] := 0;",
  "drop RelativeNetworkShortfall;",
  "restore WarmFootprintChange;",
  paste0(
    "option ", solver_name, "_options '",
    ampl_path(stage2_solver_options), "';"
  ),
  "solve;",
  "display solve_result_num, solve_result;",
  paste0(
    'printf "2_minimum_footprint_change\\t%s\\t%d\\t%.10f',
    '\\t%.8f\\t%.8f\\t%.8f\\n", ',
    'solve_result, solve_result_num, WarmFootprintChange, ',
    'sum {k in RequiredNetworks} NetworkShortfallHa[k], ',
    'WarmFootprintChange, FootprintChangeLimitHa >> "',
    ampl_path(repair_stage_summary_path), '";'
  ),
  paste0(
    "let FootprintChangeLimitHa := WarmFootprintChange + ",
    "RepairChangeToleranceHa;"
  ),
  "drop WarmFootprintChange;",
  "restore RepairQuality;",
  paste0(
    "option ", solver_name, "_options '",
    ampl_path(stage3_solver_options), "';"
  ),
  "solve;",
  "display solve_result_num, solve_result;",
  paste0(
    'printf "3_ecological_contiguity_quality\\t%s\\t%d\\t%.10f',
    '\\t%.8f\\t%.8f\\t%.8f\\n", ',
    'solve_result, solve_result_num, RepairQuality, ',
    'sum {k in RequiredNetworks} NetworkShortfallHa[k], ',
    'sum {c in Cells} RestorableAreaHa[c] * ',
    '(WarmSelected[c] * (1 - Selected[c]) + ',
    '(1 - WarmSelected[c]) * Selected[c]), ',
    'FootprintChangeLimitHa >> "',
    ampl_path(repair_stage_summary_path), '";'
  ),
  "",
  paste0(
    'printf "coarse_cell_id\\tlanduse\\trestored_ha',
    '\\tbiodiversity\\texisting_nature_contact_edges\\n" > "',
    ampl_path(solution_path), '";'
  ),
  paste0(
    'printf {(l,c) in AllowedActions: ',
    'LanduseDecision[l,c] > 0.5} ',
    '"%d\\t%s\\t%.8f\\t%.8f\\t%.8f\\n", ',
    'c, l, RestorableAreaHa[c], BioDiversity[l,c], ',
    'ExistingNatureContact[l,c] >> "',
    ampl_path(solution_path), '";'
  ),
  "",
  paste0(
    'printf "subarea_id\\tcoarse_cell_id\\trestored_ha\\n" > "',
    ampl_path(selected_subarea_path), '";'
  ),
  paste0(
    'printf {s in Subareas, c in Cells: Selected[c] > 0.5 ',
    'and RestorableAreaHaBySubarea[s,c] > 0} ',
    '"%d\\t%d\\t%.8f\\n", s, c, ',
    'RestorableAreaHaBySubarea[s,c] >> "',
    ampl_path(selected_subarea_path), '";'
  ),
  "",
  paste0(
    'printf "subarea_id\\tpolygon_ha\\treference_30pct_target_ha',
    '\\tcurrent_ha\\treference_30pct_restoration_ha',
    '\\trestored_ha\\tfinal_nature_ha',
    '\\tfinal_nature_percent\\n" > "',
    ampl_path(subarea_solution_path), '";'
  ),
  paste0(
    'printf {s in Subareas} ',
    '"%d\\t%.8f\\t%.8f\\t%.8f\\t%.8f\\t%.8f\\t%.8f',
    '\\t%.8f\\n", ',
    's, SubareaAreaHa[s], FinalNatureTargetHa[s], ',
    'CurrentNatureAreaHa[s], RestorationRequirementHa[s], ',
    'sum {c in Cells} RestorableAreaHaBySubarea[s,c] * Selected[c], ',
    'CurrentNatureAreaHa[s] + sum {c in Cells} ',
    'RestorableAreaHaBySubarea[s,c] * Selected[c], ',
    '100 * (CurrentNatureAreaHa[s] + sum {c in Cells} ',
    'RestorableAreaHaBySubarea[s,c] * Selected[c]) / ',
    'SubareaAreaHa[s] >> "', ampl_path(subarea_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tcoarse_cell_id\\tcounted_area_ha',
    '\\ttotal_restorable_ha\\n" > "',
    ampl_path(network_cell_solution_path), '";'
  ),
  paste0(
    'printf {(k,c) in EligibleNetworkCells: ',
    'NetworkCell[k,c] > 0.5} ',
    '"%s\\t%d\\t%.8f\\t%.8f\\n", ',
    'k, c, NetworkCellAreaHa[k,c], RestorableAreaHa[c] >> "',
    ampl_path(network_cell_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tpatch_id\\tcounted_area_ha',
    '\\ttotal_patch_ha\\n" > "',
    ampl_path(network_patch_solution_path), '";'
  ),
  paste0(
    'printf {(k,p) in EligibleNetworkPatches: ',
    'NetworkPatch[k,p] > 0.5} ',
    '"%s\\t%d\\t%.8f\\t%.8f\\n", ',
    'k, p, NetworkPatchAreaHa[k,p], ExistingPatchAreaHa[p] >> "',
    ampl_path(network_patch_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tpatch_id\\tcollected_area_ha\\n" > "',
    ampl_path(network_root_solution_path), '";'
  ),
  paste0(
    'printf {(k,p) in EligibleNetworkPatches: ',
    'NetworkRoot[k,p] > 0.5} ',
    '"%s\\t%d\\t%.8f\\n", k, p, CollectedAreaHa[k,p] >> "',
    ampl_path(network_root_solution_path), '";'
  ),
  "",
  paste0(
    'printf "network_name\\tminimum_ha\\tcertified_area_ha',
    '\\tmember_cells\\tmember_patches\\troot_count',
    '\\tcollected_total_area_ha\\tshortfall_ha',
    '\\tcertificate_built\\n" > "',
    ampl_path(network_solution_summary_path), '";'
  ),
  paste0(
    'printf {k in RequiredNetworks} ',
    '"%s\\t%.8f\\t%.8f\\t%d\\t%d\\t%d\\t%.8f',
    '\\t%.8f\\t%d\\n", ',
    'k, NetworkMinimumHa[k], ',
    'sum {(q,c) in EligibleNetworkCells: q = k} ',
    'NetworkCellAreaHa[q,c] * NetworkCell[q,c] + ',
    'sum {(q,p) in EligibleNetworkPatches: q = k} ',
    'NetworkPatchAreaHa[q,p] * NetworkPatch[q,p], ',
    'sum {(q,c) in EligibleNetworkCells: q = k} NetworkCell[q,c], ',
    'sum {(q,p) in EligibleNetworkPatches: q = k} NetworkPatch[q,p], ',
    'sum {(q,p) in EligibleNetworkPatches: q = k} NetworkRoot[q,p], ',
    'sum {(q,p) in EligibleNetworkPatches: q = k} ',
    'CollectedAreaHa[q,p], NetworkShortfallHa[k], ',
    'CertificateBuilt[k] >> "',
    ampl_path(network_solution_summary_path), '";'
  ),
  "",
  paste0(
    'printf "item\\tvalue\\n" > "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "solve_result\\t%s\\n", solve_result >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "solve_result_num\\t%d\\n", solve_result_num >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "repair_status\\tsolved_feasible_network_repair\\n" >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "objective\\t%.10f\\n", RepairQuality >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "warm_footprint_change_ha\\t%.8f\\n", ',
    'sum {c in Cells} RestorableAreaHa[c] * ',
    '(WarmSelected[c] * (1 - Selected[c]) + ',
    '(1 - WarmSelected[c]) * Selected[c]) >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "network_shortfall_total_ha\\t%.8f\\n", ',
    'sum {k in RequiredNetworks} NetworkShortfallHa[k] >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "spatial_contiguity_bonus\\t%.8f\\n", ',
    'SpatialContiguityBonus >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_biodiversity\\t%.10f\\n", ',
    '(sum {(l,c) in AllowedActions} BioDiversity[l,c] * ',
    'LanduseDecision[l,c]) / BiodiversityNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_new_to_new_adjacency\\t%.10f\\n", ',
    '(sum {(l,i,j) in SameHabitatEdges} Contiguity[l,i,j]) / ',
    'ContiguityNormalizer >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_existing_to_new_adjacency\\t%.10f\\n", ',
    '(sum {(l,c) in AllowedActions} ExistingNatureContact[l,c] * ',
    'LanduseDecision[l,c]) / ExistingNatureContactNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_existing_patches_joined_beyond_root\\t%.0f\\n", ',
    'sum {(k,p) in EligibleNetworkPatches: k in GlobalNetworks} ',
    'NetworkPatch[k,p] - sum {k in GlobalNetworks} ',
    'CertificateBuilt[k] >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "normalized_existing_patch_connection\\t%.10f\\n", ',
    '(sum {(k,p) in EligibleNetworkPatches: k in GlobalNetworks} ',
    'NetworkPatch[k,p] - sum {k in GlobalNetworks} ',
    'CertificateBuilt[k]) / PatchConnectionNormalizer >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "selected_cells\\t%.0f\\n", ',
    'sum {c in Cells} Selected[c] >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "restored_ha\\t%.8f\\n", ',
    'sum {c in Cells} RestorableAreaHa[c] * Selected[c] >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_polygon_area_ha\\t%.8f\\n", ',
    'GlobalPolygonAreaHa >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_current_nature_ha\\t%.8f\\n", ',
    'GlobalCurrentNatureAreaHa >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_30pct_target_ha\\t%.8f\\n", ',
    'GlobalFinalNatureTargetHa >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_required_restoration_ha\\t%.8f\\n", ',
    'GlobalRestorationRequirementHa >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_restoration_tolerance_ha\\t%.8f\\n", ',
    'GlobalRestorationToleranceHa >> "', ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_final_nature_ha\\t%.8f\\n", ',
    'GlobalCurrentNatureAreaHa + sum {c in Cells} ',
    'RestorableAreaHa[c] * Selected[c] >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "global_final_nature_percent\\t%.8f\\n", ',
    '100 * (GlobalCurrentNatureAreaHa + sum {c in Cells} ',
    'RestorableAreaHa[c] * Selected[c]) / GlobalPolygonAreaHa >> "',
    ampl_path(summary_path), '";'
  ),
  paste0(
    'printf "COMPLETE\\n" > "',
    ampl_path(repair_completion_path), '";'
  ),
  "} else {",
  paste0(
    'printf "Stage 1 ended with positive network shortfall; ',
    'stages 2 and 3 were not run. Inspect ',
    basename(repair_shortfall_path), '.\\n";'
  ),
  "}"
)
write_lines_if_changed(run_lines, run_path)

problem_audit <- tibble(
  global_budget_only = TRUE,
  polygon_area_ha = global_budget$GlobalPolygonAreaHa,
  thirty_percent_target_ha = global_budget$GlobalFinalNatureTargetHa,
  current_nature_ha = global_budget$GlobalCurrentNatureAreaHa,
  required_new_restoration_ha =
    global_budget$GlobalRestorationRequirementHa,
  restoration_upper_tolerance_ha =
    global_budget$GlobalRestorationToleranceHa,
  maximum_new_restoration_ha =
    global_budget$GlobalMaximumRestorationHa,
  maximum_final_nature_percent =
    global_budget$GlobalMaximumFinalNaturePercent,
  candidate_cells = length(candidate_cells),
  allowed_action_binaries = nrow(allowed_actions),
  same_habitat_binaries = nrow(same_habitat_edges),
  candidate_rook_edges = nrow(candidate_edges),
  retained_existing_patches = length(existing_patches),
  retained_patch_full_national_area_ha =
    sum(patch_table$ExistingPatchAreaHa),
  retained_patch_area_inside_planning_ha =
    sum(patch_area_crosscheck$ExistingPatchAreaInsidePlanningHa),
  retained_patch_area_outside_planning_ha =
    sum(patch_area_crosscheck$ExistingPatchAreaOutsidePlanningHa),
  retained_full_patches_at_least_5000_ha = sum(
    patch_table$ExistingPatchAreaHa >= global_network_minimum_ha
  ),
  candidate_patch_links = nrow(cell_patch_links),
  required_network_certificates = length(required_networks),
  certificate_membership_binaries =
    nrow(eligible_network_cells) +
      2 * nrow(eligible_network_patches) +
      length(required_networks),
  certificate_flow_variables =
    nrow(eligible_network_arcs) +
      2 * nrow(eligible_network_cell_patch_edges) +
      nrow(eligible_network_patches),
  warm_start_selected_cells = nrow(warm_start_assignment),
  warm_start_label_repairs =
    sum(warm_start_assignment$forest_rule_repair),
  complete_feasible_mip_start_available =
    complete_feasible_mip_start_available,
  effective_warm_start_strategy = effective_warm_start_strategy,
  biodiversity_checkpoint_source = biodiversity_checkpoint_path,
  spatial_contiguity_bonus = spatial_contiguity_bonus,
  new_to_new_adjacency_weight = new_to_new_adjacency_weight,
  existing_to_new_adjacency_weight =
    existing_to_new_adjacency_weight,
  connected_existing_patch_bonus = connected_existing_patch_bonus,
  patch_connection_normalizer = patch_connection_normalizer,
  network_shortfall_tolerance_ha =
    network_shortfall_tolerance_ha,
  repair_change_tolerance_ha = repair_change_tolerance_ha,
  feasibility_time_limit_seconds =
    repair_feasibility_time_limit_seconds,
  minimum_change_time_limit_seconds =
    repair_minimum_change_time_limit_seconds,
  quality_time_limit_seconds = repair_quality_time_limit_seconds
)

readr::write_csv(
  problem_audit,
  file.path(problem_dir, "problem_size_and_policy_audit.csv")
)

export_audit <- tibble(
  file = c("model", "data", "complete slack start", "run"),
  path = c(model_path, dat_path, mip_start_path, run_path),
  exists = file.exists(c(model_path, dat_path, mip_start_path, run_path)),
  size_mib = file.info(
    c(model_path, dat_path, mip_start_path, run_path)
  )$size / 1024^2
)

print(knitr::kable(t(problem_audit), digits = 3))
knitr::kable(export_audit, digits = 3)
```

# Run AMPL

From PowerShell or Command Prompt at the `BiodiversitySignatureProjects`
project root:

```{powershell, eval=FALSE}
ampl Optimization/gudenaa_new_area_200m_global30_feasible_network_repair/gudenaa_new_area_200m_global30_feasible_network_repair.run
```

Or from an open AMPL session at the project root:

```{ampl, eval=FALSE}
include "Optimization/gudenaa_new_area_200m_global30_feasible_network_repair/gudenaa_new_area_200m_global30_feasible_network_repair.run";
```

The run performs three solves. Stage 1 always has a complete feasible start
because network shortfalls are explicit. Stage 2 runs only if every Stage-1
shortfall is zero; Stage 3 then keeps that minimum repair and restores the
biodiversity plus 1.2 contiguity objective. Inspect
`*_repair_stage_summary.tsv` first. If Stage 1 does not reach zero, no final
solution tables are written and `*_stage1_network_shortfall.tsv` identifies
the unresolved certificate(s).

# Post-solve independent audit and maps

Re-knit this notebook after AMPL finishes. Matching patch checkpoints are
reused. Post-processing runs only when all required solution tables are newer
than the generated model inputs.

```{r post-solve}
required_solution_paths <- c(
  solution_path,
  subarea_solution_path,
  network_cell_solution_path,
  network_patch_solution_path,
  network_root_solution_path,
  network_solution_summary_path,
  summary_path,
  repair_stage_summary_path,
  repair_completion_path
)

solution_is_current <- all(file.exists(required_solution_paths))
if (solution_is_current) {
  solution_time <- min(file.info(required_solution_paths)$mtime)
  input_time <- max(file.info(c(
    model_path,
    dat_path,
    mip_start_path,
    run_path
  ))$mtime)
  solution_is_current <- isTRUE(solution_time >= input_time)
}
if (solution_is_current) {
  completion_status <- trimws(readLines(
    repair_completion_path,
    warn = FALSE,
    n = 1L
  ))
  solution_is_current <- identical(completion_status, "COMPLETE")
}

if (!solution_is_current) {
  message(
    "No current repaired-network solution found. Run the generated .run file, ",
    "then re-knit this notebook."
  )
} else {
  selected_solution <- readr::read_tsv(
    solution_path,
    show_col_types = FALSE
  ) |>
    mutate(coarse_cell_id = as.integer(coarse_cell_id))
  ampl_subarea_solution <- readr::read_tsv(
    subarea_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(subarea_id = as.integer(subarea_id))
  network_cell_solution <- readr::read_tsv(
    network_cell_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(coarse_cell_id = as.integer(coarse_cell_id))
  network_patch_solution <- readr::read_tsv(
    network_patch_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(patch_id = as.integer(patch_id))
  network_root_solution <- readr::read_tsv(
    network_root_solution_path,
    show_col_types = FALSE
  ) |>
    mutate(patch_id = as.integer(patch_id))
  ampl_network_summary <- readr::read_tsv(
    network_solution_summary_path,
    show_col_types = FALSE
  )
  repair_stage_summary <- readr::read_tsv(
    repair_stage_summary_path,
    show_col_types = FALSE
  )
  solution_summary <- readr::read_tsv(
    summary_path,
    show_col_types = FALSE,
    col_types = cols(.default = col_character())
  )

  selected_ids <- selected_solution$coarse_cell_id
  stopifnot(
    nrow(selected_solution) > 0L,
    !anyDuplicated(selected_ids),
    all(selected_ids %in% candidate_cells),
    all(selected_solution$landuse %in% nature_landuses),
    setequal(ampl_network_summary$network_name, required_networks),
    setequal(network_root_solution$network_name, required_networks),
    all(ampl_network_summary$shortfall_ha <=
      network_shortfall_tolerance_ha),
    all(ampl_network_summary$certificate_built == 1L),
    "3_ecological_contiguity_quality" %in%
      repair_stage_summary$stage
  )

  # ---- Exact change from the successful notebook-07 footprint -----------
  footprint_repair_cells <- candidate_table |>
    select(coarse_cell_id, EligibleAreaHa) |>
    left_join(
      warm_start_assignment |>
        select(coarse_cell_id, warm_landuse = landuse),
      by = "coarse_cell_id"
    ) |>
    left_join(
      selected_solution |>
        select(coarse_cell_id, repaired_landuse = landuse),
      by = "coarse_cell_id"
    ) |>
    mutate(
      WarmSelected = as.integer(!is.na(warm_landuse)),
      RepairedSelected = as.integer(!is.na(repaired_landuse)),
      change_type = case_when(
        WarmSelected == 1L & RepairedSelected == 1L ~ "kept",
        WarmSelected == 1L & RepairedSelected == 0L ~ "dropped",
        WarmSelected == 0L & RepairedSelected == 1L ~ "added",
        TRUE ~ "unselected"
      )
    )

  footprint_repair_summary <- footprint_repair_cells |>
    group_by(change_type) |>
    summarise(
      cells = n(),
      area_ha = sum(EligibleAreaHa),
      .groups = "drop"
    ) |>
    arrange(match(
      change_type,
      c("kept", "dropped", "added", "unselected")
    ))

  computed_footprint_change_ha <- footprint_repair_cells |>
    filter(change_type %in% c("dropped", "added")) |>
    summarise(value = sum(EligibleAreaHa)) |>
    pull(value)
  reported_footprint_change_ha <- solution_summary |>
    filter(item == "warm_footprint_change_ha") |>
    pull(value) |>
    as.numeric()
  if (
    length(reported_footprint_change_ha) != 1L ||
      abs(
        computed_footprint_change_ha - reported_footprint_change_ha
      ) > 1e-5
  ) {
    stop("Warm-footprint change audit does not reconcile.", call. = FALSE)
  }

  readr::write_csv(
    footprint_repair_cells |>
      filter(change_type != "unselected"),
    file.path(problem_dir, "warm_footprint_repair_cells.csv")
  )
  readr::write_csv(
    footprint_repair_summary,
    file.path(problem_dir, "warm_footprint_repair_summary.csv")
  )

  # ---- Whole-polygon 30% budget and subarea reference table -------------
  independent_subarea_solution <-
    restorable_area_by_subarea_cell |>
    filter(coarse_cell_id %in% selected_ids) |>
    group_by(subarea_id) |>
    summarise(
      IndependentlyRestoredHa = sum(RestorableAreaHaBySubarea),
      .groups = "drop"
    ) |>
    right_join(subarea_budget, by = "subarea_id") |>
    mutate(
      IndependentlyRestoredHa = coalesce(IndependentlyRestoredHa, 0),
      FinalNatureAreaHa =
        CurrentNatureAreaHa + IndependentlyRestoredHa,
      FinalNaturePercent =
        100 * FinalNatureAreaHa / PolygonAreaHa,
      ReferenceThirtyPercentDifferenceHa =
        FinalNatureAreaHa - FinalNatureTargetHa
    ) |>
    arrange(subarea_id)

  if (max(abs(
    ampl_subarea_solution$restored_ha -
      independent_subarea_solution$IndependentlyRestoredHa
  )) > 1e-5) {
    stop("AMPL and R subarea totals disagree.", call. = FALSE)
  }

  independent_global_solution <-
    independent_subarea_solution |>
    summarise(
      IndependentlyRestoredHa = sum(IndependentlyRestoredHa)
    ) |>
    bind_cols(global_budget) |>
    mutate(
      FinalNatureAreaHa =
        GlobalCurrentNatureAreaHa + IndependentlyRestoredHa,
      FinalNaturePercent =
        100 * FinalNatureAreaHa / GlobalPolygonAreaHa,
      ExcessAboveThirtyPercentHa =
        FinalNatureAreaHa - GlobalFinalNatureTargetHa,
      PassesGlobalThirtyPercentMinimum =
        IndependentlyRestoredHa >=
          GlobalRestorationRequirementHa - 1e-6,
      PassesGlobalThirtyPercentMaximum =
        IndependentlyRestoredHa <=
          GlobalMaximumRestorationHa + 1e-6,
      PassesGlobalThirtyPercentBudget =
        PassesGlobalThirtyPercentMinimum &
          PassesGlobalThirtyPercentMaximum
    )

  if (!independent_global_solution$PassesGlobalThirtyPercentBudget[[1]]) {
    stop("Independent whole-polygon 30% audit failed.", call. = FALSE)
  }

  # ---- Production-forest destination rule -------------------------------
  forest_rule_audit <- selected_solution |>
    select(coarse_cell_id, landuse) |>
    left_join(
      candidate_table |>
        select(
          coarse_cell_id,
          ProductionForestHa,
          AgricultureHa,
          ProductionForestShare,
          IsProductionForestCell
        ),
      by = "coarse_cell_id"
    ) |>
    mutate(
      PassesForestRule =
        IsProductionForestCell == 0L |
        landuse %in% forest_landuses
    )
  if (!all(forest_rule_audit$PassesForestRule)) {
    stop("Production-forest destination audit failed.", call. = FALSE)
  }

  # ---- Recalculate the five AMPL certificates ---------------------------
  independent_certificate_summary <- tibble(
    network_name = required_networks,
    RequiredAreaHa = unname(network_minimums[required_networks])
  ) |>
    left_join(
      bind_rows(
        network_cell_solution |>
          left_join(
            network_cell_area,
            by = c("network_name", "coarse_cell_id")
          ) |>
          transmute(
            network_name,
            area_ha = coalesce(NetworkCellAreaHa, 0),
            member_cells = 1L,
            member_patches = 0L
          ),
        network_patch_solution |>
          left_join(
            network_patch_area,
            by = c("network_name", "patch_id")
          ) |>
          transmute(
            network_name,
            area_ha = coalesce(NetworkPatchAreaHa, 0),
            member_cells = 0L,
            member_patches = 1L
          )
      ) |>
        group_by(network_name) |>
        summarise(
          IndependentlyCertifiedAreaHa = sum(area_ha),
          member_cells = sum(member_cells),
          member_patches = sum(member_patches),
          .groups = "drop"
        ),
      by = "network_name"
    ) |>
    left_join(
      network_root_solution |>
        count(network_name, name = "root_count"),
      by = "network_name"
    ) |>
    mutate(
      PassesArea =
        IndependentlyCertifiedAreaHa >= RequiredAreaHa - 1e-6,
      PassesOneRoot = root_count == 1L,
      PassesNewRestoration = member_cells >= 1L,
      PassesCoarseCertificate =
        PassesArea & PassesOneRoot & PassesNewRestoration
    )

  if (
    anyNA(independent_certificate_summary) ||
    !all(independent_certificate_summary$PassesCoarseCertificate)
  ) {
    stop("Independent AMPL certificate audit failed.", call. = FALSE)
  }
  if (any(!network_cell_solution$coarse_cell_id %in% selected_ids)) {
    stop("A certificate contains an unselected cell.", call. = FALSE)
  }
  if (max(abs(
    ampl_network_summary$certified_area_ha[
      match(
        independent_certificate_summary$network_name,
        ampl_network_summary$network_name
      )
    ] - independent_certificate_summary$IndependentlyCertifiedAreaHa
  )) > 1e-5) {
    stop("AMPL and R certificate totals disagree.", call. = FALSE)
  }

  readr::write_csv(
    independent_subarea_solution,
    file.path(problem_dir, "independent_subarea_solution_audit.csv")
  )
  readr::write_csv(
    independent_global_solution,
    file.path(problem_dir, "independent_global_30pct_solution_audit.csv")
  )
  readr::write_csv(
    forest_rule_audit,
    file.path(problem_dir, "production_forest_solution_audit.csv")
  )
  readr::write_csv(
    independent_certificate_summary,
    file.path(problem_dir, "independent_coarse_certificate_audit.csv")
  )

  # ---- Exact 10 m restoration + full national patch network -------------
  fine_restoration <- terra::ifel(
    fine_eligible == 1 & coarse_id_on_fine %in% selected_ids,
    1,
    0
  )
  fine_restoration <- terra::mask(fine_restoration, fine_landuse)
  names(fine_restoration) <- "RestorationCells"

  final_nature <- terra::ifel(
    fine_nature == 1 | fine_restoration == 1,
    1,
    0
  )
  final_nature <- terra::mask(final_nature, fine_landuse)
  names(final_nature) <- "FinalNatureCells"

  # This clipped raster remains useful for mapping final nature inside the
  # planning polygon. It is not used to truncate national patch area in the
  # 5,000 ha calculation below.
  exact_component_raster <- terra::patches(
    final_nature,
    directions = 4,
    zeroAsNA = TRUE
  )
  names(exact_component_raster) <- "component_id"

  # Restoration is first split into exact 10 m rook components. National
  # existing patches then act as supernodes. This merges restoration pieces
  # that connect through the same current patch, including connections that
  # continue outside the planning boundary.
  restoration_component_raster <- terra::patches(
    fine_restoration,
    directions = 4,
    zeroAsNA = TRUE
  )
  names(restoration_component_raster) <- "restoration_component_id"

  restoration_component_summary <- zonal_sum_table(
    fine_restoration,
    restoration_component_raster,
    "restoration_component_id"
  ) |>
    transmute(
      restoration_component_id = as.integer(round(
        restoration_component_id
      )),
      RestorationAreaHa =
        RestorationCells * fine_cell_area_ha
    ) |>
    filter(RestorationAreaHa > 0) |>
    arrange(desc(RestorationAreaHa))

  exact_pair_multiplier <- expected_number_of_subareas + 1L
  restoration_component_subarea_zone <- terra::ifel(
    is.na(restoration_component_raster) |
      is.na(subarea_id_on_fine),
    NA,
    restoration_component_raster * exact_pair_multiplier +
      subarea_id_on_fine
  )

  restoration_component_by_subarea <- zonal_sum_table(
    fine_restoration,
    restoration_component_subarea_zone,
    "pair_id"
  ) |>
    mutate(
      restoration_component_id = as.integer(floor(
        pair_id / exact_pair_multiplier
      )),
      subarea_id = as.integer(round(
        pair_id -
          restoration_component_id * exact_pair_multiplier
      )),
      RestorationAreaHa =
        RestorationCells * fine_cell_area_ha
    ) |>
    filter(
      subarea_id %in% seq_len(expected_number_of_subareas)
    ) |>
    select(
      restoration_component_id,
      subarea_id,
      RestorationAreaHa
    ) |>
    arrange(subarea_id, desc(RestorationAreaHa))

  restoration_component_polygons <- terra::as.polygons(
    restoration_component_raster,
    aggregate = TRUE,
    values = TRUE,
    na.rm = TRUE
  )
  restoration_component_polygons$restoration_component_id <-
    as.integer(
      restoration_component_polygons$restoration_component_id
    )

  restoration_contact_buffers <- terra::buffer(
    restoration_component_polygons,
    width = contact_buffer_m
  )
  exact_contact_geometry <- terra::intersect(
    restoration_contact_buffers[, "restoration_component_id"],
    patches_near_domain[, "patch_id"]
  )

  if (nrow(exact_contact_geometry)) {
    exact_contact_geometry$contact_proxy_m2 <- terra::expanse(
      exact_contact_geometry,
      unit = "m",
      transform = FALSE
    )
    exact_restoration_patch_links <- exact_contact_geometry |>
      as.data.frame() |>
      transmute(
        restoration_component_id = as.integer(
          restoration_component_id
        ),
        patch_id = as.integer(patch_id),
        contact_proxy_m2 = as.numeric(contact_proxy_m2)
      ) |>
      group_by(restoration_component_id, patch_id) |>
      summarise(
        contact_proxy_m2 = sum(contact_proxy_m2),
        .groups = "drop"
      ) |>
      filter(contact_proxy_m2 >= minimum_contact_proxy_m2) |>
      arrange(restoration_component_id, patch_id)
  } else {
    exact_restoration_patch_links <- tibble(
      restoration_component_id = integer(),
      patch_id = integer(),
      contact_proxy_m2 = numeric()
    )
  }

  exact_linked_patch_table <- exact_restoration_patch_links |>
    distinct(patch_id) |>
    left_join(patch_dictionary_full, by = "patch_id")
  if (anyNA(exact_linked_patch_table$ExistingPatchAreaHa)) {
    stop("Exact audit found a patch missing from the national dictionary.")
  }

  exact_network_vertices <- bind_rows(
    restoration_component_summary |>
      transmute(
        vertex = paste0("r_", restoration_component_id),
        node_type = "restoration",
        node_id = restoration_component_id,
        GlobalNodeAreaHa = RestorationAreaHa
      ),
    exact_linked_patch_table |>
      transmute(
        vertex = paste0("p_", patch_id),
        node_type = "existing_patch",
        node_id = patch_id,
        GlobalNodeAreaHa = ExistingPatchAreaHa
      )
  )

  exact_network_edges <- exact_restoration_patch_links |>
    transmute(
      from = paste0("r_", restoration_component_id),
      to = paste0("p_", patch_id)
    ) |>
    distinct()

  exact_augmented_graph <- igraph::graph_from_data_frame(
    exact_network_edges,
    directed = FALSE,
    vertices = exact_network_vertices
  )
  exact_membership_raw <- igraph::components(
    exact_augmented_graph
  )$membership
  exact_membership <- tibble(
    vertex = names(exact_membership_raw),
    exact_network_component = as.integer(exact_membership_raw)
  )
  exact_network_vertices <- exact_network_vertices |>
    left_join(exact_membership, by = "vertex")

  exact_augmented_component_summary <- exact_network_vertices |>
    group_by(exact_network_component) |>
    summarise(
      restoration_components = sum(node_type == "restoration"),
      existing_national_patches =
        sum(node_type == "existing_patch"),
      RestorationAreaHa = sum(
        GlobalNodeAreaHa[node_type == "restoration"]
      ),
      ExistingPatchFullNationalAreaHa = sum(
        GlobalNodeAreaHa[node_type == "existing_patch"]
      ),
      GlobalNetworkAreaHa = sum(GlobalNodeAreaHa),
      IsAnchoredRestorationNetwork =
        restoration_components > 0 & existing_national_patches > 0,
      .groups = "drop"
    ) |>
    arrange(desc(GlobalNetworkAreaHa))

  exact_restoration_membership <- exact_network_vertices |>
    filter(node_type == "restoration") |>
    transmute(
      restoration_component_id = node_id,
      exact_network_component
    )
  exact_patch_membership <- exact_network_vertices |>
    filter(node_type == "existing_patch") |>
    transmute(
      patch_id = node_id,
      exact_network_component
    )

  exact_augmented_by_subarea <- bind_rows(
    restoration_component_by_subarea |>
      left_join(
        exact_restoration_membership,
        by = "restoration_component_id"
      ) |>
      transmute(
        exact_network_component,
        subarea_id,
        RestorationAreaHa,
        ExistingNatureAreaHa = 0
      ),
    patch_area_by_subarea_all |>
      semi_join(
        exact_linked_patch_table,
        by = "patch_id"
      ) |>
      left_join(exact_patch_membership, by = "patch_id") |>
      transmute(
        exact_network_component,
        subarea_id,
        RestorationAreaHa = 0,
        ExistingNatureAreaHa = ExistingPatchAreaHaBySubarea
      )
  ) |>
    group_by(exact_network_component, subarea_id) |>
    summarise(
      RestorationAreaHa = sum(RestorationAreaHa),
      ExistingNatureAreaHa = sum(ExistingNatureAreaHa),
      NetworkAreaInsideSubareaHa =
        RestorationAreaHa + ExistingNatureAreaHa,
      .groups = "drop"
    ) |>
    left_join(
      exact_augmented_component_summary |>
        select(
          exact_network_component,
          IsAnchoredRestorationNetwork
        ),
      by = "exact_network_component"
    ) |>
    arrange(subarea_id, desc(NetworkAreaInsideSubareaHa))

  exact_anchored_components <- exact_augmented_component_summary |>
    filter(IsAnchoredRestorationNetwork)

  if (nrow(exact_anchored_components)) {
    exact_global_row <- exact_anchored_components |>
      slice_max(GlobalNetworkAreaHa, n = 1, with_ties = FALSE) |>
      transmute(
        scope = "whole_polygon",
        subarea_id = NA_integer_,
        component_id = exact_network_component,
        RequiredAreaHa = global_network_minimum_ha,
        ExactFinalNatureAreaHa = GlobalNetworkAreaHa,
        ExactCurrentNatureAreaHa =
          ExistingPatchFullNationalAreaHa,
        ExactRestorationAreaHa = RestorationAreaHa,
        ExistingAreaInterpretation = "full_national_patch_area",
        PassesExact10m =
          ExactFinalNatureAreaHa >= RequiredAreaHa - 1e-6
      )
  } else {
    exact_global_row <- tibble(
      scope = "whole_polygon",
      subarea_id = NA_integer_,
      component_id = NA_integer_,
      RequiredAreaHa = global_network_minimum_ha,
      ExactFinalNatureAreaHa = 0,
      ExactCurrentNatureAreaHa = 0,
      ExactRestorationAreaHa = 0,
      ExistingAreaInterpretation = "full_national_patch_area",
      PassesExact10m = FALSE
    )
  }

  exact_largest_by_subarea <- exact_augmented_by_subarea |>
    filter(IsAnchoredRestorationNetwork) |>
    group_by(subarea_id) |>
    slice_max(
      NetworkAreaInsideSubareaHa,
      n = 1,
      with_ties = FALSE
    ) |>
    ungroup() |>
    transmute(
      subarea_id,
      component_id = exact_network_component,
      ExactFinalNatureAreaHa = NetworkAreaInsideSubareaHa,
      ExactCurrentNatureAreaHa = ExistingNatureAreaHa,
      ExactRestorationAreaHa = RestorationAreaHa
    )

  exact_network_goal_audit <- bind_rows(
    exact_global_row,
    tibble(
      subarea_id = seq_len(expected_number_of_subareas),
      scope = subarea_labels,
      RequiredAreaHa = subarea_network_minimum_ha
    ) |>
      left_join(exact_largest_by_subarea, by = "subarea_id") |>
      mutate(
        ExactFinalNatureAreaHa = coalesce(
          ExactFinalNatureAreaHa,
          0
        ),
        ExactCurrentNatureAreaHa = coalesce(
          ExactCurrentNatureAreaHa,
          0
        ),
        ExactRestorationAreaHa = coalesce(
          ExactRestorationAreaHa,
          0
        ),
        ExistingAreaInterpretation =
          "area_inside_this_subarea_only",
        PassesExact10m =
          ExactFinalNatureAreaHa >= RequiredAreaHa - 1e-6
      ) |>
      select(
        scope,
        subarea_id,
        component_id,
        RequiredAreaHa,
        ExactFinalNatureAreaHa,
        ExactCurrentNatureAreaHa,
        ExactRestorationAreaHa,
        ExistingAreaInterpretation,
        PassesExact10m
      )
  )

  readr::write_csv(
    restoration_component_summary,
    file.path(problem_dir, "exact_10m_restoration_components.csv")
  )
  readr::write_csv(
    restoration_component_by_subarea,
    file.path(
      problem_dir,
      "exact_10m_restoration_component_by_subarea.csv"
    )
  )
  readr::write_csv(
    exact_restoration_patch_links,
    file.path(
      problem_dir,
      "exact_10m_restoration_to_national_patch_links.csv"
    )
  )
  readr::write_csv(
    exact_augmented_component_summary,
    file.path(
      problem_dir,
      "exact_10m_augmented_network_components.csv"
    )
  )
  readr::write_csv(
    exact_augmented_by_subarea,
    file.path(
      problem_dir,
      "exact_10m_augmented_network_by_subarea.csv"
    )
  )
  readr::write_csv(
    exact_network_goal_audit,
    file.path(problem_dir, "exact_10m_network_goal_audit.csv")
  )

  # ---- Rasters for inspection and the next refinement -------------------
  selected_coarse <- coarse_cell_id
  selected_values <- rep(NA_integer_, terra::ncell(selected_coarse))
  selected_values[candidate_cells] <- 0L
  selected_values[selected_ids] <- 1L
  terra::values(selected_coarse) <- selected_values
  names(selected_coarse) <- "selected_200m"

  landuse_lookup <- tibble(
    warm_start_code = seq_along(nature_landuses),
    landuse = nature_landuses
  )
  destination_coarse <- coarse_cell_id
  destination_values <- rep(
    NA_integer_,
    terra::ncell(destination_coarse)
  )
  destination_values[selected_ids] <- landuse_lookup$warm_start_code[
    match(selected_solution$landuse, landuse_lookup$landuse)
  ]
  terra::values(destination_coarse) <- destination_values
  names(destination_coarse) <- "warm_start_landuse"

  destination_fine <- align_to_template(
    destination_coarse,
    fine_landuse,
    method = "near"
  )
  destination_fine <- terra::mask(
    destination_fine,
    fine_eligible_mask
  )

  terra::writeRaster(
    selected_coarse,
    file.path(problem_dir, "selected_coarse_cells_200m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES")
  )
  terra::writeRaster(
    fine_restoration,
    file.path(problem_dir, "repaired_solution_restoration_10m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    final_nature,
    file.path(problem_dir, "repaired_solution_final_nature_10m.tif"),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    exact_component_raster,
    file.path(
      problem_dir,
      "repaired_solution_final_nature_components_10m.tif"
    ),
    overwrite = TRUE,
    datatype = "INT4U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  terra::writeRaster(
    destination_fine,
    file.path(
      problem_dir,
      "repaired_solution_landuse_warm_start_10m.tif"
    ),
    overwrite = TRUE,
    datatype = "INT1U",
    gdal = c("COMPRESS=DEFLATE", "TILED=YES", "BIGTIFF=YES")
  )
  readr::write_csv(
    landuse_lookup,
    file.path(problem_dir, "warm_start_landuse_lookup.csv")
  )

  print(independent_global_solution)
  print(independent_subarea_solution)
  print(independent_certificate_summary)
  print(exact_network_goal_audit)

  if (!all(exact_network_goal_audit$PassesExact10m)) {
    warning(
      "The AMPL 200 m certificates pass, but one or more exact 10 m ",
      "network goals fail. Use this result as the targeted input to a ",
      "10 m refinement; do not report the failed goal as achieved."
    )
  }
}
```

# Quick plots

```{r quick-plots, eval=FALSE}
result_dir <- file.path(
  "Optimization",
  "gudenaa_new_area_200m_global30_feasible_network_repair"
)

plot(terra::rast(file.path(
  result_dir,
  "repaired_solution_landuse_warm_start_10m.tif"
)))

plot(terra::rast(file.path(
  result_dir,
  "repaired_solution_final_nature_components_10m.tif"
)))
```

# Interpretation safeguards

1. The repair model certifies connectivity on the 200 m graph. The independent
   `exact_10m_network_goal_audit.csv` is the authority for fine-resolution
   reporting.
2. Existing and newly restored nature both count toward the network goals,
   but their spatial ledgers differ. Global counts the complete national area
   of every reached existing patch plus selected restoration inside the
   planning polygon. Each Subarea counts only existing and restored hectares
   physically inside its own boundary. The 30% rule is one whole-polygon
   budget; subarea final-nature percentages are reported only as diagnostics.
3. Every certificate contains at least one selected restoration cell and is
   rooted at one existing-nature patch.
4. A component may satisfy several subarea certificates. Its area counted for
   `Subarea3`, for example, is only the part physically inside subarea 3, even
   if the connecting route crosses another subarea.
5. The patch-connection term is a preference. The model must connect at least
   one existing patch because of anchoring, but it is not forced to connect two
   or more if that would make the other objectives worse.
6. The production-forest rule is a 200 m approximation: a unit is classified
   as production forest when at least 50% of its eligible area is that source
   class. A later 10 m model can enforce source-specific transitions exactly.
7. The previous warm-up folder is read-only. This stage writes only to
   `Optimization/gudenaa_new_area_200m_global30_feasible_network_repair/`; no old
   `.dat` file or output folder should be deleted.
